---
title: "Exploratory Data Analysis Practice"
output:
  github_document
date: "4/12/17"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
theme_set(theme_minimal())
```

```{r}
library(rcfss)
data("scorecard")




type_violin <- function(var, var_label){
  scorecard %>%
    ggplot(aes_string("type", var)) +
      geom_violin(scale = "count", draw_quantiles = c(.5)) +
      labs(title = sprintf("Distribution of %s by Institution Type", var_label),
           subtitle = "Medians Indicated\n Area scaled to number of observations",
           y = sprintf("%s", var_label),
           x = "Institution Type")
}

type_violin("cost", "Cost per Year")
type_violin("admrate", "Admission Rate")
type_violin("satavg", "Average SAT Score")
type_violin("avgfacsal", "Average Faculty Salary")
type_violin("pctpell", "Percentage of Students with Pell Grants")
type_violin("comprate", "Completion Rate")
type_violin("firstgen", "Percentage of First Generation Higher Edu Students")
type_violin("debt", "Debt Upon Graduation")

scorecard %>% 
  ggplot(aes(state)) +
    geom_histogram(stat = "count") + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    labs(title = "Number of Higher Institutions Types By State",
         x = "State",
         y = "Count") + 
    facet_grid(type ~ .)