---
title: "in-class-EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Question: How does share of first generation students map onto debt?

> ggplot(data = scorecard) + geom_point(mapping = aes(x = firstgen, y = debt))

Question: What schools are represented in the data?

> ggplot(data = scorecard) + geom_bar(mapping = aes(x = type))

Okay, so mostly private colleges.

Let's see what categories show up where.

>scorecard$firstgen[scorecard$firstgen>0 & scorecard$firstgen<.20] <- ("0-20%")
> scorecard$firstgen[scorecard$firstgen>.20 & scorecard$firstgen<.40] <- ("20-40%")
> scorecard$firstgen[scorecard$firstgen>.40 & scorecard$firstgen<.60] <- ("40-60%")
> scorecard$firstgen[scorecard$firstgen>.60 & scorecard$firstgen<.80] <- ("60-80%")
> scorecard$firstgen[scorecard$firstgen>.80 & scorecard$firstgen<.1] <- ("80-100%")
> ggplot(data = scorecard) + geom_bar(mapping = aes(x = firstgen))

```{r}
messyplot
```
SOrry I'm slow at this
