---
title: "Oklahoma City School Data Explorer"
resource_files:
- schooldistrict.dbf
- schooldistrict.shx
- SchoolDistrict.txt
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: bootstrap
---
```{r globals, include = FALSE, verbose=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
#knitr::opts_chunk$set(verbose=FALSE, echo=FALSE, warning=FALSE, message=FALSE)
library(flexdashboard)
library(tidyverse)
library(stringr)
library(data.table)
library(plotly)
library(shiny)
library(leaflet)
library(rgdal)

#Importing Data Sets
###############################################################################
principals<-read.csv("principals.csv", header=TRUE)
contact_info<-read.csv("contactinfo.csv", header=TRUE)
enrollment<-read.csv("OKC_enrollment.csv", header=TRUE)
af_grades_df<-read.csv("AFGrades.csv", header=TRUE)
af_grades_df$SchoolName<-toupper(af_grades_df$SchoolName)
spi<-read_csv("SchoolPerformanceIndex.csv")
testScores<-read_csv("SchoolTestScores.csv")
testScores_wide<-gather(testScores, Prof, Prct, Advanced:Satisfactory, factor_key=TRUE)

###############################################################################
#setting list names:
###############################################################################
letterGrades<-unique(af_grades_df$LetteGrade)
gradeGroups<-as.list(unique(af_grades_df$LetterGradeGroup))
schools<- unique(enrollment$SiteName)
year<- unique(enrollment$Year)
grades<-as.list(unique(enrollment$Grade))
grade_groups<-as.list(unique(af_grades_df$LetterGradeGroup))
##############################################################################
#Wide Enrollment for graphs
###############################################################################
wEnrollment <- gather(enrollment, Race_Gender, Count, Hispanic_Male:Total, factor_key = TRUE)
wEnrollment<-filter(wEnrollment, Race_Gender !='Total')
wEnrollment <- separate(wEnrollment, Race_Gender,c("Race","Gender"),sep="_")

#leaflet
#shapefile
district<-readOGR("schooldistrict.shp",layer="schooldistrict")
okc_shp<-subset(district, SD_NAME =="Oklahoma City")
#school maker
school_geo<-read_csv("school_geo.csv")
school_lat <- school_geo$Latitude
school_lon <- school_geo$Longitude
school_name<-school_geo$Site
###############################################################################
#Expenditures
expenditures<-read.csv("okcpsExp.csv", header=TRUE)
#Revenues:
revenues<-read.csv("okcpsRev.csv", header=TRUE)

```
District Map
==============================================================================
	
Over the past several years Oklahoma has been consistently ranked as one of the lowest achieving states in terms of education. The lack of achievement is often recognized by Oklahoma residents, and even more so by education professionals. As such, there have been a groundswell of reform efforts throughout the state, specifically within Oklahoma City, the state’s largest school district. Efforts such as charter school expansion to curriculum implementation have been attempted over the past several years. Though these efforts are enacted with genuine desire to improve educational outcomes in the district, they operate on minimal information. Critical data points such as a school’s A-F Grade (the State’s evaluation of a school’s performance), student achievement on tested subject, student body demographics, and socio-economic factors are often hard to find or disorganized (if available at all). 

The Oklahoma City School Data Explorer seeks to centralize critical pieces of information so that groups ranging from parents of students, to non-profit groups and even policy makers can have a more precise understanding of the different variables that makeup Oklahoma City Public Schools. It should be noted that the data included in this dashboard are only for Oklahoma Public Schools and the district’s application schools, charter school and private schools have been left out due to lack of reportable data. 

Design:

The data explorer is divided into three different sections; “District Map”. “Oklahoma City”, and “School Explorer”. The first section, “District Map”, uses the leaflet api with shiny to map the address of each school in the district. Oklahoma City is very much divided along geographic landmarks that act as neighborhood borders. As such, mapping the districts location and giving each school a color corresponding to the most recent (2016) letter grade, gives the user a sense of where OKCPS schools are struggling. A resident of Oklahoma City would understand the geography of the area and be able to draw some assumption or infer questions about why schools may be succeeding or failing. Furthermore, the map has an overlay of the district boundary.
The second tab, “Oklahoma City”, provides a summary view of the district by showing the number of schools by letter grade (A through F) since the evaluation system began in 2013, as well as a bar chart that displays the frequency of each letter grade in the specified year. Furthermore, the dashboard provides a list of schools based on the filters on the sidebar. This was done to allow the user to compile lists for their own needs as well as be given a general sense of how the district has been performing over time

The third tab, “School Explorer”, is given to the user as a means for digging deeper into individual school’s performance. The page provides seven critical fields. The first four are value boxes: address, current principal, A-F grade (for the year specified by the user), and the grades served. These value boxes are meant to provide the user with a quick summary of information. The A-F grade is color coded to provide a sense of urgency for the school (i.e. if the school as an “F” letter grade than the box turns red). Below the value boxes are 4 graphs that show student demographics for a given year, over time, Free and Reduced Lunch overtime and test performance in a given year. 

Given that Oklahoma City is very much a “neighborhood” city, it is necessary to recognize the demographics for each school, as well as how those demographics are changing. Furthermore, Free and Reduced lunch (expressed as percent of students at the school) is often used in policy making decisions when determining school funding, as well as an indicator of poverty in the school’s neighborhood. The last panel is test scores. The test scores are filtered by grade and by subject, and depict the percentage of students scoring proficient, and of those scoring proficient, the percent of students that scored advanced. Like the Free and Reduced Lunch metric, test scores often are used to determine education policies such as funding but are also an indicator of school performance as to why a school might have received the letter grade in the value box at the top of the page. These graphs were chosen to intentionally allow the reader to explore the relationship between these variables. Several of the plots were made interactive using the Plotly package to allow users to focus on specific areas and timeframes. For the test scores, I used a drop down as to reduce the noise on the graph as well as allow the user to focus on a specific subject and grade to prevent confusion and misinterpretation. 

Analysis:

Based on the first tab, “District Map”, it is clear that there are very few “A” level schools in Oklahoma City and those that are successful schools are not relegated to a geographic area within the district boundaries. That sentiment also extends to “B” level schools. It is clear that of the 80+ schools in Oklahoma City Public Schools, a majority are receiving an “F” for their performance, with a higher concentration of failing schools in the north east and southern section of the district. 

The “Oklahoma City” summary of the district performance affirms the intuition made on the map. A majority of schools in Oklahoma City are receiving a failing grade. What’s more is that the amount of F schools appears to be increasing since the initial implementation of the A-F evaluation scheme was implemented in 2013, the first graph shows an increase from 35 to 40 F schools between 2013 and 2016. 
Finally, on the third tab, “School Explorer”, when comparing an “F” school, for instance Capitol Hill High School, we see that a very large percentage of the student population qualify for “Free and Reduced Lunch”, an indicator that the areas that Capitol Hill serves are high poverty areas. Comparing Capitol Hill to a school such as Classen High school of Advanced Studies, which scored an A+ in 2016, shows a (relatively) low number of students that qualify for Free and Reduced Lunch status, suggesting that Classen does not serve as many economically disadvantaged students. This comparison is further complicated by the demographic makeup of both schools. Capitol Hill High School has a majority of Hispanic students, representing over 1000 individual students, while Classen High School of Advanced Studies is a majority of white students. Comparing these attributes are important because the imply a relationship between demographics, economic status, and the School’s Performance beyond as a simple measurement of test scores. The visualization reveal that socio-economic makeup of a school can be a heavy influencer of the school’s performance outside of tradition metrics such as amount of funding or as a function of test scores alone. 

Conclusion:

The dashboard not only centralizes reporting information for Oklahoma City, but also allows for identifying possible relationships between variables. Visualizing this data can lead to asking important questions about how poverty, race and other socio-economic factors influence student performance. Furthermore, the data explorer is not targeted to solely identify “failing” schools but also to identify those schools that are being successful. Identifying the successes in the district can help influence the outcomes of different schools be recognizing and adapting best practices. From a policy lens, the dashboard can help summarize important findings that may simply be recognized but not quantified. For instance, it seems to be common knowledge that schools on Oklahoma City’s South and North East side are failing schools, but it is rarely quantified as the dashboard does. Furthermore, the dashboard proves that just because a school is “failing” does not mean it is relegated to the South or North East side of the city, when in fact there are several failing schools in the north and western parts of the district. Ultimately, the dashboards hopes to dispel misgivings or bias by quantifying and allowing users to explore the data for themselves. 


District Map
==============================================================================

```{r map, verbose=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
af_grades<-filter(af_grades_df, af_grades_df$Year == max(af_grades_df$Year))
getColor<-function(grades){
  sapply(af_grades$LetterGradeGroup, function(LetterGradeGroup){
    if(LetterGradeGroup=="A"){
      "green"
    } else if(LetterGradeGroup=="B"){
      "blue"
    } else if(LetterGradeGroup=="C"){
      "purple"
    } else if(LetterGradeGroup=="D"){
      "orange"
    } else if(LetterGradeGroup=="F"){
      "red"
    }
  })
}

icons<- awesomeIcons(
  icon='ios-close',
  iconColor='black',
  library='ion',
  markerColor=getColor(af_grades)
)

map <- leaflet()
map<-addTiles(map, 
              group="Oklahoma")
map<- addPolygons(map,
                  group='School Districts',
                  data=okc_shp,
                  weight=1,
                  fillColor="blue")

map <- addAwesomeMarkers(map,
                         lng=af_grades$Longitude,
                         lat=af_grades$Latitude,
                         icon = icons,
                         label = paste(af_grades$SchoolName,": ", af_grades$LetteGrade))
map<-addLegend(map, "topright", labels=c("A Schools","B School","C School","D School","F School"), col=c("green","blue","purple","orange","red"), title="2016 A-F Grade Colors")

map

```

Oklahoma City
==============================================================================

Column{.sidebar}
----------------------------------------------
###Select Grade Year
```{r}
selectInput(
  "gradeYear",
  "Select Year",
  choices=unique(af_grades_df$Year)
)
```

###Select School Level
```{r}

levels <-as.list(unique(af_grades_df$School_Level))
selectInput(
  "schoolLevel",
  "School Level",
  choices= c("All", levels)
)
```
```{r}
h3("School Superintendent:")
h4("Aurora Laura")
renderText({
})
```


Row
---------------------------------------------
###Number of Schools By Grade Group 2013-2016
```{r}
af_grades_2<-filter(af_grades_df, af_grades_df$LetterGradeGroup!='N')
af_grades_agg<-count(group_by(af_grades_2,Year,LetterGradeGroup))

renderPlotly({
  
  ggplot(af_grades_agg, aes(Year,y=n, color=factor(LetterGradeGroup)))+
    geom_point()+
    scale_x_discrete("Year")+
    geom_line()+
    xlab("Year")+
    ylab("Number of Schools")+
    labs(color="Grade")+
    theme(legend.position = "bottom")
})
```

###Number of Schools By Letter Grade (Single Year)
```{r}

af_grades2<- reactive({
  if(input$schoolLevel == 'All'){
    x<-filter(af_grades_df, af_grades_df$LetterGradeGroup!='N' & Year == input$gradeYear)
    x
  } else {
    x<-filter(af_grades_df, af_grades_df$LetterGradeGroup!='N' & Year == input$gradeYear & School_Level == input$schoolLevel )
    x 
  }
})

renderPlot({
  ggplot(af_grades2(), aes(LetterGradeGroup, fill=factor(LetterGradeGroup)))+
    geom_bar(stat="count")+
    xlab("Letter Grade Group")+
    ylab("Number of Schools") +
    #theme(legend.position = "bottom")+
    theme(legend.position="none")
    
})
```

Row{.tabset}
---------------------------------------------
###Query Grades
```{r}
gradeGroups <- reactive({
  if(input$schoolLevel == 'All'){
    x<-filter(af_grades_df, af_grades_df$LetterGradeGroup!='N' & Year == input$gradeYear)
    x<-x[,c("Year", "SchoolName", "IndexScore","LetteGrade")]
    x<-x[order(x$IndexScore, decreasing=TRUE),]
    x
    
  } else {
    x<-filter(af_grades_df, af_grades_df$LetterGradeGroup!='N' & Year == input$gradeYear & School_Level == input$schoolLevel )
    x<-x[,c("Year", "SchoolName", "IndexScore","LetteGrade")]
    x<-x[order(x$IndexScore, decreasing=TRUE),]
    x
  }
})
renderTable({
  gradeGroups()
})
```

School Explorer
==============================================================================

Column {.sidebar}
-------------------------------------
###School Chooser
```{r}
#School Name
selectInput(
  "schools",
  "Select School",
  choices=schools
)

selectInput(
  "year",
  "Select Year",
  choices=year,
  selected=2016
)

selectInput(
  "grade",
  "Select Grade",
  choices = c("All", grades),
  selected="All"
)

selectInput(
  "subject",
  "Select Test Subject (for EOI select any grade between 9 and 12)",
  choices=unique(testScores$Subject),
  selected=1
)
```

Row
-------------------------------------
###Address
```{r}
address<-reactive({
  x<-filter(contact_info, contact_info$Site==input$schools)
  y<-x$Address
})

renderValueBox({
  
  h2(valueBox(address(), 
         "School Address",
         icon="fa-address-book"))
})
```

### Principal
```{r}
principal<-reactive({
  x<-filter(principals, principals$SchoolName==input$schools)
  y<-x$Principal
})
renderValueBox({
  h2(valueBox(
    principal(),
    "School Principal",
    icon="fa-id-card"
         ))
})

```

###A-F Letter Grade
```{r}
afgrade<-reactive({
  x<-filter(af_grades_df, af_grades_df$SchoolName==input$schools & af_grades_df$Year == input$year)
  y<-x$LetteGrade
})


renderValueBox({
  h3(valueBox(
    afgrade(),
    "Letter Grade",
    icon="fa-pencil",
    color=ifelse(grepl(afgrade(),"A")==TRUE,"green",
                 ifelse(grepl(afgrade()==TRUE,"B")==TRUE,"#4286f4",
                        ifelse(grepl(afgrade()==TRUE,"C"),"#def416",
                               ifelse(grepl(afgrade(),"D"),"#f47616",ifelse(grepl(afgrade(),"F")==TRUE,"red","gray")))))
         ))
})
```

###Grades Served
```{r}
gradespan<-reactive({
  x<-filter(enrollment, enrollment$SiteName==input$schools)
  if(min(x$LowGrade) == -1){
    minGrade = "Pre-School"
  } else if (min(x$LowGrade) == 0){
    minGrade = "Kindergarten"
  } else{
    minGrade = min(x$LowGrade)
  }
  
  maxGrade = max(x$HighGrade)
  paste(minGrade," - ", maxGrade)
})

renderValueBox({
  
  valueBox(
    gradespan(),
    "Grades Served",
    icon="fa-bell",
    color="gray"
  )
  
})
```

Row
-------------------------------

###School Demographics (Single Year)
```{r}
demos<-reactive({
  if(input$grade=='All'){
    x<-filter(wEnrollment, wEnrollment$Year == input$year & wEnrollment$SiteName == input$schools)
    x<-aggregate(x$Count, by=list(x$Race),FUN=sum)
  } else if(input$grade != 'All'){
    x<-filter(wEnrollment, wEnrollment$Year == input$year & wEnrollment$SiteName == input$schools & wEnrollment$Grade == input$grade)
    x<-aggregate(x$Count, by=list(x$Race),FUN=sum)
  }
})

renderPlot({
  ggplot(demos(), aes(x=Group.1, y=x, fill=Group.1))+geom_col()+xlab("Demographic")+ylab("Number of Students")+labs(fill="Demographic")
})
```


###Demographics Over Time
```{r}
demos_ot<-reactive({
  if(input$grade=='All'){
    x<-filter(wEnrollment, wEnrollment$SiteName == input$schools)
    x<-aggregate(x$Count,  by=list("Year" = x$Year, "Demographic"=x$Race),FUN=sum)
    x
  } else if(input$grade != 'All'){
    x<-filter(wEnrollment, wEnrollment$SiteName == input$schools & wEnrollment$Grade == input$grade)
    x<-aggregate(x$Count, by=list("Year" = x$Year, "Demographic"=x$Race),FUN=sum)
    x
  }
})

renderPlotly({
  demos_ot<-demos_ot()
  
  ggplotly(ggplot(demos_ot(), 
                  aes(x=Year, y=x, group=factor(Demographic),
                      color=factor(Demographic)))+
             geom_point()+
             geom_line()+
    xlab("Year")+
    ylab("Num Students")+
      labs(color="Demographic"))
})
```

Row
-------------------------------
###Free and Reduced Lunch (FLR)
```{r}
frl<-reactive({
  x<-filter(spi, spi$Site == input$schools)
})

renderPlotly({
  ggplotly(
    ggplot(frl(), aes(x=YR, y=FRL))+
      geom_point()+
      geom_line()+
      scale_y_continuous(limits=c(0,1))
  )
})

```

###Student Performance - Testing Profecient and Advanced (EOI Tests are not grade specifc )
```{r}

scores<- reactive({
  if(input$grade == 'All'){
    NULL
  } else if(input$grade=="9"|input$grade=="10"|input$grade=="11"|input$grade=="12" | input$grade =="All"){
    filter(testScores_wide, testScores_wide$Site == input$schools & testScores_wide$Subject==input$subject)
  } else if (input$grade!="9"|input$grade!="10"|input$grade!="11"|input$grade!="12" | input$grade!="All") {
    filter(testScores_wide, testScores_wide$Site == input$schools & testScores_wide$Subject==input$subject & testScores_wide$Grade==input$grade)
  }
})
renderPlot({
  if(is.null(scores())){
    print("Please Select Grade and Subject")
  } else{
     ggplot(scores(), aes(x=YR, y=Prct, group=factor(Prof), fill=factor(Prof)))+geom_col(position = "dodge")+
      scale_y_continuous(limits=c(0,1))+labs(fill="Student Scores")
  }
  
})
```



