---
title: "In class exercise"
author: "Guangyu Liu"
date: "April 12, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rcfss)
library(tidyverse)
```

### What kind of college has more percentage of first generation students
```{r}
ggplot(scorecard) +
  geom_boxplot(mapping = aes(x = type, y = firstgen))

```
For-profit private universities have larger share of first generation student and more outliers, is that because of the sample size?
```{r}
ggplot(scorecard) +
  geom_bar(mapping = aes(type))

```
### More expensive univeristies have higher faculty salary?
```{r}
ggplot(scorecard, mapping = aes(x = cost, y = avgfacsal, color = type)) +
  geom_point() +
  geom_smooth(method = "lm")
```

- Public schools have the strongest correlation between cost and faculty salary, private for-profit schools have the weakest correlation
- Private for-profit schools have more variation from the fitted line

### 3. Good universities (with higher SAT score) are more difficult to get in?
```{r}
ggplot(scorecard, mapping = aes(satavg, admrate)) +
  geom_point() +
  geom_smooth() +
  facet_grid(~ type)
```
Private for-profit schools have higher admission rate in regardless of the SAT score, but there's so small sample size for this group.    
    
For private non-profit schools, there seems to be a negative correlation between SAT score and admission rate for those SAT scores above 1100.    
    
The same trend works for public schools whose SAT scores were above 900.

### 4. First generation student rely more on debt?
```{r}
ggplot(scorecard, mapping = aes(x = firstgen, y = debt)) +
  geom_boxplot(aes(group = cut_width(firstgen, 0.1)))
```
School of larger share of first generation students have students left school with fewer debt. That might because schools with larger share of first generation students also have low cost -- needs more exploration

