---
title: "Scorecard"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rcfss)
library(tidyverse)
data("scorecard")

```


```{r}
ggplot(data = scorecard)+
  geom_bar(mapping = aes(x=type))
```

In thie dataset, most schools are Private, Nonprofit. Then it follows the Public and Private, for-profit schools.

```{r}
ggplot(data = scorecard)+
  geom_smooth(mapping=aes(x=satavg, y=admrate))
```

From the plot above, the higher average SAT scroe comes with the lower admission rate.

```{r}
admrate_satavg <- lm(data=scorecard, admrate~satavg)
admrate_satavg
```

Using Linear model, we have the rusults above. It proves the previous conclusion

```{r}
ggplot(data = scorecard)+
  geom_density(aes(admrate, color=type ))
```

The admission rate of Private, for-profit schools is the highest. Private, nonprofit schools have the lowest admission rate.

```{r}
ggplot(data=scorecard)+
  geom_point(mapping = aes(x=avgfacsal, y=cost))+
  geom_smooth(mapping = aes(x=avgfacsal, y=cost))+
  facet_wrap(~type)
```

The faculties in private nonprofit schools get the highest pay. Also, those schools have the highest costs. There is a positive-relation between this two variables.

