---
title: "Biden"
author: "Longxuan Wang"
date: "April 9, 2017"
output: github_document
---

```{r setup, include=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
```

## Visualization

```{r csv}
biden <- read_csv("C:/Users/Longxuan/Downloads/Git HW/dataviz/assignments/data/biden.csv")

biden_sum <- biden%>%
  mutate(party=dem*2+rep)%>%
  group_by(female, party)%>%
  summarise(mean=mean(biden), sd=sd(biden), n=n())%>%
  mutate(CI=sd/sqrt(n)*1.96)
  

biden_sum%>%
  ggplot(aes(x=as.factor(female), y=mean, fill=as.factor(party)))+
  geom_bar(stat = "identity",position = position_dodge())+
  scale_x_discrete(breaks=c(0,1), labels=c("Male", "Female"))+
  scale_fill_discrete("Party affiliation",breaks=c(0,1,2), labels=c("Independent", "Republican", "Democrat"))+
  xlab("Gender")+
  ylab("Rating")+
  ggtitle("Feeling of warmth towards Biden by gender and party affiliation")+
  geom_errorbar(aes(ymax=mean+CI, ymin=mean-CI), position=position_dodge(0.9), size=1,                  width=0.4)
```

## Explanation

### Brief Intro
The dataset used for this graph contains information about individuals who rate their feeling of warmth towards Joe Biden on a scale of 0 to 100. There are 1807 individuals surveyed in total, and basic information about their gender, age, years of education, and political party affiliation are recorded. Among all those variables, I choose gender and party affiliation as my variables of interest, and I study how ratings of Biden vary between those groups. 

In order to make between-group comparisons, I calculate the group mean and make comparisons based on this mean statistics. Furthermore, I calculate the 95% confidence interval for the group mean. Since we have a moderately large sample size, we can safely assume that group mean will follow a normal distribution because of the law of large numbers. Their standard deviations are calculated by dividing sample standard deviation by the square root of sample size of each subgroup. Though I lost some information by using only the group mean, I compensated this loss through the confidence interval. So I believe by adding the confidence interval I add to the truthfulness of the graph. Readers will have a better ideal about the real distribution of the data and can make judgements about whether the differences we see in the graphs are truly significant or purely by chance.  

###The story
The story that I want to tell through this graph is two-folds. First, as evidenced by the difference in height between the color-coded bars, we see that across both male and female, rating of Biden is the highest among the democrats and the lowest among the republicans. Independents fall between the republicans and democrats. This results is to be expected, and this shows the huge influence of ideology on people of judgement of a person. 

On the other hand, if we compare male and females, we see that across all parties, females invariably rates Biden higher than males. There might be a lot of different explanations for this gender difference. The most obvious explanation is of course that women like Biden more than men. However, this explanation should not be taken too seriously without further supporting data as alternative explanations exist. For example, it is plausible to argue that women are in general more lenient/kind in their ratings of anyone. So this gender difference says nothing about Biden specifically. Likewise, it is possible that women rate men higher and vice versa, so this does not say too much about Biden either.  

###Graphical presentation
I choose bar plot for this comparison because it is intuitive and eye-catching. Compared with a line chart that use a single line to represent the rating, I believe that the volume of the bars make bar plot more easy to read even from afar with ease. It requires far less attention from the reader to recognize which bar is higher and which is lower. It will be harder to compare data points far away from each other on the x-axis if I use a line chart. Furthermore, bar plot looks better when color-coded because the colors are easier to read than a line chart. 

Furthermore, instead of stacking one gender on the other, I put them side by side. This is both for easiness of comparison as well interpretability. By putting them side by side it will be very clear to the readers which bar is higher and which is lower. Moreover, since the y-axis are the ratings, stacking one bar on top of another will make the y-axis uninterpretable.  
