---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set()
```

# Assignment 3: Interactive graphics with Gapminder

[Gapminder](http://www.gapminder.org/data/) compiles detailed indicators of global development. For this assignment, you should build an interactive visualization to tell a story about the data. **This requires you to develop both the visualization as well as a written narrative.**

* You do not need to analyze all 500+ variables. You can focus on a small set of variables (whichever ones you wish, though check the indicator's coverage for all the countries and the time period available)
* You can build either a highly linear or non-linear structure to your visualization. For instance, you can write up an analysis of the data as an R Markdown document/website with interactive visualizations incorporated throughout the analysis, or build a Shiny application with which users can explore the data. Choose whichever format you find more appealing. Note that if you design a non-linear interactive visualization, you still need to situate it using an introduction or text description (a la [America's Public Bible](http://americaspublicbible.org/)^[Though with far less detail]. See FA ch 9 for more details.)
* [`gapminder`](https://cran.r-project.org/web/packages/gapminder/index.html) only contains data on three variables. You will need to obtain additional indicators from Gapminder's [data page](http://www.gapminder.org/data/). Unfortunately they don't have a friendly API to access their database - each variable has to be downloaded manually as an Excel spreadsheet and imported into R. The following basic function should allow you to import a downloaded spreadsheet into R,^[You could also rewrite it to directly import from Google Sheets using the `googlesheets` package, making it fully reproducible.] and tidy:

```{r import-gapminder, message = FALSE}
library(tidyverse)
library(readxl)

import_gapminder <- function(filename, inc_name = NA){
  # import file
  indicator <- read_excel(filename)
  
  # rename first column to country, store indicator name for later
  inc_fullname <- names(indicator)[[1]]
  names(indicator)[[1]] <- "country"
  
  # tidy data frame and add indicator name as variable
  indicator <- indicator %>%
    gather(year, value, -1, convert = TRUE) %>%
    mutate(value = as.numeric(value),
           variable = ifelse(!is.na(inc_name), inc_name, inc_fullname)) %>%
    select(country, year, variable, value)
}

# use default name
(import_gapminder("data/indicator hiv estimated prevalence% 15-49 (1).xlsx"))

# use custom label
(import_gapminder("data/indicator hiv estimated prevalence% 15-49 (1).xlsx",
                  inc_name = "hiv_15_49"))
```

## Can I use some different data?

YES! If you have another dataset around which you want to build an interactive visualization, go ahead! Just let me know before you proceed.

## Grading

Your visualization will be evaluated using the following criteria:

* Is it truthful?
* Is it functional?
* Is it beautiful?
* Is it insightful?
* Is it enlightening?

In addition to the visualization and accompanying narrative, write an approximately 500 word explanation defending your design choices. This includes (but is not limited to):

* What is the story?
* Why did you select this graphical form?
* What is the role of interactivity? How does it enhance your ability to tell the story and communicate the data?

## Submission details

Submit your visualization and analysis as a single R Markdown document rendered using the `html_document` output format. If you build a Shiny app, deploy it on [shinyapps.io](http://www.shinyapps.io) and include a link to it in your submission. Include your design explanation in a separate document. Your folder in the repo should also contain a copy of the generated data file and documentation of the survey and the treatment/condition versions of your graphic.

Your submission is due by **1:30pm on Wednesday, May 10th**.



