---
title: "Curran_Thomas_HW2"
author: "Thomas Curran"
date: "4/27/2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
```{r libraries, echo=FALSE, include=FALSE}
library(gapminder)
library(tidyverse)
library(ggplot2)
library(gganimate)
library(data.table)
library(pander)
library(psych)
library(plotly)
results<-read_csv("SurveyResults.csv")
results_long<-as.data.frame(read_csv("SurveyResults.csv"))

results_long<-gather(results, question, response, q1:q4, factor_key = TRUE)

results_long<-as.data.frame(results_long)
```

####I) Introduction:
The advent of data visualization libraries in recent years has given researchers an unprecedented avenue to communicate data. The new avenue of visualization also has drawback, those drawbacks are compounded by the blitz of media that the average user encounters on a daily bassis. Ranging from television to social media, the average consumer is constantly bombarded with visual stimuli, and the most prominent of which are video or some sort of animation component to media. Given these conditions, it is imperative to understand whether or not movement (i.e. animation) can enhance or detract from the "story" that a visualization attempts to tell. 

Data visualization frameworks as outlined in Alberto Cairo's work suggest five primary channels to expressing a visualization (insert channel here). I propose an addition to those channels; movement. Adding movement as an expression of the data "story" can enable both deeper understanding, expression and functionality to the visualization. 

####II) Hypothesis:
The question underlying this study is: does animation help or hinder a visualization's ability to convey a data's "story" better or worse than a traditional static image?

The research was conducted with the notion that adding animation (i.e. movement) to a visualization allows for viewers to interpret the data's "story" just as well if not better than if the visualization was static. 

####III) Visaulization Design:

To test the hypothesis, I considered several important factors. First, the data had to tell a story over time. Incorporating time as a method for conveying a data's story is important in testing the hypothesis because motion can easily express time and directionality much in the way that traditional scatter plots do using lines of best fit or other methods. The visualizations used the built in data set in R called gapminder. the gapminder data set contains variables for country, continent (of the country), year measurement was taken, the country's average life expectency, the country's population in that year, and the gdp per capita. 

The tested visualizations did not include any sort of lines of best fit of regression lines because it compromised the integrity and spirit of the survey questions in that a line would to easily give away the "story", making it impossibly to distinguish the level of viewer understanding because it would not be able to distinguish whether they learned the story from the line in both graphs, or from th animation thereby defeating the point of the experiment. Furthermore, the data points plotted on both graphs are the same, but only a small subset of the data for countries available in the gapminder data set.

Other controls that were included in the creation of the visualizations were the colors of the data points, both graphs having the same set of data points and colors, as well as both graphs having had the legend removed. The legend was removed in an attempt to reduce bias from the person taking the survey. 

The visualization used the R packages ggplot2, gganimate (from github), tidyverse and dataset (where gapminder data set is stored). The data used for the visualizations was a pre-existing data set sourced from R cran. 

The first graph is the control, where it reflects the traditional, static visualization:

#####Image: 'Static Graph'
```{r static_graph}
data <- filter(gapminder, 
               country == 'Nepal' | 
                 country=='United States' |
                 country == 'Morocco'
)
p <- ggplot(data, aes(gdpPercap, lifeExp, color = country, frame = year)) +
  geom_point() +
  scale_x_log10() +
  theme(legend.position="none")+
  ggtitle("GDP and Life Expectancy")
p
```

The second graph mirrored the first graph, except this time instead showing every point at once, it showed each data point for every increase in the time interval.

####Image: 'Animated Graph'
```{r animated_graph, fig.show="animate"}
#note to view the animation in the HTML version I had to install ffmpeg via
#command line and homebrew with "brew install ffmpeg"
data <- filter(gapminder, 
               country == 'Nepal' | 
                 country=='United States' |
                 country == 'Morocco'
)
p <- ggplot(data, aes(gdpPercap, lifeExp, color = country, frame = year)) +
  geom_point() +
  scale_x_log10() +
  theme(legend.position="none")+
  ggtitle("GDP and Life Expectantcy")
gganimate(p)
```

####IV) Methodology

To test the hypothesis, a survey was developed using qualtrics. The survey displayed a brief text describing the data and then the visualization. The respondent was to answer 5 basic questions about the visualizations, giving an option of 3 responses per question.

The questions were as follows:

1)If a line of best fit was drawn through each set of points on the graph, which two colors would have the most similar slope?
2)Which country has the highest life expectancy
3)Which two colors showed a similar change in GDP
4)Which statement best reflects the relationship between life expectancy and gdp based on the graph?

The survey randomized whether or not the respondent was shown the static image or the animated image, but the questions remained the same regardless of which visualization was shown. the aim of these questions was to measure the influence of motion on the respondent's understanding of the graph

The survey was then deployed on Amazon's Mechanical Turk for random respondents to take the survey. Qualtrics then collected the data for analysis. 

####V) Results:
```{r results, echo=FALSE}

results <-data.table(results)
#mean for entire survey
total_mean<-mean(results$duration)
#mean for animated
animated<-filter(results, results$version == 'Animated')
animated_mean<-mean(animated$duration)
#mean for static
static<-filter(results, results$version == 'Static')
static_mean<-mean(static$duration)

means <- c(total_mean, animated_mean, static_mean)
label<- c("Average Time", "Average time to complete animted","Average time to complete static")

mean_table<-data.frame(Time=label, means=means)

pander(mean_table)


#density plot
ggplot(results,aes(x=results$duration))+geom_histogram(aes(y=..density..),
                                                                      binwidth = 5,
                                                                      color='black', fill='white')+
  geom_density(alpha=.2, fill='#FF6666')+
  ggtitle("Density plot of time to complete survey")

#grouped historygram - static vs animated
ggplot(results, aes(x=results$duration, fill=results$version))+
  geom_density(alpha=.3)+
    ggtitle("Density plot of time to complete survey: Animated vs Static")


#Response Bar Chart
response<-ggplot(results_long, aes(factor(question)))+
  geom_bar(position='dodge', aes(fill=factor(response)))+
  theme(legend.position = "bottom") +
  facet_wrap(~version)+
  ggtitle("Count of Response Choice Per Question: Animated vs Static")
response

```


####VI) Analysis

From the data collected, the two graphs overall seem to follow the same trends. Between the four questions, question 1, and question 4 appear to follow similar response patterns. However, question 2 and 3 are less homogenous. In question 3 (Which two colors showed a similar change in GDP), a majority of people answered 'Red and Blue' while in the animated version respondents were less sure, indicated by distribution of number of responses per answer. Similarly, question 2 on the on the animated graph shows that most respondents chose the first answer, where as the respondents in the static survey were not as sure.

After exploring the differences in the responses, I analyzed the differences in time taken to complete the survey. As shown in table 1, on average the respondents that answered questions on the static image took approximately 82 second (1.5 minutes), where as respondents to the animated survey took approximately 2 minutes to complete the survey. 

####VII) Conclusion:

From the results of the survey, it appears that adding animation does not have a significantly different effect on a viewer's ability to interpret and understand the data story. From the responses on the survey, it appears that the most the animation made it difficult to determine the relationship between the data points of different groups. Specifically, question three asked about the slope of each group of data points would make up. I believe that this is telling of cases where animation may not be useful. Specifically, animation may be useful to depict the changes in one specific country, but not helpful when comparing between groups. Furthermore, the animated graph has shown that repondents take longer to answer the questions. This may be because the respondents have to watch the animated graphic several times before being able to answer the same questions asked of static graph.  








