---
title: "MACS 40700 Assignment 1"
author: "Guangyu Liu"
date: "Due April 10, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(knitr)
```

## Part 1: Visualization Critique
In 2014, Gregor Aisch, Robert Gebeloff, and Kevin Quealy charted [the migration trend of Americans since 1900](https://www.nytimes.com/interactive/2014/08/13/upshot/where-people-in-each-state-were-born.html?action=click&contentCollection=The%20Upshot&region=Footer&module=WhatsNext&version=WhatsNext&contentID=WhatsNext&moduleDetail=undefined&pgtype=Multimedia#Illinois). These interactive graphs incorporate trends of both immigration and migration between states in the past century.      
    
### Truthful
To evaluate the truthfulness of the graph, two aspects need to be examined:    
    
- Is the data reliable?    
- Is the visualization reliable?    
    
These charts are based on Census microdata obtained from [IPUMS](https://www.ipums.org/) at the University of Minnesota Population Center. IPUMS provides census and survey data from around the world. It is reasonable to believe that the data source is reliable.    
    
At the bottom of the webpage, the authors indicate how foreign born residents and U.S. born residents but without a specific state are categorized in these graphs. Indicating data source and methodology makes these charts reproducible and reduces the probability of misrepresentation. In addition, the x-axis is divided into equal time intervals, the width of the bands are in proportion to the size of each group. The graphs accurately represent the data.     
    
### Functional
In these graphs, the United States is divided into four regions: West, South, Midwest, and Northeast, each is denoted by a particular color. The division of regions is clearly showed in the map on the left bottom. Readers can easily choose a state by clicking its position in the map.    
    
Each graph contains four types of data:     
    
- Time, which is mapped to the x-axis
- Size of migration, which is showed by the width of each ribbon
- Destination (or origin) of migration, which is showed by the color of each ribbon
- Relative rank of each migration destination (or origin), which is indicated by the relative position of the ribbons    
    
Taking the graph of diaspora out of California as an example. With each ribbon stacking over the other, at the first sight, these graphs reminded me the stacked continental GDP graph we critiqued in class 2. In the GDP chart, it is difficult to compare the size of bands in the middle, since their baselines are different. An essential improvement in the migration graphs is that the ribbons intertwine with each other, with the larger size always on the top. In addition, when mouse hovers over a ribbon, the graph shows the accurate percentage of California born residents lived in other regions/states. It is easy to compare the relative size of ribbons.    
    
However, these graphs have their weaknesses too. When the migration is mainly within region and changes frequently over time, the graph may be too complex to read. An extreme example is the graph of migration into Wyoming. Too many ribbons make it difficult to differentiate and obscure the main point. For instance, less than 3% Wyoming-born residents lived in Pennsylvania since 1900. It may be more appropriate to merge it into “other states in the Northeast” to simplify the graph, and help the audience focus on its main destination. Also, since destination states are represented separately, it is difficult to add up the percentages in each region and tell the general migration trend.    

### Beautiful
Generally speaking, these visualizations graph the trends of migration in an intuitively way. The pattern of ribbons shows whether the trend of migration is steady over time or changes rapidly. However, as mentioned above, as the number of migration destination / origin increases, the graph gets less concise.

### Insightful
These graphs plot the migration trend and the variation of destinations/origins in the past century, and reveal some patterns that are otherwise difficult to see. California, for instance, has long been the destination of American residents from other states, but this situation has changed since 1980s. The percentage of domestic transplants has shrunk significantly and was replaced by immigrants from other countries. Washington, D. C., the nation’s capital, provides another astonishing example that only 17% native residents stay in the nation’s capital and more than half of the residents in the past century were born outside of the capital. The graphs also show the interrelationship between moving in and moving out. For states with large percentage of moving in migration, native residents are more likely to stay in states as well.     
    
Again, since the graphs provide too much detail at the state level, it fails to show the general trend at the country level, which, in my opinion, might provide more insightful information. In addition, migration usually closely relates to the relative position of the origin and the destination. Holding other variables constant, people are more likely to move to a place near their origin. Mapping states to ribbons loses their spatial information and makes it difficult to show the effect of distance on migration.
    
### Enlightening
These graphs tend to show patterns of migration, which is an important issue nowadays. The change of population not only has great influence on a state’s economy, politics, and culture, but also influences the development of a country. As the authors mentioned, “the Western expansion, the movement of Southern blacks to Northern cities and the migration from the Rust Belt” are telling examples. These graphs provide a picture of where people came from and where people went, usually this is the first step to understand the historical change in a place.    
    
## Part 2: `ggplot2` and the grammar of graphics
### Design
The dataset `biden.csv` includes six variables. To explore the relationship between characteristics of the respondents and their attitudes toward Joe Biden, `biden`, the feeling thermometer is regarded as the dependent variable, while the other five variables, `female`, `age`, `dem`, `rep`, and `educ` are regarded as independent variables.   
  
**The first question is how different groups of people vary in their attitudes toward Biden.** A simple regression shows that `biden` is statistically significantly associated with `female`, `dem`, and `rep`. Female and democratic respondents like Biden more than others. The associations between `biden` and `age` and `educ` are not statistically significant.   
    
```{r, include = TRUE}
# Read data
Biden <- read.csv(file = "biden.csv")

# Regression analysis
reg <- lm(biden ~ female + age + educ + dem + rep, data = Biden) %>% 
  summary()
kable(reg$coefficients, format = "markdown", col.names = c("Coefficient", "Std. Error", "t", "p"),
  caption = "Regression output")
```
  
Thus, the designed visualization aims at showing the differences between gender and political beliefs. `dem` and `rep` are transformed into a single variable, `party`, with three categories: Democratic, Other, and Republican.    
    
**The second question is how to visualize `biden`, the dependent variable.** In `biden.csv`, the feeling thermometer is encoded as an ordinal variable, ranging from 0 to 100, with an interval equals 10. One option is viewing `biden` as a continuous variable. Using dotplot or violin chart, it is easy to see the distribution within each category. But what I am interested here is not the distribution, but which group likes Biden and which group dislikes Biden. For this reason, I transformed `biden` into a categorical variable, `biden_cat`, with four categories: extreme cold (0 ~ 25), moderate cold (26 ~ 50), moderate warm (51 ~ 75), and extreme warm (76 ~ 100).    
    
```{r, echo = FALSE}
# Recode categorical variables
attach(Biden)

## Recode female
Biden$female <- factor(Biden$female, levels = c(0, 1), labels = c("male", "female"))

## Recode party
Biden$party[dem == 1] <- "Democratic"
Biden$party[rep == 1] <- "Republican"
Biden$party[dem == 0 & rep == 0] <- "Other"
Biden$party <- factor(Biden$party, levels = c("Democratic", "Other", "Republican"))

## Recode educ
Biden$educ_cat <- cut(educ, 
                      breaks = c(0, 12, 13, 16, 17, Inf),
                      labels = c("less than HS", "HS degree", "some college", "college degree", "graduate"),
                      right = FALSE)

## Recode biden
Biden$biden_cat <- cut(biden,
                       breaks = c(-Inf, 26, 51, 76, Inf),
                       labels = c("extreme cold", "moderate cold", "moderate warm", "extreme warm"),
                       right = FALSE)

## Recode age
Biden$age_cat <- cut(age, 
                     breaks = c(-Inf, 33, 48, 63, 78, Inf), 
                     labels = c("18-32", "33-47", "48-62", "63-77", "78+"),
                     right = FALSE)
detach(Biden)
```

One problem of this transformation is that the `moderate warm` category contains two values (60, 70), while other three categories all contain three values. As mentioned above, since the aim of this visualization is not to compare the distribution of feeling thermometer WITHIN each group, but ACROSS groups (i.e., I am not going to say that in the democratic female category, there are more respondents having extreme warm attitude toward Biden than having moderate warm attitude, but to say that democratic female respondents are more likely to be extreme warm toward Biden than democratic male respondents), this problem will not bias the visualization.  
    
To display the relationship between three categorical variables -- `biden`, `female`, and `party` -- bar chart is a good choice. I mapped `female` to the x-axis, stacked four attitude categories of `biden_cat` in each bar, with each political party in a different facet. Since different gender and political party groups have different number of respondents, I calculated the percentage of each category of `biden_cat` within each respondent group, so that each bar has the same length. An initial graph is as below:    
    
```{r, include = TRUE}
# Percentage of attitude within each category
biden_perc <- Biden %>%
  group_by(female, party, biden_cat) %>% 
  summarize(freq = n()) %>% 
  mutate(perc = round(freq/sum(freq)*100, 2))

# Stacked barchart
ggplot(Biden, mapping = aes(x = female)) +
  geom_bar(aes(fill = biden_cat), position = "fill", width = 0.5) +
  labs(x = "", y = "Feeling Thermometer", 
       title = "Temp Viz1: Attitudes toward Joe Biden by gender and party") +
  theme(legend.position = "bottom") +
  facet_grid(~ party)
```
    
The problem of the above barchart is that although it is easy to compare the extreme warm and extreme cold categories, which lie above the bottom line and below the upper line respectively, it is difficult to compare the middle part of the bars across groups. For instance, it might take some effort to tell democratic females and other females, which group has a larger proportion of moderate warm attitude toward Biden. On the contrary, it is more reasonable to show which group likes Biden more (i.e., has a larger proportion of moderate warm and extreme warm) and which group dislikes Biden more (i.e., has a larger proportion of moderate cold and extreme cold). Therefore, I adjusted the positions of the bars, making them align at the neutral line (the line separates warm and cold) rather than the bottom line.    
    
```{r}
# Split biden_perc into two groups: cold and warm
cold <- filter(biden_perc, biden_cat == "extreme cold" | biden_cat == "moderate cold")
warm <- filter(biden_perc, biden_cat == "moderate warm" | biden_cat == "extreme warm")

# Negate the percentage of cold attitude, reverse order
cold$perc = -cold$perc
warm$biden_cat <- ordered(warm$biden_cat, levels = rev(levels(warm$biden_cat)))

# Visualization
bar_thermo <- ggplot() +
  aes(x = female, y = perc, fill = biden_cat) +
  geom_bar(data = warm, stat = "identity", width = 0.5) +
  geom_bar(data = cold, stat = "identity", width = 0.5) +
  geom_hline(yintercept = 0, color = "ivory3") +
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "Feeling Thermometer", 
       title = "Temp Viz2: Attitudes toward Joe Biden by gender and party") +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  facet_grid(~ party)
bar_thermo
```
    
**The third question is whether the insignificant variables should be included.** I recoded `educ` into five categories: less than high school, high school degree, some college, college degree, and graduate, and plotted each education category in a different facet, as showed in `Temp Viz3`. Since the difference between different education categories is minor, and more information might distract the audience from the main point of the story, it is more appropriate to exclude insignificant variables from the visualization.    
    
```{r}
# Percentage of attitude within each category
biden_perc <- Biden %>%
  group_by(female, party, educ_cat, biden_cat) %>% 
  summarize(freq = n()) %>% 
  mutate(perc = round(freq/sum(freq)*100, 2))

# Split biden_perc into two groups: cold and warm
cold <- filter(biden_perc, biden_cat == "extreme cold" | biden_cat == "moderate cold")
warm <- filter(biden_perc, biden_cat == "moderate warm" | biden_cat == "extreme warm")

# Negate the percentage of cold attitude, reverse order
cold$perc = -cold$perc
warm$biden_cat <- ordered(warm$biden_cat, levels = rev(levels(warm$biden_cat)))

# Visualization
bar_thermo <- ggplot() +
  aes(x = female, y = perc, fill = biden_cat) +
  geom_bar(data = warm, stat = "identity", width = 0.5) +
  geom_bar(data = cold, stat = "identity", width = 0.5) +
  geom_hline(yintercept = 0, color = "ivory3") +
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "Feeling Thermometer", 
       title = "Temp Viz3: Attitudes toward Joe Biden by gender and party") +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  facet_grid(party ~ educ_cat)
bar_thermo
```


Finally, in order to more clearly emphasize the comparison between like and dislike, I changed the filled color. Now, the cold categories are from the same color set, while the warm categories are from another color set. It is easy to tell that for each political party group, female respondents like Biden more than male respondents do. Democratic respondents like Biden most, Republican respondents dislike Biden most, while respondents of other political beliefs are in between. The final visualization and the complete code is as below.     

### Code and visualization
```{r, echo = TRUE}
# Read data
Biden <- read.csv(file = "biden.csv")

# Recode categorical variables
attach(Biden)

## Recode female
Biden$female <- factor(Biden$female, levels = c(0, 1), labels = c("male", "female"))

## Recode party
Biden$party[dem == 1] <- "Democratic"
Biden$party[rep == 1] <- "Republican"
Biden$party[dem == 0 & rep == 0] <- "Other"
Biden$party <- factor(Biden$party, levels = c("Democratic", "Other", "Republican"))

## Recode educ
Biden$educ_cat <- cut(educ, 
                      breaks = c(0, 12, 13, 16, 17, Inf),
                      labels = c("less than HS", "HS degree", "some college", "college degree", "graduate"),
                      right = FALSE)

## Recode biden
Biden$biden_cat <- cut(biden,
                       breaks = c(-Inf, 26, 51, 76, Inf),
                       labels = c("extreme cold", "moderate cold", "moderate warm", "extreme warm"),
                       right = FALSE)

## Recode age
Biden$age_cat <- cut(age, 
                     breaks = c(-Inf, 33, 48, 63, 78, Inf), 
                     labels = c("18-32", "33-47", "48-62", "63-77", "78+"),
                     right = FALSE)
detach(Biden)

# Percentage of attitude within each category
biden_perc <- Biden %>%
  group_by(female, party, biden_cat) %>% 
  summarize(freq = n()) %>% 
  mutate(perc = round(freq/sum(freq)*100, 2))

# Split biden_perc into two groups: cold and warm
cold <- filter(biden_perc, biden_cat == "extreme cold" | biden_cat == "moderate cold")
warm <- filter(biden_perc, biden_cat == "moderate warm" | biden_cat == "extreme warm")

# Negate the percentage of cold attitude, reverse order
cold$perc = -cold$perc
warm$biden_cat <- ordered(warm$biden_cat, levels = rev(levels(warm$biden_cat)))

# Visualization
bar_thermo <- ggplot() +
  aes(x = female, y = perc, fill = biden_cat) +
  geom_bar(data = cold, stat = "identity", width = 0.5) +
  geom_bar(data = warm, stat = "identity", width = 0.5) +
  geom_hline(yintercept = 0, color = "ivory3") +
  scale_fill_manual(name = "", values = c("extreme warm" = "indianred4", "moderate warm" = "indianred1", "moderate cold" = "ivory3", "extreme cold" = "ivory4"),
                    breaks = c("extreme warm", "moderate warm", "moderate cold", "extreme cold")) +
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "Feeling Thermometer", 
       title = "Final Visualization: Attitudes toward Joe Biden by gender and party") +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  facet_grid(~ party)
bar_thermo
```
    
