---
title: "Day 6 - Exploratory data analysis"
output:
  html_document:
    highlight: pygments
    theme: readable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

```{r packages}
library(tidyverse)
library(broom)
library(rcfss)
```

# Get and clean scorecard data

```{r get-data}
# get data
data("scorecard")
str(scorecard)

# normalize dollar variables into thousands of dollars - easier to plot and read on the screen
scorecard <- scorecard %>%
  mutate_each(funs(. / 1000), cost, avgfacsal, debt)
```

# What's up with admit rate

```{r admit}
ggplot(scorecard, aes(admrate)) +
  geom_histogram(binwidth = 1)
```

# Eh, I'm bored - what about cost?

```{r cost}
ggplot(scorecard, aes(cost)) +
  geom_histogram()

ggplot(scorecard, aes(cost, color = type)) +
  geom_density()
```

* Different distributions for cost by type
    * Public skew right (cheapest)
    * Then private for-profit and finally private nonprofit
    
## What drives this difference in cost by college type?

### Maybe faculty salaries

```{r}
ggplot(scorecard, aes(avgfacsal, cost)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ type)

ggplot(scorecard, aes(avgfacsal, color = type)) +
  geom_density()
```

* Public universities tend to have higher average faculty salaries compared to private for profits, yet charge a lot less - clearly the money isn't going towards faculty salaries at these schools

### What about Pell grants?

Pell grants are free money given to low-income students to subsidize tuition. Perhaps private for-profits target these students to capitalize on the free money.

```{r}
ggplot(scorecard, aes(pctpell, color = type)) +
  geom_density()

ggplot(scorecard, aes(pctpell, cost)) +
  geom_point(alpha = .2) +
  geom_smooth()

ggplot(scorecard, aes(pctpell, cost, color = type)) +
  geom_point(alpha = .2) +
  geom_smooth()
```

* Generally as percent of Pell recipients increases, costs go down
* Overall though, private institutions have higher costs even after controlling for Pell grant students

How about using a regression model to test this?

```{r}
cost_pell <- lm(cost ~ pctpell, data = scorecard)
tidy(cost_pell)
glance(cost_pell)

cost_pell_type <- lm(cost ~ pctpell + type, data = scorecard)
tidy(cost_pell_type)
glance(cost_pell_type)

cost_pell_type_x <- lm(cost ~ pctpell * type, data = scorecard)
tidy(cost_pell_type_x)
glance(cost_pell_type_x)

# compare residuals for interactive and non-interactive models
bind_rows(plain = augment(cost_pell_type),
          x = augment(cost_pell_type_x), .id = "model") %>%
  ggplot(aes(.resid, color = model)) +
  geom_density()
```

* Larger $R^2$ with the interactive model, plus smaller residuals (larger clustering around 0)
* Could do some statistical tests (ANOVA, $\chi^2$, etc.) to confirm if the models really are statistically distinguishable from one another, but EDA is a quick check - plus more directly intuitive to non-stats trained people


