---
title: "Assignment 3"
output: github_document
---
# Armed Forces Personnel

This report focueses on the number of armed forces personnel all over the world.

Armed forces personnel are active duty military personnel, including paramilitary forces if the training, organization, equipment, and control suggest they may be used to support or replace regular military forces. Labor force comprises all people who meet the International Labour Organization's definition of the economically active population. Note: Data for some countries are based on partial or uncertain data or rough estimates.

# Import the data

I used two datasets from the website of "gapminder". The first one includes the percentage of armed forces personel of total labour force. The second one inclueds the total number of armed forces personnel.

```{r, warning=FALSE, message=FALSE}

library(rworldmap)
library(tidyverse)
library("ggmap")
library(maptools)
library(maps)
library(graphics)
library(readxl)
library(readr)
library(gapminder)

  # import file
  percent <- read_excel("indicator army_percent.xlsx")

  # rename first column to country, store indicator name for later
  inc_fullname <- names(percent)[[1]]
  names(percent)[[1]] <- "country"
  
  inc_name=NA
  # tidy data frame and add indicator name as variable
  percent <- percent %>%
    gather(year, value, -1, convert = TRUE) %>%
    mutate(value = as.numeric(value),
           variable = ifelse(!is.na(inc_name), inc_name, inc_fullname)) %>%
    select(country, year, variable, value)


  # import file
  total <- read_excel("indicator army_total.xlsx")

  # rename first column to country, store indicator name for later
  inc_fullname <- names(total)[[1]]
  names(total)[[1]] <- "country"
  
  inc_name=NA
  # tidy data frame and add indicator name as variable
  total <- total %>%
    gather(year, value, -1, convert = TRUE) %>%
    mutate(value = as.numeric(value),
           variable = ifelse(!is.na(inc_name), inc_name, inc_fullname)) %>%
    select(country, year, variable, value)


percent <- mutate(percent, percent=value) %>%
  select(-3,-4)

total <- mutate(total, total=value) %>%
  select(-3,-4)

all <- left_join(percent, total) 
all <- left_join(all, gapminder)
  
```

# The Status quo

The latest data I can get is the data of year 2008. I draw two tables below. The first one is the Top 10 in terms of Percent of labour force (2008). The second one is Top 10 in terms of Total Amount (2008).

## Armed forces personnel over labour force

```{r}
data2008_percent <- filter(all, year==2008) %>%
arrange(desc(percent)) %>%
  select(1:3) %>%
  as.data.frame() %>%
  head(n=10) 

knitr::kable(data2008_percent, caption="'Top 10 in terms of Percent (2008)'")
```

```{r}
ggplot(data=data2008_percent)+
  geom_bar(mapping = aes(x=country, y=percent, fill=country), stat = "identity") +
  ggtitle("Top 10 countries in terms of Percent (2008)")

```


In order to better show the top 10 countries, I draw them on a world map.

```{r, warning=FALSE, message=FALSE}

visited <- data2008_percent$country
ll.visited <- geocode(visited)
visit.x <- ll.visited$lon
visit.y <- ll.visited$lat

mp <- NULL
mapWorld <- borders("world", colour="gray50", fill="gray50") # create a layer of borders
mp <- ggplot() +   mapWorld

#Now Layer the cities on top
mp <- mp+ geom_point(aes(x=visit.x, y=visit.y) ,color="red", size=0.5) 
mp
```

Obviously, Mid-Eastern is quite an unstable area. The Arabian countries and Israel are always in tensions. So they tend to keep a largy armed force. The case is same for North Korea. North Korea is prepared a war with the South and it is developing nuclear weapons.

The outlier is Singapore. It also has a large armed force in terms of the percentage.

## The total number of armed forces personnel

```{r, warning=FALSE, message=FALSE}
data2008_total <- filter(all, year==2008) %>%
arrange(desc(total)) %>%
  select(-3) %>%
  select(1:3) %>%
  as.data.frame() %>%
  head(n=10)

knitr::kable(data2008_total, caption='Top 10 in terms of Total Amount (2008)')

data2008_pop <- filter(all, year==2008) %>%
arrange(desc(total)) %>%
  select(-3) %>%
  select(1:3) %>%
  as.data.frame() %>%
  head(n=10)

knitr::kable(data2008_pop, caption='Top 10 in terms of Total Amount (2008)')


```

```{r}
ggplot(data=data2008_total)+
  geom_bar(mapping = aes(x=country, y=total, fill=country), stat = "identity") +
  ggtitle("Top 10 countries in terms of total number of armed forces personnel (2008)")

```


Based on general knowledge, I doubt if there is a correlation between population and total number of armed forces. So I draw a table of top 10 countries in terms of polulation in 2007.

```{r}

data2007_pop <- filter(all, year==2007) %>%
arrange(desc(pop)) %>%
  select(c(1,2,7)) %>%
  as.data.frame() %>%
  head(n=10)

knitr::kable(data2007_pop, caption='Top 10 countries in terms of Popluation (2007)')

```

There are five countries listed in both the tables. So I conclude that the countries with larger population have more military personnels.

Again, I draw them on a world map.

```{r, warning=FALSE, message=FALSE}
visited <- data2008_total$country
ll.visited <- geocode(visited)
visit.x <- ll.visited$lon
visit.y <- ll.visited$lat

mp <- NULL
mapWorld <- borders("world", colour="gray50", fill="gray50") # create a layer of borders
mp <- ggplot() +   mapWorld

#Now Layer the cities on top
mp <- mp+ geom_point(aes(x=visit.x, y=visit.y) ,color="red", size=0.5) 
mp


```

In terms of total amount of armed forces personnel, the top 10 countries are scattered distributed across the world.

# Correlations

Now, I'm trying to depict the correlations between different variables.

## GDP and the Armed Force

Firstly, I draw the smooth lines indicating the relations between per capita GDP, the percentage of armed forces personnel and the total amount of armed forces personnel. This is to show the rough relations between those variables.

```{r, warning=FALSE, message=FALSE}
data_gdp_2007 <- filter(all, is.na(gdpPercap)==F) %>%
  filter(year==2007)

ggplot(data = data_gdp_2007)+
  geom_point(mapping = aes(x=gdpPercap, y=percent), na.rm = T)+
  geom_smooth(mapping = aes(x=gdpPercap, y=percent), na.rm = T)+
  ggtitle("Percentage and per capita GDP")

ggplot(data = data_gdp_2007)+
  geom_point(mapping = aes(x=gdpPercap, y=total), na.rm = T)+
  geom_smooth(mapping = aes(x=gdpPercap, y=total), na.rm = T)+
  ggtitle("Total amount and per capita GDP")
```

To make it more significant in the graph, I adopt log to draw the lines again.

```{r}
ggplot(data = data_gdp_2007)+
  geom_point(mapping = aes(x=gdpPercap, y=log(percent)), na.rm = T)+
  geom_smooth(mapping = aes(x=gdpPercap, y=log(percent)), na.rm = T)+
  ggtitle("Percentage(log) and per capita GDP")

ggplot(data = data_gdp_2007)+
  geom_point(mapping = aes(x=gdpPercap, y=log(total)), na.rm = T)+
  geom_smooth(mapping = aes(x=gdpPercap, y=log(total)),, na.rm = T)+
  ggtitle("Total amount(log) and per capita GDP")
```

For most countries both the percentage and total amount of armed forces have are positive correlated with per capita GDP.

However, when per capita GDP is high enough, the percentage and total amount of armed forces are tend to maintain at a fixed level.

To better illustrate the data, I use OLS method to analyze it.

```{r}
percentage_gdp_lm <- lm(percent ~ gdpPercap, data = data_gdp_2007)
summary(percentage_gdp_lm)
```

```{r}
total_gdp_lm <- lm(total ~ gdpPercap, data = data_gdp_2007)
summary(total_gdp_lm)
```


```{r, warning=FALSE, message=FALSE}
data2007 <- filter(all, year==2005)
ggplot(data=data2007)+
geom_point(mapping = aes(x=gdpPercap, y=percent), na.rm = T)+
  geom_smooth(mapping = aes(x=pop, y=percent), na.rm = T)
```

## GDP, Population and the Armed Forces Personnel

1.Regress the total number of Armed Forces Personnel on population and per capita GDP.
2.Regress the percentage of Armed Forces Personnel over total labor force on population and per capita GDP.

```{r}
percentage_gdp_pop_lm <- lm(percent ~ pop * gdpPercap, data = data_gdp_2007)
summary(percentage_gdp_pop_lm)
```

```{r}
total_gdp_pop_lm <- lm(total ~ pop * gdpPercap, data = data_gdp_2007)
summary(total_gdp_pop_lm)
```

Results:
1.regress the percentage of Armed Forces Personnel over total labor force on population and per capita GDP
The coefficient of population and per capita GDP are -7.292e-10 and 5.792e-06 respectively. 
2.Regress the Total number of Armed Forces Personnel over total labor force on population and per capita GDP.
The coefficient of population and per capita GDP are  1.979e-03 and -1.589e+00 respectively. 

It's very interesting that I get two seemingly opposite results. 

The countries with larger populations tend to have more soldiers. That is quite obvious. The higher per capita GDP normally indicates a higher level of technology. Those countries can make up their shortage in total numbers of soldiers by their advanced technology.

As for the per percentage of Armed Forces over total labor force, 
