---
title: "Day 15 - Network visualization"
output:
  html_document:
    highlight: pygments
    theme: readable
    toc: yes
    toc_float: yes
    code_folding: show
    includes:
      in_header: "header_include_d3.html"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE)
```

# Objectives

* Define core concepts of network theory
* Summarize [*The Rise of Partisanship in the U.S. House of Representatives*](http://www.mamartino.com/projects/rise_of_partisanship/) and it's use of netowrk analysis
* Review the [*Love Actually*](http://varianceexplained.org/r/love-actually-network/) example
* Identify core packages in R for network analysis
* Demonstrate major types of network visualizations
    * Node-edge diagrams
    * Adjacency matrix
    * Trees and hierarchies (aka dendrograms)

```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(modelr)
library(forcats)
library(ggmap)
library(plotly)

options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())
```



# Acknowledgments {.toc-ignore}

* [Analyzing networks of characters in 'Love Actually'](http://varianceexplained.org/r/love-actually-network/)

# Session Info {.toc-ignore}

```{r cache = FALSE}
devtools::session_info()
```


