---
title: "Day 12 - Interactivity (cont.)"
output:
  html_document:
    highlight: pygments
    theme: readable
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE)
```

# Objectives

* Introduce `plot_ly()` generating interactive graphs
* Introduce `shiny` for interactive visualizations
* Introduce `flexdashboard` for creating interactive dashboards

```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(modelr)
library(forcats)
library(haven)
library(plotly)

options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())
```

# Plotly

We previously discussed using `plotly` and the `ggplotly()` function to modify existing `ggplot()` objects. However we can also build plotly.js objects directly using the `plot_ly()` function. It is based on the layered grammar of graphics, but instead it directly calls components of the plotly.js JavaScript library to generate graphs. Originally it was intended to be used for plot types not supported by `ggplot2` (like 3d graphs), but has been extended so that you could use it entirely instead of `ggplot2`.

```{r}
txhousing

# ggplot
p <- ggplot(txhousing, aes(date, median)) +
  geom_line(aes(group = city), alpha = 0.2)
ggplotly(p)

# initialize plotly object
p <- txhousing %>%
  group_by(city) %>%
  plot_ly(x = ~date, y = ~median)
p

# specify geometric relation (i.e. what kind of geom to draw?)
# add a line highlighting houston
add_lines(
  # plots one line per city since p knows city is a grouping variable
  add_lines(p, alpha = 0.2, name = "Texan Cities", hoverinfo = "none"),
  name = "Houston", data = filter(txhousing, city == "Houston")
)

# redone completely with piping
allCities <- txhousing %>%
  group_by(city) %>%
  plot_ly(x = ~date, y = ~median) %>%
  add_lines(alpha = 0.2, name = "Texan Cities", hoverinfo = "none")

allCities %>%
  filter(city == "Houston") %>%
  add_lines(name = "Houston")
```

We can add "layers" by using the piped notation used by `dplyr` and other `tidyverse` packages (not `+` in `ggplot2`).

## More support for plotly

* [Plotly R library](https://plot.ly/r/)
* [plotly for R](https://cpsievert.github.io/plotly_book/) - good introductory book explaining some of the fundamental concepts of plotly

# Shiny

See [here](http://cfss.uchicago.edu/shiny001_abc.html).

# `flexdashboard`

[`flexdashboard`](http://rmarkdown.rstudio.com/flexdashboard/) is a package for R that allows you to create visual dashboards. You can incorporate tables, graphs (both static and dynamic), text, and all sorts of components to provide insight and analysis. Layouts are customizable by dividing the page into columns and rows.

The entire dashboard is basically one large R Markdown document, so you can use the same syntax and code/output/analysis format that you've used before (though generally hiding the code with `echo = FALSE`).

## Review layout and components

* [Layout](http://rmarkdown.rstudio.com/flexdashboard/using.html#layout)
* [Components](http://rmarkdown.rstudio.com/flexdashboard/using.html#components)

## Examples of flexdashboards

* [NBA scoring](https://beta.rstudioconnect.com/jjallaire/htmlwidgets-d3heatmap/htmlwidgets-d3heatmap.html)
* [ggplotly: various examples](https://beta.rstudioconnect.com/jjallaire/htmlwidgets-ggplotly-examples/htmlwidgets-ggplotly-examples.html)

## Shiny with `flexdashboard`

One concern with a traditional dashboard is that while it may be "interactive", it is not dynamic - that is, you cannot update the dashboard by changing parameters or underlying data values. Shiny applications allow you to do so, but they are very clunky to write due to the deeply nested structure of the code. Instead, we can combine the two approaches by generating a Shiny dashboard. In our header, we simply add `runtime: shiny` to convert the standard dashboard to a Shiny dashboard.

### Comparison

* [Pure Shiny app](age-rule/app.R)
* [Shiny flexdashboard](age-rule/age-rule.Rmd)

## Examples of Shiny flexdashboards

* [`ggplot2` Diamonds Explorer](https://jjallaire.shinyapps.io/shiny-ggplot2-diamonds/)
* [`ggplot2` Brushing](https://jjallaire.shinyapps.io/shiny-ggplot2-brushing/)

# Session Info {.toc-ignore}

```{r cache = FALSE}
devtools::session_info()
```


