---
title: "Exploratory data analysis"
author: |
  | MACS 40700
  | University of Chicago
date: "April 12, 2017"
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(ggthemes)

options(digits = 3)
set.seed(1234)

base_size <- 18

theme_set(theme_gray(base_size = base_size))
```

## Exploratory data analysis

1. Generate questions about your data.
1. Search for answers by visualising, transforming, and modeling your data.
1. Use what you learn to refine your questions and or generate new questions.
1. Rinse and repeat until you publish a paper.

## Exploratory data analysis

1. What type of variation occurs within my variables?
1. What type of covariation occurs between my variables?
1. Are there outliers in the data?
1. Do I have missingness? Are there patterns to it?
1. How much variation/error exists in my statistical estimates? Is there a pattern to it?

## Differences between EDA and modeling

Tips dataset

Variable | Explanation
---------|-------------
`obs` | Observation number
`totbill` | Total bill (cost of the meal), including tax, in US dollars
`tip` | Tip (gratuity) in US dollars
`sex` | Sex of person paying for the meal (0=male, 1=female)
`smoker` | Smoker in party? (0=No, 1=Yes)
`day` | 3=Thur, 4=Fri, 5=Sat, 6=Sun
`time` | 0=Day, 1=Night
`size` | Size of the party

```{r get-tips, include = FALSE}
tips <- read_csv("data/tips.csv")
str(tips)
```

## Tips regression

```{r tips-lm}
tips <- tips %>%
  mutate(tiprate = tip / totbill)

tips_lm <- lm(tiprate ~ sex + smoker + day + time + size, data = tips)
tidy(tips_lm)
```

```{r tips-rsq}
glance(tips_lm)
```

## Exploring `tips`

```{r tips-hist}
ggplot(tips, aes(tip)) +
  geom_histogram(binwidth = 1)
```

## Exploring `tips`

```{r tips-round}
ggplot(tips, aes(tip)) +
  geom_histogram(binwidth = .1)
```

## Exploring `tips`

```{r tips-scatter}
ggplot(tips, aes(totbill, tip)) +
  geom_point() +
  geom_abline(slope = .18, linetype = 2) +
  expand_limits(x = c(0, NA),
                y = c(0, NA)) +
  geom_text(data = tips %>%
              summarize(rsq = format(cor(totbill, tip), digits = 3)),
            aes(x = 2, y = 9, label = rsq))
```

## Exploring `tips`

```{r tips-scatter-many}
ggplot(tips, aes(totbill, tip)) +
  geom_point() +
  geom_abline(slope = .18, linetype = 2) +
  expand_limits(x = c(0, NA),
                y = c(0, NA)) +
  facet_grid(smoker ~ sex, labeller = "label_both") +
  geom_text(data = tips %>%
              group_by(smoker, sex) %>%
              summarize(rsq = format(cor(totbill, tip), digits = 3)),
            aes(x = 2, y = 9, label = rsq))
```

## EDA vs. CDA

* Exploratory data analysis
* Confirmatory data analysis

## EDA in R

![Source: [R for Data Science](http://r4ds.had.co.nz/introduction.html)](http://r4ds.had.co.nz/diagrams/data-science.png)

## Tidy data

![Figure 12.1 from [*R for Data Science*](r4ds.had.co.nz)](http://r4ds.had.co.nz/images/tidy-1.png)

1. Each variable must have its own column.
2. Each observation must have its own row.
3. Each value must have its own cell.

## Functions for tidy data {.scrollable}

```{r example-tidy-graphs, echo = TRUE}
ggplot(gapminder::gapminder, aes(lifeExp)) +
  geom_histogram()

ggplot(diamonds, aes(carat)) +
  geom_histogram()

ggplot(mpg, aes(hwy)) +
  geom_histogram()

ggplot(rcfss::scorecard, aes(admrate)) +
  geom_histogram()
```

## Keeping a record of exploration

* Importance of keeping a record
* Programmatic approach
* Where to store records
    * GitHub
    * R Markdown documents
    * Jupyter notebooks
