---
title: "Day 8 - Multivariate data visualization"
output:
  html_document:
    highlight: pygments
    theme: readable
    toc: yes
    toc_float: yes
    code_folding: hide
  github_document:
    toc: yes
    toc_depth: 3
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

# Objectives

* Demonstrate `lubridate` for working with dates in R and `ggplot2`
* Define correlation
* Review how to implement smoothing lines
* Generate and interpret scatterplot matricies and correlation heatmaps
* Introduce and generate parallel coordinate plots
* Identify methods for implementing three-dimensional graphs in R

```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(ggthemes)
library(knitr)
library(broom)
library(stringr)

options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())
```

# Types of graphs to cover/demo

* Working with dates in R using `lubridate`
    * Demo drawing calendar heatmap in `ggplot2`
    * For more in-depth time series analysis, read [this](https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/)
* Correlation coefficient
    * Coefficient of correlation ($R$)
    * Coefficient of determination ($R^2$)
* Smoothing lines
    * Brief intro to the different `stat_smooth()` methods
* Scatterplot matricies
    * Heatmaps of correlation coefficients
* Parallel coordinate plots
* Contour plots
* 3D plots using plotly
    * Surface plots
    * Scatterplots
    * Line plots

# Session Info {.toc-ignore}

```{r cache = FALSE}
devtools::session_info()
```


