---
title: "Multivariate data visualization"
author: |
  | MACS 40700
  | University of Chicago
date: "April 19, 2017"
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE, include = FALSE}
library(tidyverse)
library(ggthemes)
library(knitr)
library(broom)
library(stringr)
library(modelr)
library(plotly)

options(digits = 3)
set.seed(1234)

base_size <- 22

theme_set(theme_minimal(base_size = base_size))
```

## Working with dates in R

* Numeric (`20174018`)
* String (`2016-04-18`, `April 18th, 2017`, etc.)
* Split into components
* `lubridate`

## Formatting dates

```{r lubridate}
library(lubridate)
```

```{r format-date}
ymd("2017-01-31")
mdy("January 31st, 2017")
dmy("31-Jan-2017")
```

## Extracting date components {.scrollable}

```{r flights}
(flights <- read_csv("data/flights-departed.csv"))
```

```{r flights-over-time}
ggplot(flights, aes(date, value)) +
  geom_line() +
  labs(x = NULL,
       y = "Number of departing commercial flights")
```

## Extracting date components {.scrollable}

```{r flights-year}
(flights <- flights %>%
  mutate(year = year(date),
         yday = yday(date),
         # hack to label the x-axis with months
         days = dmy(format(date,"%d-%m-2016"))))

ggplot(flights, aes(days, value)) +
  geom_line(aes(group = year), alpha = .2) +
  geom_smooth(se = FALSE) +
  scale_x_date(labels = scales::date_format("%b")) +
  labs(x = NULL,
       y = "Number of departing commercial flights")
```

## Extracting date components {.scrollable}

```{r flights-by-month}
(flights <- flights %>%
  mutate(month = month(date, label = TRUE)))

ggplot(flights, aes(month, value)) +
  geom_violin() +
  geom_boxplot(width = .1, outlier.shape = NA) +
  labs(x = NULL,
       y = "Number of departing commercial flights")
```

## Calendar heatmap

* Layer
    * Data - flights
    * Mapping
        * $x$ - weekday (e.g. Sunday, Monday, Tuesday)
        * $y$ - week in month (e.g. first week, second week, third week)
        * Fill - `value` (number of departing flights)
    * Statistical transformation (stat) - `identity`
    * Geometric object (geom) - `geom_tile()`
    * Position adjustment (position) - none
* Scale
    * Fill - low and high-end colors (use shading to identify in-between values)
* Coordinate system - Cartesian coordinate plane
* Faceting - `facet_grid()` (year X month)

## Calendar heatmap

* Year
* Month
* Weekday
* Week-in-month

## Calendar heatmap {.scrollable}

```{r parse-components}
(flights <- flights %>%
  mutate(weekday = wday(date, label = TRUE)))
```

```{r parse-week-in-month}
(flights <- flights %>%
  # generate variables for week in the year (1-54) and the day in the year (1-366)
  mutate(week = week(date),
         yday = yday(date)) %>%
  # normalize to draw calendar correctly - wday should represent the number of days from the Sunday of the week containing January 1st, then adjust based on that
  group_by(year) %>%
  mutate(yday = yday + wday(date)[1] - 2,
         week = floor(yday / 7)) %>%
  group_by(year, month) %>%
  mutate(week_month = week - min(week) + 1))
```

## {.scrollable}

```{r heatmap, fig.asp = 2}
ggplot(flights, aes(weekday, week_month, fill = value)) +
  facet_grid(year ~ month) +
  geom_tile(color = "black") +
  scale_fill_continuous(low = "green", high = "red") +
  scale_x_discrete(labels = NULL) +
  scale_y_reverse(labels = NULL) +
  labs(title = "Domestic commercial flight activity",
       x = NULL,
       y = NULL,
       fill = "Number of departing flights") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.text = element_text(angle = 45))
```

## Smoothing lines

```{r diamonds-point, echo = FALSE}
p <- ggplot(diamonds, aes(carat, price)) +
  geom_point() +
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "Carat size",
       y = "Price")
p
```

----

```{r diamonds-lm}
p +
  geom_smooth(method = "lm", se = FALSE)
```

----

```{r diamonds-gam}
p +
  geom_smooth(se = FALSE)
```

## LOESS

```{r loess, echo = FALSE, warning = FALSE, message = FALSE}
library(lattice)

mod <- loess(NOx ~ E, ethanol, degree = 1, span = .75)
fit <- augment(mod)

mod0 <- loess(NOx ~ E, ethanol, degree = 0, span = .75)
mod1 <- loess(NOx ~ E, ethanol, degree = 1, span = .75)
mod2 <- loess(NOx ~ E, ethanol, degree = 2, span = .75)

fit_all <- ethanol %>%
  gather_predictions(mod0, mod1, mod2) %>%
  mutate(model = factor(model, levels = c("mod0", "mod1", "mod2"),
                        labels = c("Constant", "Linear", "Quadratic")))

ggplot(fit_all, aes(E, NOx)) +
  geom_point() +
  geom_line(aes(y = pred, color = model)) +
  labs(title = "Local linear regression",
       x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J",
       color = "Regression")
```

## LOESS

```{r loess_buildup, dependson="loess", fig.show = "animate", echo = FALSE, warning = FALSE, message = FALSE}
dat <- ethanol %>%
  inflate(center = unique(ethanol$E)) %>%
  mutate(dist = abs(E - center)) %>%
  filter(rank(dist) / n() <= .75) %>%
  mutate(weight = (1 - (dist / max(dist)) ^ 3) ^ 3)

library(gganimate)

p <- ggplot(dat, aes(E, NOx)) +
  geom_point(aes(alpha = weight, frame = center)) +
  geom_smooth(aes(group = center, frame = center, weight = weight), method = "lm", se = FALSE) +
  geom_vline(aes(xintercept = center, frame = center), lty = 2) +
  geom_line(aes(y = .fitted), data = fit, color = "red") +
  labs(x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J")
gg_animate(p)
```

## LOESS

```{r loess_span, dependson="loess", fig.show = "animate", echo = FALSE, warning = FALSE, message = FALSE}
spans <- c(.25, .5, .75, 1)

# create loess fits, one for each span
fits <- data_frame(span = spans) %>%
  group_by(span) %>%
  do(augment(loess(NOx ~ E, ethanol, degree = 1, span = .$span)))

# calculate weights to reproduce this with local weighted fits
dat <- ethanol %>%
  inflate(span = spans, center = unique(ethanol$E)) %>%
  mutate(dist = abs(E - center)) %>%
  filter(rank(dist) / n() <= span) %>%
  mutate(weight = (1 - (dist / max(dist)) ^ 3) ^ 3)

# create faceted plot with changing points, local linear fits, and vertical lines,
# and constant hollow points and loess fit
p <- ggplot(dat, aes(E, NOx)) +
  geom_point(aes(alpha = weight, frame = center)) +
  geom_smooth(aes(group = center, frame = center, weight = weight), method = "lm", se = FALSE) +
  geom_vline(aes(xintercept = center, frame = center), lty = 2) +
  geom_point(shape = 1, data = ethanol, alpha = .25) +
  geom_line(aes(y = .fitted, frame = E, cumulative = TRUE), data = fits, color = "red") +
  facet_wrap(~span) +
  ylim(0, 5) +
  ggtitle("x0 = ") +
  labs(x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J")

gg_animate(p)
```

## Coefficient of correlation ($r$)

* Pearson's $r$
* Scales between $-1$ and $+1$
    * $-1$ -- perfect negative association between the variables
    * $+1$ -- perfect positive association between the variables
    * $0$ -- no relationship between the variables
* Unit-less measure

## Coefficient of correlation ($r$) {.scrollable}

```{r pearson-r}
r_plot <- function(r, n = 100){
  xy <- ecodist::corgen(len = n, r = r) %>%
    bind_cols
  
  ggplot(xy, aes(x, y)) +
    geom_point() +
    ggtitle(str_c("Pearson's r = ", r))
}

r <- c(.8, 0, -.8)

for(r in r){
  print(r_plot(r))
}
```

## Scatterplot matrix

```{r credit-import, include = FALSE}
credit <- read_csv("data/Credit.csv") %>%
  # remove first ID column
  select(-X1)
names(credit) <- stringr::str_to_lower(names(credit))   # convert column names to lowercase
str(credit)
```

```{r credit-scatter-matrix}
pairs(select_if(credit, is.numeric))
```

## Scatterplot matrix {.scrollable}

```{r credit-scatter-ggpairs}
library(GGally)

ggpairs(select_if(credit, is.numeric))
```

## Scatterplot matrix {.scrollable}

```{r credit-scatter-matrix-gender}
ggpairs(credit, mapping = aes(color = gender),
        columns = c("income", "limit", "rating", "cards", "age", "education", "balance"))
```

## Scatterplot matrix {.scrollable}

```{r credit-scatter-smoother}
ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = "smooth"
        )
)
```

## Scatterplot matrix {.scrollable}

```{r wrap}
ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = wrap("smooth", alpha = .1, color = "blue")
        )
)
```

## Scatterplot matrix {.scrollable}

```{r credit-scatter-point-smoother, message = FALSE, warning = FALSE}
scatter_smooth <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    # make data points transparent
    geom_point(alpha = .2) +
    # add default smoother
    geom_smooth(se = FALSE)
}

ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = scatter_smooth
        )
)

ggpairs(credit, mapping = aes(color = gender),
        columns = c("income", "limit", "rating", "cards", "age", "education", "balance"),
        lower = list(
          continuous = scatter_smooth
        )
)
```

## Scatterplot matrix {.scrollable}

```{r diamonds-scatter-matrix}
ggpairs(select(rcfss::scorecard, type:debt))
```

## Heatmap of correlation coefficients {.scrollable}

```{r mpg-cor}
(mpg_lite <- select_if(mpg, is.numeric))

(cormat <- mpg_lite %>%
  cor %>%
  round(2))
```

```{r cormat-tidy}
library(reshape2)
(melted_cormat <- melt(cormat))
```

```{r cormat-tile}
ggplot(melted_cormat, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile()
```

```{r cormat-reduce}
# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

upper_tri <- get_upper_tri(cormat)
upper_tri
```

```{r cormat-tile-tidy}
melted_cormat <- melt(upper_tri, na.rm = TRUE)

ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white") +
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1)) +
 coord_fixed()
```

```{r reorder-cormat}
reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)

# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)

# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

# Print the heatmap
print(ggheatmap)
```

```{r add-coef-heatmap}
ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "bottom")
```

```{r coef-heatmap-function}
cormat_heatmap <- function(data){
  # generate correlation matrix
  cormat <- round(cor(data), 2)
  
  # melt into a tidy table
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
  upper_tri <- get_upper_tri(cormat)
  
  # reorder matrix based on coefficient value
  reorder_cormat <- function(cormat){
    # Use correlation between variables as distance
    dd <- as.dist((1-cormat)/2)
    hc <- hclust(dd)
    cormat <-cormat[hc$order, hc$order]
  }
  
  cormat <- reorder_cormat(cormat)
  upper_tri <- get_upper_tri(cormat)
  
  # Melt the correlation matrix
  melted_cormat <- melt(upper_tri, na.rm = TRUE)
  
  # Create a ggheatmap
  ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
    geom_tile(color = "white")+
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                         midpoint = 0, limit = c(-1,1), space = "Lab", 
                         name="Pearson\nCorrelation") +
    theme_minimal()+ # minimal theme
    theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                     size = 12, hjust = 1))+
    coord_fixed()
  
  # add correlation values to graph
  ggheatmap + 
    geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      panel.grid.major = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      axis.ticks = element_blank(),
      legend.position = "bottom")
}

cormat_heatmap(select_if(mpg, is.numeric))
cormat_heatmap(select_if(credit, is.numeric))
cormat_heatmap(select_if(diamonds, is.numeric))
```

## Parallel coordinate plots {.scrollable}

```{r pcp}
ggparcoord(data = iris, columns = 1:4, groupColumn = 5)

# with the iris data, order the axes by overall class (Species) separation
# using the anyClass option
ggparcoord(data = iris, columns = 1:4, groupColumn = 5, order = "anyClass")

# add points to the plot, add a title, and use an alpha scalar to make the
# lines transparent
p <- ggparcoord(data = iris, columns = 1:4, groupColumn = 5, order = "anyClass", 
    showPoints = TRUE, title = "Parallel Coordinate Plot for the Iris Data", 
    alphaLines = 0.3)
p

# add some basic interactivity
ggplotly(p)
```

## Three dimensions

```{r add-dimensions, echo = FALSE}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point()
```

## Three dimensions {.scrollable}

```{r vote96-model, echo = FALSE}
# import data
(vote <- rcfss::mental_health)

# estimate model
vote_glm <- glm(vote96 ~ age + educ, data = vote, family = "binomial")
tidy(vote_glm)

# extract predicted probabilities
vote_prob <- vote %>%
  data_grid(age = 18:89, educ = 0:20) %>%
  modelr::add_predictions(vote_glm) %>%
  # convert predicted values to probabilities
  mutate(prob = rcfss::logit2prob(pred))

ggplot(vote_prob, aes(age, educ, fill = prob)) +
  geom_tile() +
  scale_fill_gradient2(midpoint = .5, label = scales::percent) +
  labs(title = "Probability of voter turnout in 1996",
       x = "Age",
       y = "Education (in years)",
       fill = "Probability\nof voting")

# cleaner image using geom_raster and interpolate = TRUE
ggplot(vote_prob, aes(age, educ, fill = prob)) +
  geom_raster(interpolate = TRUE) +
  scale_fill_gradient2(midpoint = .5, label = scales::percent) +
  labs(title = "Probability of voter turnout in 1996",
       x = "Age",
       y = "Education (in years)",
       fill = "Probability\nof voting")
```

## 3D scatter plot with mesh

```{r 3d-vote}
plot_ly(vote_prob, x = ~age, y = ~educ, z = ~prob) %>%
  add_mesh()
```

## 3D scatter plot with mesh

```{r 3d-credit}
plot_ly(credit, x = ~limit, y = ~balance, z = ~income) %>%
  add_mesh()
```

## 3D surface plot {.scrollable}

```{r 3d-surface}
plot_ly(z = ~volcano) %>% add_surface()

volcano %>%
  melt %>%
  ggplot(aes(Var1, Var2, z = value)) +
  geom_contour(aes(color = ..level..))
```

