---
title: "Day 8 - Multivariate data visualization"
output:
  html_document:
    highlight: pygments
    theme: readable
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

# Objectives

* Demonstrate `lubridate` for working with dates in R and `ggplot2`
* Define correlation
* Review how to implement smoothing lines
* Generate and interpret scatterplot matricies and correlation heatmaps
* Introduce and generate parallel coordinate plots
* Identify methods for implementing three-dimensional graphs in R

```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(ggthemes)
library(knitr)
library(broom)
library(stringr)
library(modelr)

options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())
```

# Working with dates in R

> For more details on the `lubridate` package, check out [*R for Data Science*](http://r4ds.had.co.nz/dates-and-times.html).

When importing data, date variables can be somewhat tricky to correctly store and utilize. In a spreadsheet, tabular format, dates by default will appear either as numeric (`20174018`) or string (`2016-04-18`, `April 18th, 2017`, etc.) columns. If you want to perform tasks such as: extracting and summarizing over individual components (year, month, day, etc.), we need to represent dates in a different, yet standardized, format.

`lubridate` is a tidyverse package that facilitates working with dates (and date-times) in R.

```{r lubridate}
library(lubridate)
```

## Formatting dates

When using `readr` to import data files, R will use `parse_date()` or `parse_datetime()` to try and format any columns it thinks contain dates or date-times. To manually format dates from strings, use the appropriate function combining `y`, `m`, and `d` in the proper order depending on the original format of the date:

```{r format-date}
ymd("2017-01-31")
mdy("January 31st, 2017")
dmy("31-Jan-2017")
```

## Extracting date components

Let's practice extracting components of dates using an example dataset. `flights-departed.csv` is a time series data file containing the daily number of departing commercial flights in the United States from 1988-2008.

```{r flights}
(flights <- read_csv("data/flights-departed.csv"))
```

We will use `ggplot2` to generate several graphs based on the data. The first is a simple line plot over time of the daily commercial flights. To build this, we don't need to modify `flights`:

```{r flights-over-time}
ggplot(flights, aes(date, value)) +
  geom_line() +
  labs(x = NULL,
       y = "Number of departing commercial flights")
```

But this is quite noisy. Instead, let's draw a line plot depicting commercial flights over a one-year period, with separate lines for each year in the data (1988, 1989, 1990, etc.). To do that, we need to create a new variable `year` which will serve as our grouping variable in `ggplot()`:

```{r flights-year}
(flights <- flights %>%
  mutate(year = year(date),
         yday = yday(date),
         # hack to label the x-axis with months
         days = dmy(format(date,"%d-%m-2016"))))

ggplot(flights, aes(days, value)) +
  geom_line(aes(group = year), alpha = .2) +
  geom_smooth(se = FALSE) +
  scale_x_date(labels = scales::date_format("%b")) +
  labs(x = NULL,
       y = "Number of departing commercial flights")
```

Or we could summarize the distribution of departing commercial flights by days in each month over the 20 year time period:

```{r flights-by-month}
(flights <- flights %>%
  mutate(month = month(date, label = TRUE)))

ggplot(flights, aes(month, value)) +
  geom_violin() +
  geom_boxplot(width = .1, outlier.shape = NA) +
  labs(x = NULL,
       y = "Number of departing commercial flights")
```

Hmmm, there seems to be an outlier in September. What's up with that?

Finally, we can generate a heatmap depicting the change over time of this data by creating a calendar-like visualization.^[A la [U.S. Commercial Flights, 1995-2008](http://mbostock.github.io/d3/talk/20111018/calendar.html).] In order do this, we need the following grammar for the graph:

* Layer
    * Data - flights
    * Mapping
        * $x$ - weekday (e.g. Sunday, Monday, Tuesday)
        * $y$ - week in month (e.g. first week, second week, third week)
        * Fill - `value` (number of departing flights)
    * Statistical transformation (stat) - `identity`
    * Geometric object (geom) - `geom_tile()`
    * Position adjustment (position) - none
* Scale
    * Fill - low and high-end colors (use shading to identify in-between values)
* Coordinate system - Cartesian coordinate plane
* Faceting - `facet_grid()` (year X month)

In order to generate this graph then, we need to create several new variables for `flights`:

* Year
* Month
* Weekday
* Week-in-month

We can use `lubridate` to directly generate three of those variables (we've already generated `year` and `month`):

```{r parse-components}
(flights <- flights %>%
  mutate(weekday = wday(date, label = TRUE)))
```

We use `label = TRUE` to generate factor labels for these values (January, February, March) instead of the numeric equivalent (1, 2, 3).

To generate the final week-in-month variable, we need to combine a few `lubridate` functions to get exactly what we want:

```{r parse-week-in-month}
(flights <- flights %>%
  # generate variables for week in the year (1-54) and the day in the year (1-366)
  mutate(week = week(date),
         yday = yday(date)) %>%
  # normalize to draw calendar correctly - wday should represent the number of days from the Sunday of the week containing January 1st, then adjust based on that
  group_by(year) %>%
  mutate(yday = yday + wday(date)[1] - 2,
         week = floor(yday / 7)) %>%
  group_by(year, month) %>%
  mutate(week_month = week - min(week) + 1))
```

Now that we have the data correctly formatted and all the components are extracted, we can draw the graph:

```{r heatmap, fig.asp = 2}
ggplot(flights, aes(weekday, week_month, fill = value)) +
  facet_grid(year ~ month) +
  geom_tile(color = "black") +
  scale_fill_continuous(low = "green", high = "red") +
  scale_x_discrete(labels = NULL) +
  scale_y_reverse(labels = NULL) +
  labs(title = "Domestic commercial flight activity",
       x = NULL,
       y = NULL,
       fill = "Number of departing flights") +
  theme_void() +
  theme(legend.position = "bottom",
        legend.text = element_text(angle = 45))
```

Aha, now the outlier makes sense. In the days following the September 11th attacks, the United States grounded virtually all commercial air traffic.

# Smoothing lines

When examining multivariate continuous data, scatterplots are a quick and easy visualization to assess relationships. However if the data points become too densely clustered, interpreting the graph becomes difficult. Consider the `diamonds` dataset:

```{r diamonds-point}
p <- ggplot(diamonds, aes(carat, price)) +
  geom_point() +
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "Carat size",
       y = "Price")
p
```

What is the relationship between carat size and price? It appears positive, but there are also a lot of densely packed data points in the middle of the graph. **Smoothing lines** are a method for summarizing the relationship between variables to capture important patterns by approximating the functional form of the relationship. The functional form can take on many shapes. For instance, a very common functional form is a **best-fit line**, also known as **ordinary least squares (OLS)** or **simple linear regression**. We can estimate the model directly using `lm()`, or we can directly plot the line by using `geom_smooth(method = "lm")`:

```{r diamonds-lm}
p +
  geom_smooth(method = "lm", se = FALSE)
```

The downside to a linear best-fit line is that it assumes the relationship between the variables is **additive** and **monotonic**. Therefore the summarized relationship between carat size and price seems wildly incorrect for diamonds with a carat size larger than 3. Instead we could use a [**generalized additive model**](http://cfss.uchicago.edu/persp007_nonlinear.html#generalized_additive_models) which allow for flexible, non-linear relationships between the variables while still implementing a basic regression approach:^[`geom_smooth()` automatically implements the `gam` method for datasets with greater than 1000 observations.]

```{r diamonds-gam}
p +
  geom_smooth(se = FALSE)
```

**Locally weighted scatterplot smoothing** (local regression, LOWESS, or LOESS) fits a separate non-linear function at each target point $x_0$ using only the nearby training observations. This method estimates a regression line based on localized subsets of the data, building up the global function $f$ point-by-point.

Here is an example of a local linear regression on the `ethanol` dataset in the `lattice` package:

```{r loess, echo = FALSE, warning = FALSE, message = FALSE}
library(lattice)

mod <- loess(NOx ~ E, ethanol, degree = 1, span = .75)
fit <- augment(mod)

mod0 <- loess(NOx ~ E, ethanol, degree = 0, span = .75)
mod1 <- loess(NOx ~ E, ethanol, degree = 1, span = .75)
mod2 <- loess(NOx ~ E, ethanol, degree = 2, span = .75)

fit_all <- ethanol %>%
  gather_predictions(mod0, mod1, mod2) %>%
  mutate(model = factor(model, levels = c("mod0", "mod1", "mod2"),
                        labels = c("Constant", "Linear", "Quadratic")))

ggplot(fit_all, aes(E, NOx)) +
  geom_point() +
  geom_line(aes(y = pred, color = model)) +
  labs(title = "Local linear regression",
       x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J",
       color = "Regression")
```

The LOESS is built up point-by-point:

```{r loess_buildup, dependson="loess", fig.show = "animate", echo = FALSE, warning = FALSE, message = FALSE}
dat <- ethanol %>%
  inflate(center = unique(ethanol$E)) %>%
  mutate(dist = abs(E - center)) %>%
  filter(rank(dist) / n() <= .75) %>%
  mutate(weight = (1 - (dist / max(dist)) ^ 3) ^ 3)

library(gganimate)

p <- ggplot(dat, aes(E, NOx)) +
  geom_point(aes(alpha = weight, frame = center)) +
  geom_smooth(aes(group = center, frame = center, weight = weight), method = "lm", se = FALSE) +
  geom_vline(aes(xintercept = center, frame = center), lty = 2) +
  geom_line(aes(y = .fitted), data = fit, color = "red") +
  labs(x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J")
gg_animate(p)
```

One important argument you can control with LOESS is the **span**, or how smooth the LOESS function will become. A larger span will result in a smoother curve, but may not be as accurate. A smaller span will be more local and wiggly, but improve our fit to the training data.

```{r loess_span, dependson="loess", fig.show = "animate", echo = FALSE, warning = FALSE, message = FALSE}
spans <- c(.25, .5, .75, 1)

# create loess fits, one for each span
fits <- data_frame(span = spans) %>%
  group_by(span) %>%
  do(augment(loess(NOx ~ E, ethanol, degree = 1, span = .$span)))

# calculate weights to reproduce this with local weighted fits
dat <- ethanol %>%
  inflate(span = spans, center = unique(ethanol$E)) %>%
  mutate(dist = abs(E - center)) %>%
  filter(rank(dist) / n() <= span) %>%
  mutate(weight = (1 - (dist / max(dist)) ^ 3) ^ 3)

# create faceted plot with changing points, local linear fits, and vertical lines,
# and constant hollow points and loess fit
p <- ggplot(dat, aes(E, NOx)) +
  geom_point(aes(alpha = weight, frame = center)) +
  geom_smooth(aes(group = center, frame = center, weight = weight), method = "lm", se = FALSE) +
  geom_vline(aes(xintercept = center, frame = center), lty = 2) +
  geom_point(shape = 1, data = ethanol, alpha = .25) +
  geom_line(aes(y = .fitted, frame = E, cumulative = TRUE), data = fits, color = "red") +
  facet_wrap(~span) +
  ylim(0, 5) +
  ggtitle("x0 = ") +
  labs(x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J")

gg_animate(p)
```

LOESS lines are best used for datasets with fewer than 1000 observations, otherwise the time and memory usage required to compute the line increases exponentially.

# Coefficient of correlation ($r$)

* Produces a measure of association, known as Pearson's $r$, that gauges the direction and strength of a relationship between two continuous variables
* Scales between $-1$ and $+1$
  * $-1$ -- perfect negative association between the variables
  * $+1$ -- perfect positive association between the variables
  * $0$ -- no relationship between the variables
* Unit-less measure - no matter what scale the variables fall on (e.g. turnout, education, income), the number will always fall between $-1$ and $+1$

```{r pearson-r}
r_plot <- function(r, n = 100){
  xy <- ecodist::corgen(len = n, r = r) %>%
    bind_cols
  
  ggplot(xy, aes(x, y)) +
    geom_point() +
    ggtitle(str_c("Pearson's r = ", r))
}

r <- c(.8, 0, -.8)

for(r in r){
  print(r_plot(r))
}
```

# Scatterplot matricies

To quickly visualize several variables in a dataset and their relation to one another, a **scatterplot matrix** is a quick and detailed tool for generating a series of scatterplots for each combination of variables. Consider `credit.csv` which contains a sample of individuals from a credit card company, identifying their total amount of credit card debt and other financial/demographic variables:

```{r credit-import}
credit <- read_csv("data/Credit.csv") %>%
  # remove first ID column
  select(-X1)
names(credit) <- stringr::str_to_lower(names(credit))   # convert column names to lowercase
str(credit)
```

If we want to quickly assess the relationship between all of the variables (in preparation for more advanced statistical learning techniques), we could generate a matrix of scatterplots using the base `graphics` package:

```{r credit-scatter-matrix}
pairs(select_if(credit, is.numeric))
```

* This only works well if we use strictly quantitative variables (hence the use of `select_if()`)
* We don't automatically see correlation information
* It's not built using `ggplot2` so it's hard to modify using techniques with which we are already familiar

Instead, we can use `GGally::ggpairs()` to generate a scatterplot matrix. `GGally` is a package for R that extends `ggplot2` by adding helper functions for common multivariate data structures. `ggpairs()` is a function that allows us to quickly generate a scatterplot matrix.

```{r credit-scatter-ggpairs}
library(GGally)

ggpairs(select_if(credit, is.numeric))
```

When applied to strictly numeric variables, the lower triangle generates scatterplots, the upper triangle prints the correlation coefficient, and the diagonal panels are density plots of the variable.

Because `ggpairs()` is ultimately based on `ggplot()`, we can use the same types of commands to modify the graph. For instance, if we want to use the color aesthetic to distinguish between men and women in the dataset:

```{r credit-scatter-matrix-gender}
ggpairs(credit, mapping = aes(color = gender),
        columns = c("income", "limit", "rating", "cards", "age", "education", "balance"))
```

Or if we wanted to draw a smoothing line instead of scatterplots, we can modify the graph's [matrix sections](http://ggobi.github.io/ggally/index.html#matrix_sections):

```{r credit-scatter-smoother}
ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = "smooth"
        )
)
```

Hmm, too difficult to see the smoothers because the points are so dense. We can use `wrap()` to pass through individual parameters to the underlying `geom_()`:

```{r wrap}
ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = wrap("smooth", alpha = .1, color = "blue")
        )
)
```

Or we can write a custom function and apply it to the lower triangle panels:

```{r credit-scatter-point-smoother, message = FALSE, warning = FALSE}
scatter_smooth <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    # make data points transparent
    geom_point(alpha = .2) +
    # add default smoother
    geom_smooth(se = FALSE)
}

ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = scatter_smooth
        )
)

ggpairs(credit, mapping = aes(color = gender),
        columns = c("income", "limit", "rating", "cards", "age", "education", "balance"),
        lower = list(
          continuous = scatter_smooth
        )
)
```

`ggpairs()` also works on datasets with a mix of qualitative and quantitative variables, drawing appropriate graphs based on whether the variables are continuous or discrete:

```{r diamonds-scatter-matrix}
ggpairs(select(rcfss::scorecard, type:debt))
```

## Heatmap of correlation coefficients

Scatterplot matricies can provide lots of information, but can also be very densely packed. Perhaps instead we want to quickly visualize the correlation between each of the variables.^[Example drawn from [ ggplot2 : Quick correlation matrix heatmap - R software and data visualization](http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization).] We can easily calculate the correlation coefficients using `cor()`:

```{r mpg-cor}
(mpg_lite <- select_if(mpg, is.numeric))

(cormat <- mpg_lite %>%
  cor %>%
  round(2))
```

But who likes yucky tables. Instead let's turn this into a heatmap. First we need to reshape the data into a tidy structure:

1. Each row contains a single observation
1. Each column contains a single variable
1. Each cell contains a single value

What we need is a data frame with three columns:

1. First variable name
1. Second variable name
1. Correlation coefficient

We can use `reshape2::melt()` to quickly accomplish this:

```{r cormat-tidy}
library(reshape2)
(melted_cormat <- melt(cormat))
```

We can then use `geom_tile()` to visualize the correlation matrix:

```{r cormat-tile}
ggplot(melted_cormat, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile()
```

Not exactly pretty. We can clean it up first by reducing redundancy (remember the upper and lower triangles provide duplicate information):

```{r cormat-reduce}
# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

upper_tri <- get_upper_tri(cormat)
upper_tri
```

Now melt `upper_tri` and repeat the same process, cleaning up the colors for the heatmap as well to distinguish between positive and negative coefficients:

```{r cormat-tile-tidy}
melted_cormat <- melt(upper_tri, na.rm = TRUE)

ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white") +
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1)) +
 coord_fixed()
```

We can also reorder the correlation matrix according to correlation coefficient to help reveal additional trends:

```{r reorder-cormat}
reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)

# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)

# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

# Print the heatmap
print(ggheatmap)
```

Finally we can directly label the correlation coefficient values on the graph, so we have both the color channel and exact values:

```{r add-coef-heatmap}
ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "bottom")
```

To make it more flexible, we can also turn all of this into a function that works for any dataset:

```{r coef-heatmap-function}
cormat_heatmap <- function(data){
  # generate correlation matrix
  cormat <- round(cor(data), 2)
  
  # melt into a tidy table
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
  upper_tri <- get_upper_tri(cormat)
  
  # reorder matrix based on coefficient value
  reorder_cormat <- function(cormat){
    # Use correlation between variables as distance
    dd <- as.dist((1-cormat)/2)
    hc <- hclust(dd)
    cormat <-cormat[hc$order, hc$order]
  }
  
  cormat <- reorder_cormat(cormat)
  upper_tri <- get_upper_tri(cormat)
  
  # Melt the correlation matrix
  melted_cormat <- melt(upper_tri, na.rm = TRUE)
  
  # Create a ggheatmap
  ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
    geom_tile(color = "white")+
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                         midpoint = 0, limit = c(-1,1), space = "Lab", 
                         name="Pearson\nCorrelation") +
    theme_minimal()+ # minimal theme
    theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                     size = 12, hjust = 1))+
    coord_fixed()
  
  # add correlation values to graph
  ggheatmap + 
    geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      panel.grid.major = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      axis.ticks = element_blank(),
      legend.position = "bottom")
}

cormat_heatmap(select_if(mpg, is.numeric))
cormat_heatmap(select_if(credit, is.numeric))
cormat_heatmap(select_if(diamonds, is.numeric))
```

# Types of graphs to cover/demo

* Smoothing lines
    * Brief intro to the different `stat_smooth()` methods
* Scatterplot matricies
    * Heatmaps of correlation coefficients
* Parallel coordinate plots
* Contour plots
* 3D plots using plotly
    * Surface plots
    * Scatterplots
    * Line plots

# Session Info {.toc-ignore}

```{r cache = FALSE}
devtools::session_info()
```


