---
title: "`flexdashboard`"
author: |
  | MACS 40700
  | University of Chicago
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE, include = FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(forcats)

options(digits = 3)
set.seed(1234)

base_size <- 22
theme_set(theme_minimal(base_size = base_size))
```

## `flexdashboard`

* Use R Markdown to publish a group of related data visualizations as an information dashboard
* Supports wide range of components, both static and interactive
* Flexible layouts and auto-resizing (similar to Shiny)
* Can be both static and interactive
* Allows for storyboard layouts

## Major layouts

> [See examples here](https://rmarkdown.rstudio.com/flexdashboard/layouts.html)

* Columns/rows
* Scrolling layouts
* Tabsets
* Multiple pages
* Storyboards

## Components

> [See examples here](https://rmarkdown.rstudio.com/flexdashboard/using.html#components)

* `htmlwidgets`
* R graphs
* Tabular data
* Value boxes
* Gauges
* Navigation bar
* Adding text annotations

## Appearance

* Defined by themes
* Large selection of core themes
* Can define custom themes using CSS

## Examples of static dashboards

* [`ggplotly`: `ggplot2` geoms](https://beta.rstudioconnect.com/jjallaire/htmlwidgets-ggplotly-geoms/)
* [Storyboard: `htmlwidgets` showcase](https://beta.rstudioconnect.com/jjallaire/htmlwidgets-showcase-storyboard/)

## Shiny `flexdashboards`

* `runtime: shiny`
* Allows users to change underlying parameters and see the results immediately
* Need to be run on a Shiny server

## Difference from `Shiny`

* Uses an R Markdown interface to design the application
* Far more control over the layout
* Content auto-resizing is different

## Comparison of apps

* [Dating age rule in pure Shiny](https://bensoltoff.shinyapps.io/cfss-age-rule/)
* [Dating age rule in `flexdashboard` + Shiny](https://bensoltoff.shinyapps.io/age-rule-flexdashboard/)
* [Source code](https://github.com/bensoltoff/age-rule)

## Difference from `shinydashboard`

* [`shinydashboard`](https://rstudio.github.io/shinydashboard/index.html)
* Requires Shiny server
* Allows easier creation of dashboard layouts
* Doesn't allow "static interactivity"
* Uses more familiar coding scheme for Shiny

## Examples of Shiny `flexdashboard`

* [Shiny: Neighborhood diversity](https://walkerke.shinyapps.io/neighborhood_diversity/)
* CFSS grade report
