---
title: "`flexdashboard`"
output:
  html_document:
    highlight: pygments
    theme: readable
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE)
```


```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(forcats)

options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())
```

# `flexdashboard`

* Use R Markdown to publish a group of related data visualizations as an **information dashboard**
* Supports wide range of components, both static and interactive
* Flexible layouts and auto-resizing (similar to Shiny)
* Can be both static and interactive
* Allows for storyboard layouts

## Major layouts

> [See examples here.](https://rmarkdown.rstudio.com/flexdashboard/layouts.html)

* Columns/rows
* Scrolling layouts
* Tabsets
* Multiple pages
* Storyboards

* Auto resizing for small/mobile screens

## Components

> [See examples here.](https://rmarkdown.rstudio.com/flexdashboard/using.html#components)

* `htmlwidgets`
* R graphs
* Tabular data
* Value boxes
* Gauges
* Navigation bar
* Adding text annotations

## Appearance

* Defined by themes
* Large selection of core themes
* Can define custom themes using CSS

## Examples of static dashboards

* [`ggplotly`: `ggplot2` geoms](https://beta.rstudioconnect.com/jjallaire/htmlwidgets-ggplotly-geoms/)
* [Storyboard: `htmlwidgets` showcase](https://beta.rstudioconnect.com/jjallaire/htmlwidgets-showcase-storyboard/)

# Shiny `flexdashboards`

* `runtime: shiny`
* Allows users to change underlying parameters and see the results immediately
* Need to be run on a Shiny server (no longer just a standalone HTML file)

## Difference from `Shiny`

* Uses an R Markdown interface to design the application (no longer functions nested in functions nested in functions)
* Far more control over the layout

## Comparison of apps

* [Dating age rule in pure Shiny](https://bensoltoff.shinyapps.io/cfss-age-rule/)
* [Dating age rule in `flexdashboard` + Shiny](https://bensoltoff.shinyapps.io/age-rule-flexdashboard/)
* [Source code](https://github.com/bensoltoff/age-rule)

# Difference from `shinydashboard`

* [`shinydashboard`](https://rstudio.github.io/shinydashboard/index.html)
* Requires Shiny server
* Allows easier creation of dashboard layouts
* Doesn't allow "static interactivity" a la `flexdashboard`
* Uses more familiar coding scheme for Shiny
* `shinydashboard` came before `flexdashboard`
* Both are under active development, though `flexdashboard` receives more attention

# Examples of Shiny `flexdashboard`

* [Shiny: Neighborhood diversity](https://walkerke.shinyapps.io/neighborhood_diversity/)
* CFSS grade report

# Session Info {.toc-ignore}

```{r cache = FALSE}
devtools::session_info()
```
