---
title: "Geospatial visualization"
author: |
  | MACS 40700
  | University of Chicago
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE, include = FALSE}
library(tidyverse)
library(stringr)
library(forcats)
library(ggmap)

options(digits = 3)
set.seed(1234)

base_size <- 22
theme_set(theme_minimal(base_size = base_size))
```

## Geospatial visualization

* History
* [Google Maps](https://www.google.com/maps)
* Major components
    * Scale
    * Projection
    * Symbols
    
## Map boundaries

* Layer process
* Fill in with:
    * Points
    * Symbols
    * Fills (choropleth)

## Storing map boundaries

* Geographic information system (GIS)
* Specialized software
* GIS in R

## `maps`

* Basic boundary files
* The world
* United States
    * Continental USA
    * States
    * Counties
* France
* Italy
* New Zealand

## `map_data()`

```{r maps}
library(maps)

map_data("state") %>%
  as_tibble()
```

## `map_data()` {.scrollable}

```{r group}
library(gapminder)

# no group aesthetic
ggplot(gapminder, aes(year, lifeExp)) +
  geom_line()

# with grouping by country
ggplot(gapminder, aes(year, lifeExp, group = country)) +
  geom_line()
```

## `map_data()` {.scrollable}

```{r michigan}
map("state", "michigan")
```

## Shapefiles

* `.shp` stores the geographic coordinates of the geographic features
* `.dbf` stores data associated with the geographic features
* `.prj` stores information about the projection of the coordinates in the shapefile

## Shapefiles

```{r readshapepoly}
library(rgdal)

usa <- readOGR("data/census_bureau/cb_2013_us_state_20m/cb_2013_us_state_20m.shp")
str(usa, max.level = 2)
```

## Shapefiles

```{r states-convert}
usa %>%
  fortify() %>%
  head()
```

## Shapefiles

```{r at-data}
usa@data %>%
  as_tibble
```

## `fortify()` {.scrollable}

```{r fortify-id}
# state name
usa %>%
  fortify(region = "NAME") %>%
  head

# FIPS code
usa %>%
  fortify(region = "STATEFP") %>%
  head

# keep it all
(usa2 <- usa %>%
  fortify(region = "NAME") %>%
  as_tibble %>%
  left_join(usa@data, by = c("id" = "NAME")))
```

## Simple features

* Standard of how to represent real-world objects on a computer
* Emphasizes spatial geometry
* Widely implemented in commercial software
* `sf` package

## What is a feature?

* A thing or an object in the real world
* Frequently consist of other objects
* Features have a geometry describing where on Earth the feature is located
* They have attributes, which describe other properties of the feature

## Dimensions

* Geometries composed of points
* Coordinates in 2-, 3- or 4-dimensional space
* X and Y coordinates
* Z coordinate
* M coordinate

## Simple feature geometry types

| type | description                                        |
| ---- | -------------------------------------------------- |
| `POINT` | zero-dimensional geometry containing a single point |
| `LINESTRING` | sequence of points connected by straight, non-self intersecting line pieces; one-dimensional geometry |
| `POLYGON` | geometry with a positive area (two-dimensional); sequence of points form a closed, non-self intersecting ring; the first ring denotes the exterior ring, zero or more subsequent rings denote holes in this exterior ring |
| `MULTIPOINT` | set of points; a MULTIPOINT is simple if no two Points in the MULTIPOINT are equal |
| `MULTILINESTRING` | set of linestrings |
| `MULTIPOLYGON` | set of polygons |
| `GEOMETRYCOLLECTION` | set of geometries of any type except GEOMETRYCOLLECTION |

## Simple features in R

* `sf` stores simple features as basic R data structures (lists, matrix, vectors, etc.)
* Data frame storage
* One row per feature
* Geometries and list columns
* The three classes used to represent simple features
    * `sf`
    * `sfc`
    * `sfg`

## Example: North Carolina

```{r nc-import}
library(sf)
nc <- st_read(system.file("shape/nc.shp", package="sf"))
```

## Example: North Carolina

```{r nc-class}
class(nc)
```

## Example: North Carolina

```{r nc-column}
attr(nc, "sf_column")
```

## Example: North Carolina

```{r nc-rows}
print(nc[9:15], n = 3)
```

## Plotting boundaries with `maps` {.scrollable}

```{r maps-plot}
# map of the world
map()

# usa boundaries
map("usa")
map("state")

# county map of illinois
map("county", "illinois")
```

## Plotting boundaries with `ggplot2` {.scrollable}

```{r usa-bound}
usa <- map_data("usa") %>%
  as_tibble
usa
```

## Simple black map {.scrollable}

```{r usa-map}
ggplot() +
  geom_polygon(data = usa, aes(x = long, y = lat, group = group))
```

## Simple black map {.scrollable}

```{r coord-fixed}
ggplot() +
  geom_polygon(data = usa, aes(x = long, y = lat, group = group)) +
  coord_fixed()
```

## Simple black map {.scrollable}

```{r coord-fixed-13}
ggplot() +
  geom_polygon(data = usa, aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3)
```

## Change the colors {.scrollable}

```{r fill-color}
ggplot() +
  geom_polygon(data = usa, aes(x = long, y = lat, group = group),
               fill = NA, color = "red") + 
  coord_fixed(1.3)

gg1 <- ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group),
               fill = "violet", color = "blue") + 
  coord_fixed(1.3)
gg1
```

## Always remember to use the group aesthetic {.scrollable}

```{r no-group}
ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat),
               fill = "violet", color = "blue") + 
  coord_fixed(1.3)
```

## State maps

```{r states}
states <- map_data("state") %>%
  as_tibble()
states
```

## State maps {.scrollable}

```{r states-no-fill}
ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, group = group), color = "white") + 
  coord_fixed(1.3)
```

## State maps {.scrollable}

```{r states-fill}
ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_fixed(1.3) +
  # turn off color legend
  theme(legend.position = "none")
```

## Plot a subset of states {.scrollable}

```{r midwest}
midwest <- subset(states, region %in% c("illinois", "indiana", "iowa",
                                        "kansas", "michigan", "minnesota",
                                        "missouri", "nebraska", "north dakota",
                                        "ohio", "south dakota", "wisconsin"))

ggplot(data = midwest) + 
  geom_polygon(aes(x = long, y = lat, group = group),
               fill = "palegreen", color = "black") + 
  coord_fixed(1.3)
```

## But what about Alaska and Hawaii?

```{r all-50, echo = FALSE}
library(maptools)
library(rgeos)
library(rgdal)

#state shape file from
# http://www.arcgis.com/home/item.html?id=f7f805eb65eb4ab787a0a3e1116ca7e5
loc <- file.path(tempdir(), "stats_dat")
unzip(system.file("extdata", "states_21basic.zip", package = "fiftystater"),
      exdir = loc)
fifty_states_sp <- readOGR(dsn = loc, layer = "states", verbose = FALSE)

fifty_states_sp %>%
  fortify(region = "STATE_NAME") %>%
  mutate(id = tolower(id)) %>%
  ggplot(mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
```

## `fiftystater` {.scrollable}

```{r fiftystater}
library(fiftystater)

data("fifty_states")
fifty_states %>%
  as_tibble

ggplot(data = fifty_states, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
```

## From a shapefile

```{r usa2-str}
glimpse(usa2)
```

## From a shapefile {.scrollable}

```{r usa2-all}
ggplot(data = usa2, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
```

## From a shapefile {.scrollable}

```{r usa2-filter}
usa2 <- usa2 %>%
  filter(id != "Alaska", id != "Hawaii", id != "Puerto Rico")

ggplot(data = usa2, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
```

## An `sf` object

```{r nc-str}
glimpse(nc)
```

## `geom_sf()` {.scrollable}

```{r sf-plot}
ggplot() +
  geom_sf(data = nc)
```

## `ggmap` {.scrollable}

```{r ggmap}
library(ggmap)
```

```{r google-maps}
get_googlemap("Saieh Hall For Economics, East 58th Street, Chicago, IL", zoom = 10) %>%
  ggmap()
```

## `leaflet` {.scrollable}

> [Leaflet for R](https://rstudio.github.io/leaflet/)

## Map projections

* [Cartographers for social equality](https://www.youtube.com/watch?v=vVX-PrBRtTY)

## Changing map projections {.scrollable}

```{r projections}
ggplot(data = usa2, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map() +
  ggtitle("Mercator projection (default)")

ggplot(data = usa2, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  ggtitle("Albers equal-area projection")

ggplot(data = map_data("world"), mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map(projection = "mollweide", xlim = c(-180, 180)) +
  ggtitle("Mollweide projection")
```

## Points

```{r nycflights}
library(nycflights13)
airports
```

## Points {.scrollable}

```{r scatter}
ggplot(airports, aes(lon, lat)) +
  geom_point()
```

## Points {.scrollable}

```{r flights-usa}
ggplot() + 
  coord_map() + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "gray") +
  geom_point(data = airports, aes(x = lon, y = lat), shape = 1)
```

## Points {.scrollable}

```{r crop}
ggplot() + 
  coord_map(xlim = c(-130, -60),
            ylim = c(20, 50)) + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "gray") +
  geom_point(data = airports, aes(x = lon, y = lat), shape = 1)
```

## Points {.scrollable}

```{r crop-proj}
ggplot() + 
  coord_map(projection = "albers", lat0 = 25, lat1 = 50,
            xlim = c(-130, -60),
            ylim = c(20, 50)) + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "gray") +
  geom_point(data = airports, aes(x = lon, y = lat), shape = 1)
```

## Symbols {.scrollable}

```{r airport-alt}
ggplot() + 
  coord_map(xlim = c(-130, -60),
            ylim = c(20, 50)) + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "white") +
  geom_point(data = airports, aes(x = lon, y = lat, size = alt),
             fill = "grey", color = "black", alpha = .2) +
  theme_void() +
  theme(legend.position = "none")
```

## Symbols {.scrollable}

```{r airport-dest}
airports_n <- flights %>%
  count(dest) %>%
  left_join(airports, by = c("dest" = "faa"))

ggplot() + 
  coord_map(xlim = c(-130, -60),
            ylim = c(20, 50)) + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "white") +
  geom_point(data = airports_n, aes(x = lon, y = lat, size = n),
             fill = "grey", color = "black", alpha = .2) +
  theme_void() +
  theme(legend.position = "none")
```

## Drawing choropleth maps {.scrollable}

```{r county-shp}
usa <- readOGR("data/census_bureau/cb_2013_us_state_20m/cb_2013_us_state_20m.shp")

usa2 <- usa %>%
  fortify(region = "GEOID") %>%
  as_tibble %>%
  left_join(usa@data, by = c("id" = "GEOID")) %>%
  # filter out Alaska, Hawaii, Puerto Rico via FIPS codes
  filter(!(STATEFP %in% c("02", "15", "72")))

counties <- readOGR("data/census_bureau/cb_2013_us_county_20m/cb_2013_us_county_20m.shp")

counties2 <- counties %>%
  fortify(region = "GEOID") %>%
  as_tibble %>%
  left_join(counties@data, by = c("id" = "GEOID")) %>%
  # filter out Alaska, Hawaii, Puerto Rico via FIPS codes
  filter(!(STATEFP %in% c("02", "15", "72")))

ggplot(counties2, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map()
```

## Drawing choropleth maps {.scrollable}

```{r import-foreign}
(fb_state <- read_csv("data/census_bureau/ACS_13_5YR_B05012_state/ACS_13_5YR_B05012.csv") %>%
  mutate(rate = HD01_VD03 / HD01_VD01))

(fb_county <- read_csv("data/census_bureau/ACS_13_5YR_B05012_county/ACS_13_5YR_B05012.csv") %>%
  mutate(rate = HD01_VD03 / HD01_VD01))
```

## Joining the data to regions {.scrollable}

```{r geom-map-state}
ggplot(fb_state, aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate), map = usa2) +
  expand_limits(x = usa2$long, y = usa2$lat)
```

## Joining the data to regions {.scrollable}

```{r geom-map-state-neat}
ggplot(fb_state, aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate), map = usa2) +
  expand_limits(x = usa2$long, y = usa2$lat) +
  scale_fill_continuous(labels = scales::percent) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50)
```

## Joining the data to regions {.scrollable}

```{r geom-map-county}
ggplot(fb_county, aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate), map = counties2) +
  expand_limits(x = counties2$long, y = counties2$lat) +
  scale_fill_continuous(labels = scales::percent) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50)
```

## `sf` {.scrollable}

```{r nc-fill}
ggplot() +
  geom_sf(data = nc, aes(fill = BIR74))
```

## `sf` {.scrollable}

```{r nc-facet}
(nc2 <- nc %>%
  select(SID74, SID79, geometry) %>%
  gather(VAR, SID, -geometry))

ggplot() +
  geom_sf(data = nc2, aes(fill = SID)) +
  facet_wrap(~ VAR, ncol = 1)
```

## Selecting color palettes

```{r color-wheel, echo = FALSE}
# generate simulated data points
sim_points <- data_frame(x = factor(1:6))

plots <- purrr::map(1:6, ~ ggplot(sim_points[1:.x, ], aes(x, x, color = x)) +
  geom_point(size = 5) +
    ggtitle(paste(.x, "color")) +
  theme(legend.position = "none"))

gridExtra::marrangeGrob(plots, nrow = 2, ncol = 3, top = NULL)
```

## RColorBrewer

> [Color Brewer](http://colorbrewer2.org/)

## Sequential {.scrollable}

```{r cb-seq}
fb_county %>%
  mutate(rate_cut = cut_number(rate, 6)) %>%
  ggplot(aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate_cut), map = counties2) +
  expand_limits(x = counties2$long, y = counties2$lat) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  scale_fill_brewer(palette = "BuGn")

fb_county %>%
  mutate(rate_cut = cut_number(rate, 6)) %>%
  ggplot(aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate_cut), map = counties2) +
  expand_limits(x = counties2$long, y = counties2$lat) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  scale_fill_brewer(palette = "YlGn")

fb_county %>%
  mutate(rate_cut = cut_number(rate, 6)) %>%
  ggplot(aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate_cut), map = counties2) +
  expand_limits(x = counties2$long, y = counties2$lat) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  scale_fill_brewer(palette = "Blues")
```

## Qualitative {.scrollable}

```{r qual}
state_data <- data_frame(name = state.name,
                         region = state.region,
                         subregion = state.division,
                         abb = state.abb) %>%
  bind_cols(as_tibble(state.x77)) %>%
  # get id variable into data frame
  left_join(usa2 %>%
              select(id, NAME) %>%
              distinct,
            by = c("name" = "NAME")) %>%
  # remove Alaska and Hawaii
  na.omit

# set region base plot
region_p <- ggplot(state_data, aes(map_id = id)) +
  geom_map(aes(fill = region), map = usa2) +
  expand_limits(x = usa2$long, y = usa2$lat) +
  labs(fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50)
region_p

# try different color brewers
region_p +
  scale_fill_brewer(palette = "Paired")

region_p +
  scale_fill_brewer(palette = "Dark2")

region_p +
  scale_fill_brewer(palette = "Pastel2")

# set subregion base plot
subregion_p <- ggplot(state_data, aes(map_id = id)) +
  geom_map(aes(fill = subregion), map = usa2) +
  expand_limits(x = usa2$long, y = usa2$lat) +
  labs(fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50)
subregion_p

subregion_p +
  scale_fill_brewer(palette = "Paired")

subregion_p +
  scale_fill_brewer(palette = "Set1")

subregion_p +
  scale_fill_brewer(palette = "Pastel1")
```
