---
title: "Geospatial visualization"
output:
  html_document:
    highlight: pygments
    theme: readable
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE)
```

```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(modelr)
library(forcats)
library(ggmap)
library(plotly)

options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())
```

# Overview

Geospatial visualizations are some of the oldest data visualization methods in human existence. Data maps were first popularized in the seventeenth century and have grown in complexity and detail since then. Consider [Google Maps](https://www.google.com/maps), the sheer volume of data depicted, and the analytical pathways available to its users.

Geometric visualizations are used to depict spatial features, and with the incorporation of data reveal additional attributes and information. The main features of a map are defined by its **scale** (the proportion between distances and sizes on the map), its **projection** (how the three-dimensional Earth is represented on a two-dimensional surface), and its **symbols** (how data is depicted and visualized on the map).^[See chapter 10 in **The Truthful Art** for a more detailed introduction to these features.]

Drawing maps in R is a layer-like process. Typically you start by drawing the boundaries of the geographic regions you wish to visualize, then you add additional layers defined by other variables:

* Points
* Symbols
* Fills (choropleths)

# Storing geospatial data

A **geographic information system** (GIS) is software that is "designed to capture, store, manipulate, analyze, manage, and present spatial or geographic data".^[Source: [Wikipedia](https://en.wikipedia.org/wiki/Geographic_information_system).] There are many specialized software packages for spatial data analysis, many of which are commercial or proprietary software. For serious spatial data analysis tasks, you probably want to learn how to use these products. However for casual usage, R has a number of tools for drawing maps, most importantly `ggplot2`.

## Built-in packages

The `maps` package includes the `map()` function for drawing maps based on bundled geodatabases using the `graphics` package. `maps` includes basic boundary files for:

* The world
* United States
    * Continental USA
    * States
    * Counties
* France
* Italy
* New Zealand

We can extract the boundary files using `map_data()`, which converts the data object to a data frame:

```{r maps}
library(maps)

# world
map_data("world") %>%
  as_tibble()

# usa
map_data("usa") %>%
  as_tibble()

# states
map_data("state") %>%
  as_tibble()
```

`map_data()` converts the geodatabases into data frames where each row is a single point on the map. The resulting data frame contains the following variables:

* `long` - longitude. Things to the west of the prime meridian are negative
* `lat` - latitude
* `order` - this identifies the order `ggplot()` should follow to "connect the dots" and draw the borders
* `region` and `subregion` identify what region or subregion a set of points surrounds
* `group` - this is perhaps the most important variable in the data frame. `ggplot()` uses the `group` aesthetic to determine whether adjacent points should be connected by a line. If they are in the same group, then the points are connected. If not, then they are not connected. This is the same basic principle for standard `geom_line()` plots:

    ```{r group}
    library(gapminder)
    
    # no group aesthetic
    ggplot(gapminder, aes(year, lifeExp)) +
      geom_line()
    
    # with grouping by country
    ggplot(gapminder, aes(year, lifeExp, group = country)) +
      geom_line()
    ```

    Note that `group` is not the same thing as `region` or `subregion`. If a region contains landmasses that are discontiguous, there should be multiple groups to properly draw the region:
    
    ```{r michigan}
    map("state", "michigan")
    ```

## Shapefiles

`maps` contains a very limited number of geodatabases. If you want to import a different country's borders or some other geographic information, you will likely find the data in a **shapefile**. This is a special file format that encodes points, lines, and polygons in geographic space. Files appear with a `.shp` extension, sometimes with accompanying files ending in `.dbf` and `.prj`.

* `.shp` stores the geographic coordinates of the geographic features (e.g. country, state, county)
* `.dbf` stores data associated with the geographic features (e.g. unemployment rate, crime rates, percentage of votes cast for Donald Trump)
* `.prj` stores information about the projection of the coordinates in the shapefile (we'll handle this shortly)

[There are a crap ton of packages for R that allow you to interact with shapefiles and spatial data.](https://cran.r-project.org/web/views/Spatial.html) Let's start with a shapefile for state boundaries in the United States.^[Originally obtained from the Census Bureau.] We'll use `readOGR()` from the `rgdal` package to read in the data file:

```{r readshapepoly}
library(rgdal)

usa <- readOGR("data/census_bureau/cb_2013_us_state_20m/cb_2013_us_state_20m.shp")
str(usa, max.level = 2)
```

This is decidedly not a tidy data frame. Once you import the shapefile, it's best to convert it to a data frame for `ggplot()`. We can do this using `fortify()`:

```{r states-convert}
usa %>%
  fortify() %>%
  head()
```

Under this approach, the `id` variable is just a number assigned to each region (in this case, each state/territory). However the shapefile contains linked data with attributes for each region. We can access this using the `@data` accessor:

```{r at-data}
usa@data %>%
  as_tibble
```

We can keep these variables in the new data frame through parameters to `fortify(region = "")`:

```{r fortify-id}
# state name
usa %>%
  fortify(region = "NAME") %>%
  head

# FIPS code
usa %>%
  fortify(region = "STATEFP") %>%
  head

# keep it all
(usa2 <- usa %>%
  fortify(region = "NAME") %>%
  as_tibble %>%
  left_join(usa@data, by = c("id" = "NAME")))
```

## Simple features

[Simple features](https://en.wikipedia.org/wiki/Simple_Features) or [_simple feature access_](http://www.opengeospatial.org/standards/sfa) refers to a formal standard (ISO 19125-1:2004) that describes how objects in the real world can be represented in computers, with emphasis on the _spatial_ geometry of these objects. It also describes how such objects can be stored in and retrieved from databases, and which geometrical operations should be defined for them.

The standard is widely implemented in spatial databases (such as PostGIS), commercial GIS (e.g., [ESRI ArcGIS](http://www.esri.com/)) and forms the vector data basis for libraries such as [GDAL](http://www.gdal.org/). A subset of simple features forms the [GeoJSON](http://geojson.org/) standard.

R has well-supported classes for storing spatial data ([sp](https://CRAN.R-project.org/package=sp)) and interfacing to the above mentioned environments ([rgdal](https://CRAN.R-project.org/package=rgdal), [rgeos](https://CRAN.R-project.org/package=rgeos)), but has so far lacked a complete implementation of simple features, making conversions at times convoluted, inefficient or incomplete. The package [sf](http://github.com/r-spatial/sf) tries to fill this gap, and aims at succeeding [sp](https://CRAN.R-project.org/package=sp) in the long term.

### What is a feature?

* A thing or an object in the real world
* Frequently consist of other objects
    * A set of features can form a single feature
    * i.e. a tree can be a feature, but a set of trees can form a forest which is itself a stand-alone feature
* Features have a **geometry** describing where on Earth the feature is located
* They have attributes, which describe other properties of the feature

### Dimensions

All geometries are composed of points. Points are coordinates in a 2-, 3- or 4-dimensional space. All points in a geometry have the same dimensionality. In addition to X and Y coordinates, there are two optional additional dimensions:

* a Z coordinate, denoting altitude
* an M coordinate (rarely used), denoting some _measure_ that is associated with the point, rather than with the feature as a whole (in which case it would be a feature attribute); examples could be time of measurement, or measurement error of the coordinates

The four possible cases then are:

1. two-dimensional points refer to x and y, easting and northing, or longitude and latitude, we refer to them as XY
2. three-dimensional points as XYZ
3. three-dimensional points as XYM
4. four-dimensional points as XYZM (the third axis is Z, fourth M)

### Simple feature geometry types

The following seven simple feature types are the most common, and are for instance the only ones used for [GeoJSON](https://tools.ietf.org/html/rfc7946):

| type | description                                        |
| ---- | -------------------------------------------------- |
| `POINT` | zero-dimensional geometry containing a single point |
| `LINESTRING` | sequence of points connected by straight, non-self intersecting line pieces; one-dimensional geometry |
| `POLYGON` | geometry with a positive area (two-dimensional); sequence of points form a closed, non-self intersecting ring; the first ring denotes the exterior ring, zero or more subsequent rings denote holes in this exterior ring |
| `MULTIPOINT` | set of points; a MULTIPOINT is simple if no two Points in the MULTIPOINT are equal |
| `MULTILINESTRING` | set of linestrings |
| `MULTIPOLYGON` | set of polygons |
| `GEOMETRYCOLLECTION` | set of geometries of any type except GEOMETRYCOLLECTION |

### Coordinate reference system

Coordinates can only be placed on the Earth's surface when their coordinate reference system (CRS) is known; this may be an spheroid CRS such as WGS84, a projected, two-dimensional (Cartesian) CRS such as a UTM zone or Web Mercator, or a CRS in three-dimensions, or including time. Similarly, M-coordinates need an attribute reference system, e.g. a [measurement unit](https://CRAN.R-project.org/package=units).

### Simple features in R

* `sf` stores simple features as basic R data structures (lists, matrix, vectors, etc.)
* Typical data structure stores geometric and feature attributes as a data frame
* One row per feature
* Since feature geometries are not single-valued, they are put in a list-column with each list element holding teh simple feature geometry of that feature
* The three classes used to represent simple features are:
    * `sf`, the table (`data.frame`) with feature attributes and feature geometries, which contains
    * `sfc`, the list-column with the geometries for each feature (record), which is composed of
    * `sfg`, the feature geometry of an individual simple feature

#### Example: North Carolina

`st_read(0` imports a shapefile and converts it to a simple feature data frame):

```{r nc-import}
library(sf)
nc <- st_read(system.file("shape/nc.shp", package="sf"))
```

The short report printed gives the file name, the driver (ESRI Shapefile), mentions that there are 100 features (records, represented as rows) and 14 fields (attributes, represented as columns). This object is of class

```{r nc-class}
class(nc)
```

meaning it extends (and "is" a) `data.frame`, but with a single list-column with geometries, which is held in the column with name

```{r nc-column}
attr(nc, "sf_column")
```

If we print the first three features, we see their attribute values and an abridged version of the geometry

```{r nc-rows}
print(nc[9:15], n = 3)
```

In the output we see:

* Each row is a simple feature: a single record, or `data.frame` row, consisting of attributes and geometry
* The `geometry` column is a simple feature list-column (an object of class `sfc`, which is a column in the `data.frame`)
* Each value in `geometry` is a single simple feature geometry (an object of class `sfg`)

# Plotting methods for boundaries

## `maps`

The `maps` package includes the `map()` function for drawing maps based on bundled geodatabases using the `graphics` package:

```{r maps-plot}
# map of the world
map()

# usa boundaries
map("usa")
map("state")

# county map of illinois
map("county", "illinois")
```

These are fine, but we'd rather use them with our friendly `ggplot2` library. We can do this by converting the geodatabases into data frames for plotting with `ggplot2` using the `map_data()` function we saw before.

## `ggplot2`

### `map_data()`

Let's draw a map of the United States. First we need to store the USA map boundaries in a data frame:

```{r usa-bound}
usa <- map_data("usa") %>%
  as_tibble
usa
```

#### Simple black map

* We can use `geom_polygon()` to draw lines between points and close them up (connect the last point with the first point)
* `x = long` and `y = lat`
* We map the `group` aesthetic to the `group` column

```{r usa-map}
ggplot() +
  geom_polygon(data = usa, aes(x = long, y = lat, group = group))
```

Ta-da! A few things we want to immediately start thinking about. First, because latitude and longitude have absolute relations to one another, we need to fix the aspect ratio so that we don't accidentially compress the graph in one dimension. Fixing the aspect ratio also means that even if you change the outer dimensions of the graph (i.e. adjust the window size), then the aspect ratio of the graph itself remains unchanged. We can do this using the `coord_fixed()` function:

```{r coord-fixed}
ggplot() +
  geom_polygon(data = usa, aes(x = long, y = lat, group = group)) +
  coord_fixed()
```

Now it looks a little squished. You can play around with the aspect ratio to find a better projection:^[Or as we'll see shortly, apply a map projection to the graph.]

```{r coord-fixed-13}
ggplot() +
  geom_polygon(data = usa, aes(x = long, y = lat, group = group)) +
  coord_fixed(1.3)
```

#### Change the colors

Since this is a `ggplot()` object, we can change the fill and color aesthetics for the map:

```{r fill-color}
ggplot() +
  geom_polygon(data = usa, aes(x = long, y = lat, group = group),
               fill = NA, color = "red") + 
  coord_fixed(1.3)

gg1 <- ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group),
               fill = "violet", color = "blue") + 
  coord_fixed(1.3)
gg1
```

#### Always remember to use the group aesthetic

What happens if we plot the map without the group aesthetic?

```{r no-group}
ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat),
               fill = "violet", color = "blue") + 
  coord_fixed(1.3)
```

Oops. The map doesn't connect the dots in the correct order.

### State maps

`maps` also comes with a state boundaries geodatabase:

```{r states}
states <- map_data("state") %>%
  as_tibble()
states
```

By default, each state is filled with the same color:

```{r states-no-fill}
ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, group = group), color = "white") + 
  coord_fixed(1.3)
```

We can adjust this by using the `fill` aesthetic. Here, let's map `region` to fill:

```{r states-fill}
ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_fixed(1.3) +
  # turn off color legend
  theme(legend.position = "none")
```

Here, each state is assigned a different color at random. You can start to imagine how we might build a choropleth by mapping a different variable to `fill`, but we'll return to that in a little bit.

#### Plot a subset of states

We can use `filter()` to subset the `states` data frame and draw a map with only a subset of the states. For example, if we want to only graph states in the Midwest:

```{r midwest}
midwest <- subset(states, region %in% c("illinois", "indiana", "iowa",
                                        "kansas", "michigan", "minnesota",
                                        "missouri", "nebraska", "north dakota",
                                        "ohio", "south dakota", "wisconsin"))

ggplot(data = midwest) + 
  geom_polygon(aes(x = long, y = lat, group = group),
               fill = "palegreen", color = "black") + 
  coord_fixed(1.3)
```

#### Zoom in on Illinois and its counties

First let's get the Illinois boundaries:

```{r il-state}
il_df <- filter(states, region == "illinois")
```

Now let's get the accompanying counties:

```{r il-counties}
counties <- map_data("county") %>%
  as_tibble
il_county <- filter(counties, region == "illinois")
il_county
```

Plot the state first. This time, lets' remove all the axes gridlines and background junk using `theme_void()`:

```{r il-map}
il_base <- ggplot(data = il_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

il_base +
  theme_void()
```

Now let's plot the county boundaries in white:

```{r il-county-plot}
il_base +
  theme_void() + 
  geom_polygon(data = il_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA)  # get the state border back on top
```

### But what about Alaska and Hawaii?

If you were observant, you noticed `map_data("states")` only includes the 48 contiguous states in the United States. This is because Alaska and Hawaii exist far off from the rest of the states. What happens if you try to draw a map with them in it?

```{r all-50, echo = FALSE}
library(maptools)
library(rgeos)
library(rgdal)

#state shape file from
# http://www.arcgis.com/home/item.html?id=f7f805eb65eb4ab787a0a3e1116ca7e5
loc <- file.path(tempdir(), "stats_dat")
unzip(system.file("extdata", "states_21basic.zip", package = "fiftystater"),
      exdir = loc)
fifty_states_sp <- readOGR(dsn = loc, layer = "states", verbose = FALSE)

fifty_states_sp %>%
  fortify(region = "STATE_NAME") %>%
  mutate(id = tolower(id)) %>%
  ggplot(mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
```

Yup, that doesn't look right. Most maps of the United States place Alaska and Hawaii as **insets** to the south of California. Until recently, in R this was an extremely tedious task that required manually changing the latitude and longitude coordinates for these states to place them in the correct location. Fortunately a new package is available that has already done the work for you. [`fiftystater`](https://cran.r-project.org/web/packages/fiftystater/vignettes/fiftystater.html) includes the `fifty_states` data frame which contains adjusted coordinates for Alaska and Hawaii to plot them with the mainland:

```{r fiftystater}
library(fiftystater)

data("fifty_states")
fifty_states %>%
  as_tibble

ggplot(data = fifty_states, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
```

### From a shapefile

Recall how to import a shapefile and convert it to a data frame. Once we have it in that form:

```{r usa2-str}
glimpse(usa2)
```

Now we can plot it like normal using `ggplot()`:

```{r usa2-all}
ggplot(data = usa2, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
```

Because the data file comes from the Census Bureau, we also get boundaries for Alaska, Hawaii, and Puerto Rico. To remove them from the data, just use `filter()`:

```{r usa2-filter}
usa2 <- usa2 %>%
  filter(id != "Alaska", id != "Hawaii", id != "Puerto Rico")

ggplot(data = usa2, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
```

### An `sf` object

Recall the North Carolina simple features dataset:

```{r nc-str}
glimpse(nc)
```

We have information on each county in the state of North Carolina. To plot using `ggplot2`, make sure you have the development version of `ggplot2` installed using `devtools::install_github("tidyverse/ggplot2")`. We can use `geom_sf()` to plot the shapefile:

```{r sf-plot}
ggplot() +
  geom_sf(data = nc)
```

This produces just the boundaries.

### `ggmap`

Rather than relying on geodatabases or shapefiles which store boundaries as numeric data, we can use `ggmap` to retrieve raster map tiles from online mapping services.

```{r ggmap}
library(ggmap)
```

```{r google-maps}
get_googlemap("Saieh Hall For Economics, East 58th Street, Chicago, IL", zoom = 10) %>%
  ggmap()
```

This generates a `ggplot()` which uses the Google Maps image as a raster image, or background you can draw on top of.

## `leaflet`

[Leaflet](https://leafletjs.com/) is an open-source JavaScript library for drawing interactive maps. `leaflet` is a package for R that allows you to generate Leaflet maps using R functions and data objects. These maps are highly customizable and accept a wide range of data sources, including:

* Data frames
* Shapefiles
* GeoJSON

`leaflet` integrates nicely into the `tidyverse` in that the map objects (or widgets) are built as a series of piped operations. The basic process is:

1. Create a map widget using `leaflet()`
1. Add layers (or features) to the map by using layer functions (e.g. `addTiles`, `addMarkers`, `addPolygons`)
1. Build multiple layers as necessary
1. Print the map widget to display it

```{r leaflet-chicago}
library(leaflet)

m <- leaflet() %>%
  setView(lng = -87.596154, lat = 41.790115, zoom = 14)
m %>%
  addTiles()
```

# Changing map projections

As you saw in **The Truthful Art**, representing portions of the globe on a flat surface can be challenging. Depending on how you project the map, you can distort or emphasize certain features of the map. Fortunately, `ggplot()` includes the `coord_map()` function which allows us to easily implement different projection methods.^[This function replaces `coord_fixed()`.] Depending on the projection method, you may need to pass additional arguments to `coord_map()` to define the standard parallel lines used in projecting the map:

```{r projections}
ggplot(data = usa2, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map() +
  ggtitle("Mercator projection (default)")

ggplot(data = usa2, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  ggtitle("Albers equal-area projection")

ggplot(data = map_data("world"), mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map(projection = "mollweide", xlim = c(-180, 180)) +
  ggtitle("Mollweide projection")
```

# Adding data to the map

## Points

Region boundaries serve as the background in geospatial data visualization - so now we need to add data. Some types of channels (points and symbols) are overlaid on top of the boundaries, whereas other channels (fill) are incorporated into the region layer itself. Let's look at the first set of channels.

## Points

Let's use our `usa2` map data to add some points. The `airports` data frame in the `nycflights13` package includes geographic info on airports in the United States.

```{r nycflights}
library(nycflights13)
airports
```

Each airport has it's geographic location encoded through `lat` and `lon`. To draw these points on the map, basically we draw a scatterplot with `x = lon` and `y = lat`. In fact we could simply do that:

```{r scatter}
ggplot(airports, aes(lon, lat)) +
  geom_point()
```

Let's overlay it with the mapped state borders:

```{r flights-usa}
ggplot() + 
  coord_map() + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "gray") +
  geom_point(data = airports, aes(x = lon, y = lat), shape = 1)
```

Slight problem. We have airports listed outside of the continental United States. There are a couple ways to rectify this. Unfortunately `airports` does not include a variable identifying state so the `filter()` operation is not that simple. The easiest solution is to crop the limits of the graph to only show the mainland:

```{r crop}
ggplot() + 
  coord_map(xlim = c(-130, -60),
            ylim = c(20, 50)) + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "gray") +
  geom_point(data = airports, aes(x = lon, y = lat), shape = 1)
```

If we want to change the projection method, the points will automatically adjust too:

```{r crop-proj}
ggplot() + 
  coord_map(projection = "albers", lat0 = 25, lat1 = 50,
            xlim = c(-130, -60),
            ylim = c(20, 50)) + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "gray") +
  geom_point(data = airports, aes(x = lon, y = lat), shape = 1)
```

## Symbols

We can change the size or type of symbols on the map. For instance, we can draw a **bubble plot** (also known as a **proportional symbol map**) and encode the altitude of the airport through the size channel:

```{r airport-alt}
ggplot() + 
  coord_map(xlim = c(-130, -60),
            ylim = c(20, 50)) + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "white") +
  geom_point(data = airports, aes(x = lon, y = lat, size = alt),
             fill = "grey", color = "black", alpha = .2) +
  theme_void() +
  theme(legend.position = "none")
```

Circle area is proportional to the airport's altitude (in feet). Or we could scale it based on the number of arriving flights in `flights`:

```{r airport-dest}
airports_n <- flights %>%
  count(dest) %>%
  left_join(airports, by = c("dest" = "faa"))

ggplot() + 
  coord_map(xlim = c(-130, -60),
            ylim = c(20, 50)) + 
  geom_polygon(data = usa2, mapping = aes(x = long, y = lat, group = group),
               color = "black", fill = "white") +
  geom_point(data = airports_n, aes(x = lon, y = lat, size = n),
             fill = "grey", color = "black", alpha = .2) +
  theme_void() +
  theme(legend.position = "none")
```

> `airports` contains a list of virtually all commercial airports in the United States. However `flights` only contains data on flights departing from New York City airports (JFK, LaGuardia, or Newark) and only services a few airports around the country.

## Fill (choropleths)

**Choropleth maps** encode information by assigning shades of colors to defined areas on a map (e.g. countries, states, counties, zip codes). There are lots of ways to tweak and customize these graphs, which is generally a good idea because remember that color is one of the harder-to-decode channels.

### With shapefiles

We'll continue to use the `usa2` shapefile. Let's reload it and also load and tidy the county shapefile:

```{r county-shp}
usa <- readOGR("data/census_bureau/cb_2013_us_state_20m/cb_2013_us_state_20m.shp")

usa2 <- usa %>%
  fortify(region = "GEOID") %>%
  as_tibble %>%
  left_join(usa@data, by = c("id" = "GEOID")) %>%
  # filter out Alaska, Hawaii, Puerto Rico via FIPS codes
  filter(!(STATEFP %in% c("02", "15", "72")))

counties <- readOGR("data/census_bureau/cb_2013_us_county_20m/cb_2013_us_county_20m.shp")

counties2 <- counties %>%
  fortify(region = "GEOID") %>%
  as_tibble %>%
  left_join(counties@data, by = c("id" = "GEOID")) %>%
  # filter out Alaska, Hawaii, Puerto Rico via FIPS codes
  filter(!(STATEFP %in% c("02", "15", "72")))

ggplot(counties2, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map()
```

We'll draw choropleths for the number of foreign-born individuals in each region (state or county). We can get those files from the `census_bureau` folder. Let's also normalize our measure by the total population to get the rate of foreign-born individuals in the population:

```{r import-foreign}
(fb_state <- read_csv("data/census_bureau/ACS_13_5YR_B05012_state/ACS_13_5YR_B05012.csv") %>%
  mutate(rate = HD01_VD03 / HD01_VD01))

(fb_county <- read_csv("data/census_bureau/ACS_13_5YR_B05012_county/ACS_13_5YR_B05012.csv") %>%
  mutate(rate = HD01_VD03 / HD01_VD01))
```

#### Joining the data to regions

Now that we have our data, we want to draw it on the map. To do that, we have to join together our data sources - the shapefiles and the CSVs. Normally joining data files requires a `_join()` operation of some sort. However when using `ggplot2`, we don't have to do this. Remember that we can pass different data frames into different layers of a `ggplot()` object. Rather than using `geom_polygon()` to draw our maps, now we switch to `geom_map()`:

```{r geom-map-state}
ggplot(fb_state, aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate), map = usa2) +
  expand_limits(x = usa2$long, y = usa2$lat)
```

Let's break down what just happened:

* `fb_state` is the data frame with the variables we want to visualize
* `map_id = GEO.id2` identifies the column in `fb_state` that uniquely matches each observation to a region in `usa2`
* `geom_map(aes(fill = rate), map = usa2) `
    * `fill = rate` identifies the column in `fb_state` we will use to determine the color of each region
    * `map = usa2` is the data frame containing the boundary coordinates
* `expand_limits(x = usa2$long, y = usa2$lat)` ensures the graph is drawn to the proper window. Because the default data frame for this `ggplot()` object is `fb_state`, it won't contain the necessary information to size the window

We can then tweak this up by adding a title, removing the background (but retaining the legend), and projecting the map using a different method:

```{r geom-map-state-neat}
ggplot(fb_state, aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate), map = usa2) +
  expand_limits(x = usa2$long, y = usa2$lat) +
  scale_fill_continuous(labels = scales::percent) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50)
```

We could do the same thing for the county-level data:

```{r geom-map-county}
ggplot(fb_county, aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate), map = counties2) +
  expand_limits(x = counties2$long, y = counties2$lat) +
  scale_fill_continuous(labels = scales::percent) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50)
```

### `sf`

If we want to plot feature attributes, the attributes are already joined with the geometric features. Just use the `fill` aesthetic:

```{r nc-fill}
ggplot() +
  geom_sf(data = nc, aes(fill = BIR74))
```

Because `nc` is a data frame, we can also apply basic reshaping functions to it. For instance, what if we want a facet plot of the SID rate in 1974 and 1979?

```{r nc-facet}
(nc2 <- nc %>%
  select(SID74, SID79, geometry) %>%
  gather(VAR, SID, -geometry))

ggplot() +
  geom_sf(data = nc2, aes(fill = SID)) +
  facet_wrap(~ VAR, ncol = 1)
```

# Selecting color palettes

Recall that Cleveland and McGill identify the color channel as one of the most difficult channels for humans to properly decode and interpret. Selection of your **color palette** is perhaps the most important decision to make when drawing a choropleth.

By default, `ggplot2` picks evenly spaced hues around the [Hue-Chroma-Luminance (HCL) color space](https://en.wikipedia.org/wiki/HCL_color_space):^[Check out chapter 6.6.2 in *`ggplot2`: Elegant Graphics for Data Analysis* for a much more thorough explanation of the theory behind this selection process]

```{r color-wheel}
# generate simulated data points
sim_points <- data_frame(x = factor(1:6))

plots <- purrr::map(1:6, ~ ggplot(sim_points[1:.x, ], aes(x, x, color = x)) +
  geom_point(size = 5) +
    ggtitle(paste(.x, "color")) +
  theme(legend.position = "none"))

gridExtra::marrangeGrob(plots, nrow = 2, ncol = 3, top = NULL)
```

`ggplot2` gives you many different ways of defining and customizing your `scale_color_` and `scale_fill_` palettes, but will not tell you if they are optimal for your specific usage in the graph.

### RColorBrewer

[Color Brewer](http://colorbrewer2.org/) is a diagnostic tool for selecting optimal color palettes for maps with discrete variables. The authors have generated different color palettes designed to make differentiating between categories easy depending on the scaling of your variable. All you need to do is define the number of categories in the variable, the nature of your data (sequential, diverging, or qualitative), and a color scheme. There are also options to select palettes that are colorblind safe, print friendly, and photocopy safe. Depending on the combination of options, you may not find any color palette that matches your criteria. In such a case, consider reducing the number of data classes.

#### Sequential

```{r cb-seq}
fb_county %>%
  mutate(rate_cut = cut_number(rate, 6)) %>%
  ggplot(aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate_cut), map = counties2) +
  expand_limits(x = counties2$long, y = counties2$lat) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  scale_fill_brewer(palette = "BuGn")

fb_county %>%
  mutate(rate_cut = cut_number(rate, 6)) %>%
  ggplot(aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate_cut), map = counties2) +
  expand_limits(x = counties2$long, y = counties2$lat) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  scale_fill_brewer(palette = "YlGn")

fb_county %>%
  mutate(rate_cut = cut_number(rate, 6)) %>%
  ggplot(aes(map_id = GEO.id2)) +
  geom_map(aes(fill = rate_cut), map = counties2) +
  expand_limits(x = counties2$long, y = counties2$lat) +
  labs(title = "Rate of foreign-born individuals in the population",
       fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  scale_fill_brewer(palette = "Blues")
```

### Qualitative

```{r qual}
state_data <- data_frame(name = state.name,
                         region = state.region,
                         subregion = state.division,
                         abb = state.abb) %>%
  bind_cols(as_tibble(state.x77)) %>%
  # get id variable into data frame
  left_join(usa2 %>%
              select(id, NAME) %>%
              distinct,
            by = c("name" = "NAME")) %>%
  # remove Alaska and Hawaii
  na.omit

# set region base plot
region_p <- ggplot(state_data, aes(map_id = id)) +
  geom_map(aes(fill = region), map = usa2) +
  expand_limits(x = usa2$long, y = usa2$lat) +
  labs(fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50)
region_p

# try different color brewers
region_p +
  scale_fill_brewer(palette = "Paired")

region_p +
  scale_fill_brewer(palette = "Dark2")

region_p +
  scale_fill_brewer(palette = "Pastel2")

# set subregion base plot
subregion_p <- ggplot(state_data, aes(map_id = id)) +
  geom_map(aes(fill = subregion), map = usa2) +
  expand_limits(x = usa2$long, y = usa2$lat) +
  labs(fill = NULL) +
  ggthemes::theme_map() +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50)
subregion_p

subregion_p +
  scale_fill_brewer(palette = "Paired")

subregion_p +
  scale_fill_brewer(palette = "Set1")

subregion_p +
  scale_fill_brewer(palette = "Pastel1")
```

# Session Info {.toc-ignore}

```{r cache = FALSE}
devtools::session_info()
```
