---
title: "Grammar of graphics and `ggplot2`"
output:
  html_document:
    highlight: pygments
    theme: readable
    toc: yes
    toc_float: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Objectives

* Define the layered grammar of graphics
* Identify function naming scheme for `ggplot2` functions and elements of the layers
* Review how to build a complex graph using `ggplot2`

```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(lubridate)
library(ggmap)
library(ggrepel)
library(gridExtra)
library(pander)

options(digits = 3)
set.seed(1234)
```

# The layered grammar of graphics

Google defines a **grammar** as "the whole system and structure of a language or of languages in general, usually taken as consisting of syntax and morphology (including inflections) and sometimes also phonology and semantics".^[[Google](https://www.google.com/search?q=grammar)] Others consider a grammar to be "the fundamental principles or rules of an art or science".[^layer] Applied to visualizations, a **grammar of graphics** is a grammar used to describe and create a wide range of statistical graphics.^[[Wilkinson, Leland. (2005). *The Grammar of Graphics*. (UChicago authentication required)](http://link.springer.com.proxy.uchicago.edu/book/10.1007%2F0-387-28695-0)]

The **layered grammar of graphics** approach is implemented in [`ggplot2`](https://cran.r-project.org/web/packages/ggplot2/index.html), a widely used graphics library for R. All graphics in this library are built using a layered approach, building layers up to create the final graphic.

# Components of the layered grammar of graphics

* Layer
    * Data
    * Mapping
    * Statistical transformation (stat)
    * Geometric object (geom)
    * Position adjustment (position)
* Scale
* Coordinate system (coord)
* Faceting (facet)
* Defaults
    * Data
    * Mapping

## Layer

**Layers** are used to create the objects on a plot. They are defined by five basic parts:

1. Data
1. Mapping
1. Statistical transformation (stat)
1. Geometric object (geom)
1. Position adjustment (position)

Layers are typically related to one another and share many common features. For instance, multiple layers can be built using the same underlying data. An example would be a scattterplot overlayed with a smoothed regression line to summarize the relationship between the variables:

```{r layers, echo = FALSE}
set.seed(1234)

data_frame(x = runif(50),
           y = x + rnorm(50, 0, .2)) %>%
  ggplot(aes(x, y)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "A plot with two layers",
       subtitle = "A scatterplot and a smoothing line")
```

### Data and mapping

**Data** defines the source of the information to be visualized, but is independent from the other elements. So a layered graphic can be built which can utilize different data sources while keeping the other components the same. Here is a portion of a dataset contained in the `ggplot2` package: `mpg`.

```{r mpg}
head(mpg) %>%
  kable(caption = "Dataset of automobiles")
```

**Mapping** defines how the variables are applied to the graphic. So if we were graphing information from `mpg`, we might map a car's engine displacement to the $x$ position and highway mileage to the $y$ position.

```{r mapping}
mpg %>%
  select(displ, hwy) %>%
  rename(x = displ,
         y = hwy)
```

In `ggplot2`, data and mapping can be defined either in the initial call to create a graphic:

```r
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))
```

Or when you create a geometric object using a `geom_()` function:

```r
ggplot() +
  geom_(data = mpg, mapping = aes(x = displ, y = hwy))
```

### Statistical transformation

A **statistical transformation** (*stat*) transforms the data, generally by summarizing the information. For instance, in a bar graph you typically are not trying to graph the raw data because this doesn't make any inherent sense. Instead, you might summarize the data by graphing the total number of observations within a set of categories. Or if you have a dataset with many observations, you might transform the data into a smoothing line which summarizes the overall pattern of the relationship between variables.

A stat takes a dataset as input and returns a dataset as output, and so a stat can add new variables to the original dataset. So instead of graphing this data in its raw form:

```{r stat_raw}
mpg %>%
  select(cyl)
```

You would transform it to:

```{r stat_transform}
mpg %>%
  count(cyl)
```

> Sometimes you don't need to make a statistical transformation. For example, in a scatterplot you use the raw values for the $x$ and $y$ variables to map onto the graph. In these situations, the statistical transformation is an *identity* transformation - the stat simply passes in the original dataset and exports the exact same dataset.

Statistical transformation functions are called using the `stat_()` syntax. Used outside of a `ggplot` object, their output is rather meaningless.

```{r stat-count}
stat_output <- stat_count(data = mpg, mapping = aes(x = cyl))
class(stat_output)
stat_output
```

### Geometric objects

**Geometric objects** (*geoms*) control the type of plot you create. Geoms are classified by their dimensionality:

* 0 dimensions - point, text
* 1 dimension - path, line
* 2 dimensions - polygon, interval

> A geom is the `ggplot2` name for a "mark".

Each geom can only display certain aesthetics. For example, a point geom has position, color, shape, and size aesthetics.

```{r geom_point}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point() +
  ggtitle("A point geom with position and color aesthetics")
```

In this example, the aesthetics for shape and size are not used, so instead they take on a constant value for all observations.

A bar geom has position, height, width, and fill.

```{r geom_bar}
ggplot(mpg, aes(cyl)) +
  geom_bar() +
  ggtitle("A bar geom with position and height aesthetics")
```

Again, with this example the fill aesthetic is not used.

Geometric objects are created using `geom_()` functions. The type of variable(s) you seek to map will dictate which kinds of `geom_()` functions you can use. For instance, you cannot create a `geom_histogram()` if your variable is disrete. Instead, you need to use `geom_bar()`.

### Position adjustment

Sometimes with dense data we need to adjust the position of elements on the plot, otherwise data points might obscure one another. Bar plots frequently stack or dodge the bars to avoid overlap:

```{r position_dodge}
count(mpg, class, cyl) %>%
  ggplot(aes(cyl, n, fill = class)) +
  geom_col(position = "identity") +
  ggtitle("An identity bar chart")

count(mpg, class, cyl) %>%
  ggplot(aes(cyl, n, fill = class)) +
  geom_col() +
  ggtitle("A stacked bar chart")

count(mpg, class, cyl) %>%
  ggplot(aes(cyl, n, fill = class)) +
  geom_col(position = "dodge") +
  ggtitle("A dodged bar chart")
```

Sometimes scatterplots with few unique $x$ and $y$ values are **jittered** (random noise is added) to reduce overplotting.

```{r position}
ggplot(mpg, aes(cyl, hwy)) +
  geom_point() +
  ggtitle("A point geom with obscured data points")

ggplot(mpg, aes(cyl, hwy)) +
  geom_jitter() +
  ggtitle("A point geom with jittered data points")
```

Position adjustments are defined as functions in R (e.g. `position_()`), however the normal syntax is to specify an adjustment as an argument to the `geom_()` function, like:

```r
geom_bar(position = "fill")
geom_bar(position = position_fill())
```

Typically you only directly call the function if you want to specify non-default arguments to that position function.

## Scale

A **scale** controls how data is mapped to aesthetic attributes, so we need one scale for every aesthetic property employed in a layer. For example, this graph defines a scale for color:

```{r scale_color}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point()
```

Note that the scale is consistent - every point for a compact car is drawn in tan, whereas SUVs are drawn in pink. The scale can be changed to use a different color palette:

```{r scale_color_palette}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point() +
  scale_color_brewer(palette = "Dark2")
```

Now we are using a different palette, but the scale is still consistent: all compact cars utilize the same color, whereas SUVs use a different color **but each SUV uses the same color**.

> A scale is the `ggplot2` name for a "channel".

Scales are found in the `scale_()` family of functions.

## Coordinate system

A **coordinate system** (*coord*) maps the position of objects onto the plane of the plot, and control how the axes and grid lines are drawn. Plots typically use two coordinates ($x, y$), but could use any number of coordinates. Most plots are drawn using the [**Cartesian coordinate system**](https://en.wikipedia.org/wiki/Cartesian_coordinate_system):

```{r coord_cart}
x1 <- c(1, 10)
y1 <- c(1, 5)
p <- qplot(x1, y1, geom = "blank", xlab = NULL, ylab = NULL) +
  theme_bw()
p 
```

This system requires a fixed and equal spacing between values on the axes. That is, the graph draws the same distance between 1 and 2 as it does between 5 and 6. The graph could be drawn using a [**semi-log coordinate system**](https://en.wikipedia.org/wiki/Semi-log_plot) which logarithmically compresses the distance on an axis:

```{r coord_semi_log}
p + coord_trans(y = "log10")
```

Or could even be drawn using [**polar coordinates**](https://en.wikipedia.org/wiki/Polar_coordinate_system):

```{r coord_polar}
p + coord_polar()
```

When drawing maps in `ggplot2`, you will need to employ a map projection coordinate system:

```{r map-proj}
library(fiftystater)
data("fifty_states")

ggplot(data = fifty_states, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  ggtitle("A cartesian coordinate system")

ggplot(data = fifty_states, mapping = aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black", fill = "gray") +
  coord_map(projection = "albers", lat0 = 25, lat1 = 50) +
  ggtitle("Albers equal-area projection")
```

## Faceting

**Faceting** can be used to split the data up into subsets of the entire dataset. This is a powerful tool when investigating whether patterns are the same or different across conditions, and allows the subsets to be visualized on the same plot (known as **conditioned** or **trellis** plots). The faceting specification describes which variables should be used to split up the data, and how they should be arranged.

```{r facet}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~ class)
```

`ggplot2` contains two faceting functions:

* `facet_grid()` - form a matrix of panels defined by row and column faceting variables. Works best with two discrete variables, and all combinations of the variables exist in the data.
* `facet_wrap()` - wraps a one dimensional sequence of panels into two dimensions. Automatically adjusts the number of panels per row to maximize the use of horizontal and vertical space.

Why use `facet_grid()` instead of `facet_wrap()`?^[Example drawn from [this StackOverflow thread](https://stackoverflow.com/questions/20457905/whats-the-difference-between-facet-wrap-and-facet-grid-in-ggplot2).] `facet_grid(x ~ y)` will display $x \times y$ plots even if some plots are empty. For example:

```{r facet-grid}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  facet_grid(cyl ~ class)
```

There are 4 distinct `cyl` and 7 distinct `class` values. This plot displays $4 \times 7 = 28$ plots, even if some are empty (because some classes do not have corresponding cylinder values, like rows with `class = "midsize"` doesn't have any corresponding `cyl = 5` value ).

`facet_wrap(x ~ y)` displays only the plots having actual values.

```{r facet-wrap}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~ cyl + class)
```

There are 19 plots displayed now, one for every combination of `cyl` and `class`.

## Themes

**Themes** control the display of all non-data elements of the plot. To modify individual elements, use `theme()`. To select between different premade themes, use the appropriate `theme_()` function.

```{r themes}
sample_plot <- ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point() +
  scale_color_brewer(palette = "Dark2")

sample_plot +
  theme_gray() +
  ggtitle("Standard gray theme")

sample_plot +
  theme_bw() +
  ggtitle("Black and white theme")

sample_plot +
  theme_light() +
  ggtitle("Light theme")

sample_plot +
  theme_dark() +
  ggtitle("Dark theme")

sample_plot +
  theme_minimal() +
  ggtitle("Minimal theme")

sample_plot +
  theme_void() +
  ggtitle("Void theme")

sample_plot +
  ggthemes::theme_excel() +
  labs(title = "Excel theme",
       caption = "From the ggthemes package")
```

## Defaults

Rather than explicitly declaring each component of a layered graphic (which will use more code and introduces opportunities for errors), we can establish intelligent defaults for specific geoms and scales. For instance, whenever we want to use a bar geom, we can default to using a stat that counts the number of observations in each group of our variable in the $x$ position.

```{r default}
ggplot() +
  layer(
    data = mpg, mapping = aes(x = displ, y = hwy),
    geom = "point", stat = "identity", position = "identity"
  ) +
  scale_x_continuous() +
  scale_y_continuous() +
  coord_cartesian()
```

The above code:

* Creates a new plot object (`ggplot`)
* Adds a layer (`layer`)
    * Specifies the data (`mpg`)
    * Maps engine displacement to the $x$ position and highway mileage to the $y$ position (`mapping`)
    * Uses the point geometric transformation (`geom = "point"`)
    * Implements an identity transformation and position (`stat = "identity"` and `position = "identity"`)
* Establishes two continuous position scales (`scale_x_continuous` and `scale_y_continuous`)
* Declares a cartesian coordinate system (`coord_cartesian`)

How can we simplify this using intelligent defaults?

1. We only need to specify one geom and stat, since each geom has a default stat.
1. Cartesian coordinate systems are most commonly used, so it is the default.
1. Default scales can be added based on the aesthetic and type of variables.
    * Continuous values are transformed with a linear scaling.
    * Discrete values are mapped to integers.
    * Scales for aesthetics such as color, fill, and size can also be intelligently defaulted.

Using these defaults, we can rewrite the above code as:

```{r default2}
ggplot() +
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy))
```

Because multiple layers can use the same components (data, mapping, etc.), we can specify that information in the `ggplot` function rather than in the `layer` function:

```{r default3}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point()
```

And function arguments in R use specific ordering, so we can omit the explicit call to `data` and `mapping`:

```{r default4}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

With this specification, it is easy to build the graphic up with additional layers, without modifying the original code:

```{r default5}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth()
```

Because we called `aes(x = displ, y = hwy)` within the `ggplot` function, it is automatically passed along to both `geom_point()` and `geom_smooth()`. If we fail to do this, we get an error:

```{r default6, error = TRUE}
ggplot(mpg) +
  geom_point(aes(displ, hwy)) +
  geom_smooth()
```

# How to build a complicated, layered graphic

## Charles Minard's map of Napoleon's disastrous Russian campaign of 1812

![["Carte figurative des pertes successives en hommes de l'Armee Français dans la campagne de Russe 1812–1813" by Charles Joseph Minard](https://en.wikipedia.org/wiki/Charles_Joseph_Minard)](https://upload.wikimedia.org/wikipedia/commons/2/29/Minard.png)

The graphic is notable for its representation in two dimensions of six types of data: the number of Napoleon's troops; distance; temperature; the latitude and longitude; direction of travel; and location relative to specific dates.^[This exercise is drawn from [Wickham, Hadley. (2010) "A Layered Grammar of Graphics". *Journal of Computational and Graphical Statistics*, 19(1).](http://www.jstor.org.proxy.uchicago.edu/stable/25651297) and [Andrew Heiss's extension using `ggplot2`](https://github.com/andrewheiss/fancy-minard)]

## Building Minard's map in R

```{r minard_data}
# get data on troop movements, city names, and temperatures
troops <- read_table("data/minard-troops.txt")
cities <- read_table("data/minard-cities.txt")
temps <- read_table("data/minard-temps.txt") %>%
  mutate(date = dmy(date))

glimpse(troops)
glimpse(cities)
glimpse(temps)
```

### Grammar of Minard's graphic

**Exercise: Write out what the grammar of graphics will look for this graph, just focusing on the map portion.**

<details> 
  <summary>**Click here for solution**</summary>
  <p>
* Layer
    * Data - `troops`
    * Mapping
        * $x$ and $y$ - troop position (`lat` and `long`)
        * Size - `survivors`
        * Color - `direction`
    * Statistical transformation (stat) - `identity`
    * Geometric object (geom) - `path`
    * Position adjustment (position) - none
* Layer
    * Data - `cities`
    * Mapping
        * $x$ and $y$ - city position (`lat` and `long`)
        * Label - `city`
    * Statistical transformation (stat) - `identity`
    * Geometric object (geom) - `text`
    * Position adjustment (position) - none
* Scale
    * Size - range of widths for troop `path`
    * Color - colors to indicate advancing or retreating troops
* Coordinate system - map projection (Mercator or something else)
* Faceting - none
  </p>
</details>

### Create the troop movement layer

First we want to build the layer for the troop movement.^[Example drawn from [Exploring Minard's 1812 plot with `ggplot2`](https://github.com/andrewheiss/fancy-minard).] If we just include the troop movement information (ignoring `surviviors` and `direction`), we could write this:

```{r troops-1}
ggplot(data = troops,
       mapping = aes(x = long, y = lat, group = group)) +
  geom_path()
```

We can now add the `survivors` and `direction` aesthetics:

```{r troops-2}
ggplot(data = troops,
       mapping = aes(x = long, y = lat, group = group,
                     color = direction, size = survivors)) +
  geom_path()
```

The individual segments of the path don't fit together very well and leave big gaps. We can fix that by adding a rounded line ending to each segment.

```{r troops-3}
ggplot(data = troops,
       mapping = aes(x = long, y = lat, group = group,
                     color = direction, size = survivors)) +
  geom_path(lineend = "round")
```

> This is not adding a new channel, it is simply adjusting the appearance of the existing channel and how the lines visually appear at the end of each segment.

The size of the path hides the drama of the plot. Napoleon started the 1812 campaign with 422,000 troops and returned with only 10,000. `scale_size()` automatically uses a default range of `c(0,6)`. We can adjust the scale to allow for more variation in size, highlighting the devastation of the army:

```{r troops-4}
ggplot(data = troops,
       mapping = aes(x = long, y = lat, group = group,
                     color = direction, size = survivors)) +
  geom_path(lineend = "round") +
  scale_size(range = c(0.5, 15))
```

Finally, we can remove the labels, legends, and change the colors to match the shade of brown from Minard's original plot.

```{r troops-5}
ggplot(data = troops,
       mapping = aes(x = long, y = lat, group = group,
                     color = direction, size = survivors)) +
  geom_path(lineend = "round") +
  scale_size(range = c(0.5, 15)) + 
  scale_color_manual(values = c("#DFC17E", "#252523")) +
  labs(x = NULL,
       y = NULL) + 
  guides(color = FALSE,
         size = FALSE)
```

### Create the cities layer

Because this graph is actually a map (with vertical and horizontal spatial position corresponding to latitude and longitude), we can easily overlay geographic details like city names and locations. We can locate the city location using `geom_point()` and label them using `geom_text()`.

Because we are adding a second layer to the graph with a different data source, we can also move the `data` and `mapping` arguments to specific `geom_()` functions

```{r cities-1}
ggplot() +
  geom_path(data = troops,
            mapping = aes(x = long, y = lat, group = group,
                          color = direction, size = survivors),
            lineend = "round") +
  geom_point(data = cities, aes(x = long, y = lat)) +
  geom_text(data = cities, aes(x = long, y = lat, label = city)) +
  scale_size(range = c(0.5, 15)) + 
  scale_color_manual(values = c("#DFC17E", "#252523")) +
  labs(x = NULL,
       y = NULL) + 
  guides(color = FALSE,
         size = FALSE)
```

Doing this places the city names directly at the defined geographical location. This is a bit problematic and makes the graph harder to read. We can adjust the city names using the `vjust` argument to `geom_text()`:

```{r cities-2}
ggplot() +
  geom_path(data = troops,
            mapping = aes(x = long, y = lat, group = group,
                          color = direction, size = survivors),
            lineend = "round") +
  geom_point(data = cities, aes(x = long, y = lat)) +
  geom_text(data = cities, aes(x = long, y = lat, label = city),
            vjust = 1.5) +
  scale_size(range = c(0.5, 15)) + 
  scale_color_manual(values = c("#DFC17E", "#252523")) +
  labs(x = NULL,
       y = NULL) + 
  guides(color = FALSE,
         size = FALSE)
```

Not perfect, but better. We could also adjust the color and font of the text labels to make them more readable:

```{r cities-3}
ggplot() +
  geom_path(data = troops,
            mapping = aes(x = long, y = lat, group = group,
                          color = direction, size = survivors),
            lineend = "round") +
  geom_point(data = cities, aes(x = long, y = lat),
             color = "#DC5B44") +
  geom_text(data = cities, aes(x = long, y = lat, label = city),
            vjust = 1.5,
            color = "#DC5B44", family = "sans") +
  scale_size(range = c(0.5, 15)) + 
  scale_color_manual(values = c("#DFC17E", "#252523")) +
  labs(x = NULL,
       y = NULL) + 
  guides(color = FALSE,
         size = FALSE)
```

To finish this component, let's remove all the background noise (e.g. grid, gridlines, axis tick marks, axis tick labels). A quick solution is to use the `theme_void()` function:

```{r cities-4}
troops_cities <- ggplot() +
  geom_path(data = troops,
            mapping = aes(x = long, y = lat, group = group,
                          color = direction, size = survivors),
            lineend = "round") +
  geom_point(data = cities, aes(x = long, y = lat),
             color = "#DC5B44") +
  geom_text(data = cities, aes(x = long, y = lat, label = city),
            vjust = 1.5,
            color = "#DC5B44", family = "sans") +
  scale_size(range = c(0.5, 15)) + 
  scale_color_manual(values = c("#DFC17E", "#252523")) +
  labs(x = NULL,
       y = NULL) + 
  guides(color = FALSE,
         size = FALSE) +
  theme_void()
troops_cities
```

### Temperatures and time

So far we have four of the variables from Minard's original plot -- we're still missing the temperatures during the retreat and the days of the retreat. Minard put this information in a separate plot under the map, which is fairly easy to do with [`gridExtra`](https://cran.r-project.org/web/packages/gridExtra/index.html).

First we have to create the panel, which is a basic line graph with longitude along the x-axis and temperature along the y-axis, with text added at each point.

```{r temps-1}
ggplot(data = temps, aes(x = long, y = temp)) +
  geom_line() +
  geom_label(aes(label = temp), vjust = 1.5)
```

We can create a new variable for nicer labels, combining temperature with the date.

```{r temps-2}
temps <- temps %>%
  mutate(nice_label = str_c(temp, "°, ", month, ". ", day))

ggplot(data = temps, aes(x = long, y = temp)) +
  geom_line() +
  geom_label(aes(label = nice_label), vjust = 1.5)
```

We'll also clean up the theme, move the axis label to the right, and only include major horizontal gridlines.

```{r temps-3}
ggplot(data = temps, aes(x = long, y = temp)) +
  geom_line() +
  geom_label(aes(label = nice_label),
            family = "sans", size = 2.5) + 
  labs(x = NULL,
       y = "° Celsius") +
  scale_y_continuous(position = "right") +
  coord_cartesian(ylim = c(-35, 5)) +  # Add some space above/below
  theme_bw(base_family = "sans") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks = element_blank(),
        panel.border = element_blank())
```

When we overlay the two plots, we have to make sure the x-axes align, so we need to use the same x-axis limits used in `troops_cities`. Those limits are buried inside the plot object, the parts of which can be accessed with `ggplot_build()`:

```{r axis-limits}
xrange <- ggplot_build(troops_cities)$layout$panel_ranges[[1]]$x.range
xrange
```

```{r temps-4}
temps_plot <- ggplot(data = temps, aes(x = long, y = temp)) +
  geom_line() +
  geom_label(aes(label = nice_label),
            family = "sans", size = 2.5) + 
  labs(x = NULL,
       y = "° Celsius") +
  scale_x_continuous(limits = xrange) +
  scale_y_continuous(position = "right") +
  coord_cartesian(ylim = c(-35, 5)) +  # Add some space above/below
  theme_bw(base_family = "sans") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks = element_blank(),
        panel.border = element_blank())
temps_plot
```

### Combining the plots

Finally, we use functions in the `gridExtra` package to combine the two plots. The easiest way to combine plot objects with `gridExtra` is to use `grid.arrange()`, but doing so doesn't align the axes of the plot. For instance, look at these two example plots — they're no longer comparable vertically because the left side of the bottom plot extends to the edge of the plot, expanding under the long axis label in the top plot:

```{r align-example}
example.data <- data_frame(
  x = 1:10,
  y = rnorm(10)
)

plot1 <- ggplot(example.data, aes(x = x, y = y)) +
  geom_line() +
  labs(y = "This is a really\nreally really really\nreally tall label")

plot2 <- ggplot(example.data, aes(x = x, y = y)) +
  geom_line() +
  labs(y = NULL)

grid.arrange(plot1, plot2)
```

Insetad of using `grid.arrange`, we can use `gridExtra`'s special version of `rbind()` (or `cbind()`) for `ggplotGrob` objects. A **grid graphical object** (`grob`) is a description of a graphical item. These basic classes provide default behavior for validating, drawing, and modifying graphical objects. These are used by `grid` and `gridExtra` to redraw graphical objects. `ggplotGrob()` converts a `ggplot` object into a `grob` suitable for use with `gridExtra`.

```{r align-example-good}
plot.both <- gtable_rbind(ggplotGrob(plot1),
                          ggplotGrob(plot2))

grid::grid.newpage()
grid::grid.draw(plot.both)
```

Now that we can align plots correctly, we can combine the map and the temperature:

```{r align-minard}
both_plot <- gtable_rbind(ggplotGrob(troops_cities),
                          ggplotGrob(temps_plot))

grid::grid.newpage()
grid::grid.draw(both_plot)
```

They're aligned, but there's an obvious problem—the map is way too small and the temperatures are too tall. With `grid.arrange` it's possible to pass a vector of relative panel heights, which would let us shrink the bottom panel. While using `gtable_rbind()` does let us align the two plots, it doesn't provide an easy way to mess with panel heights. Following [this StackOverflow answer](https://stackoverflow.com/questions/24331107/the-perils-of-aligning-plots-in-ggplot/24333504#24333504), though, we can mess with the `ggplot` object and adjust the panels manually.

```{r adjust-panels}
# Identify which layout elements are panels
panels <- both_plot$layout$t[grep("panel", both_plot$layout$name)]
panels

# Let's try a 3:1 ratio
both_plot$heights[panels] <- unit(c(3, 1), "null")

grid::grid.newpage()
grid::grid.draw(both_plot)
```

# Session Info {.toc-ignore}

```{r cache = FALSE}
devtools::session_info()
```

[^layer]: [Wickham, Hadley. (2010) "A Layered Grammar of Graphics". *Journal of Computational and Graphical Statistics*, 19(1).](http://www.jstor.org.proxy.uchicago.edu/stable/25651297)

