---
title: "More `ggplot2`"
author: |
  | MACS 40700
  | University of Chicago
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(forcats)
library(stringr)
library(ggrepel)

library(socviz)

options(digits = 3)
set.seed(1234)

base_size <- 18
theme_set(theme_minimal(base_size = base_size))
```

## Group data and the group aesthetic {.scrollable}

```{r gapminder}
library(gapminder)

ggplot(data = gapminder,
       mapping = aes(x = year,
                     y = lifeExp)) +
  geom_line()
```

## Group data and the group aesthetic {.scrollable}

```{r gapminder-country}
ggplot(data = gapminder,
       mapping = aes(x = year,
                     y = lifeExp,
                     group = country)) +
  geom_line()
```

## Group data and the group aesthetic {.scrollable}

```{r gapminder-continent}
ggplot(data = gapminder,
       mapping = aes(x = lifeExp,
                     y = gdpPercap,
                     color = continent)) +
  geom_point()

ggplot(data = gapminder,
       mapping = aes(x = lifeExp,
                     y = gdpPercap,
                     color = continent,
                     # redundant
                     group = continent)) +
  geom_point()
```

## Statistical transformations and temporary variables {.scrollable}

```{r geom-smooth}
ggplot(data = gapminder,
       mapping = aes(x = year,
                     y = lifeExp)) +
  geom_line(aes(group = country), alpha = .1) +
  geom_smooth()
```

## Statistical transformations and temporary variables {.scrollable}

```{r geom-smooth-ols}
ggplot(data = gapminder,
       mapping = aes(x = year,
                     y = lifeExp)) +
  geom_line(aes(group = country), alpha = .1) +
  geom_smooth(method = "lm")
```

## Statistical transformations and temporary variables {.scrollable}

```{r bigregion}
ggplot(data = gss_sm,
            mapping = aes(x = bigregion)) +
  geom_bar()
```

## Statistical transformations and temporary variables {.scrollable}

```{r prop}
ggplot(data = gss_sm,
            mapping = aes(x = bigregion)) +
  geom_bar(aes(y = ..prop..))
```

## Statistical transformations and temporary variables {.scrollable}

```{r prop-correct}
ggplot(data = gss_sm,
            mapping = aes(x = bigregion,
                          group = 1)) +
  geom_bar(aes(y = ..prop..))
```

## Frequency plots {.scrollable}

```{r bar-color}
ggplot(data = gss_sm,
       mapping = aes(x = religion,
                     fill = religion)) +
  geom_bar()

ggplot(data = gss_sm,
       mapping = aes(x = religion, fill = religion)) +
  geom_bar() +
  guides(fill = FALSE) 
```

## The awkward way {.scrollable}

```{r region-religion}
ggplot(data = gss_sm,
       mapping = aes(x = bigregion,
                     fill = religion)) +
  geom_bar()
```

## The awkward way {.scrollable}

```{r region-religion-fill}
ggplot(data = gss_sm,
       mapping = aes(x = bigregion,
                     fill = religion)) +
  geom_bar(position = "fill")
```

## The awkward way {.scrollable}

```{r region-religion-dodge}
ggplot(data = gss_sm,
       mapping = aes(x = bigregion,
                     fill = religion)) +
  geom_bar(position = "dodge")
```

## The awkward way {.scrollable}

```{r region-religion-dodge-prop}
ggplot(data = gss_sm,
       mapping = aes(x = bigregion,
                     fill = religion)) +
  geom_bar(aes(y = ..prop..), position = "dodge")
```

## The awkward way {.scrollable}

```{r region-religion-dodge-prop-religion}
ggplot(data = gss_sm,
       mapping = aes(x = bigregion,
                     fill = religion)) +
  geom_bar(aes(y = ..prop..,
               group = religion), position = "dodge")
```

## The awkward way {.scrollable}

```{r region-religion-facet}
ggplot(data = gss_sm,
       mapping = aes(x = religion)) +
  geom_bar(aes(y = ..prop..,
               group = bigregion), position = "dodge") +
    facet_wrap(~ bigregion)
```

## Calculate manually {.scrollable}

```{r rel-by-region}
glimpse(gss_sm)

(rel_by_region <- gss_sm %>%
    group_by(bigregion, religion) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N),
           pct = round((freq * 100), 0)))
```

## Calculate manually {.scrollable}

```{r rel-by-region-plot}
ggplot(data = rel_by_region,
       mapping = aes(x = bigregion,
                     y = pct,
                     fill = religion)) +
  geom_col(position = "dodge2") +
  labs(x = "Region", y = "Percent", fill = "Religion") +
  theme(legend.position = "top")
```

## Calculate manually {.scrollable}

```{r rel-by-region-plot-facet}
ggplot(data = rel_by_region,
       mapping = aes(x = religion,
                     y = pct,
                     fill = religion)) +
  geom_col(position = "dodge2") +
  labs(x = "Region", y = "Percent", fill = "Religion") +
  guides(fill = FALSE) + 
  coord_flip() + 
  facet_grid(~ bigregion)
```

## Continuous variables by group or category

```{r}
glimpse(organdata)
```

## Scatterplot {.scrollable}

```{r}
ggplot(data = organdata,
            mapping = aes(x = year, y = donors)) +
  geom_point()
```

## Line plots {.scrollable}

```{r}
ggplot(data = organdata,
            mapping = aes(x = year, y = donors)) +
  geom_line(aes(group = country)) +
  facet_wrap(~ country)
```

## Boxplots {.scrollable}

```{r}
ggplot(data = organdata,
            mapping = aes(x = country, y = donors)) +
  geom_boxplot()
```

## `coord_flip()` {.scrollable}

```{r}
ggplot(data = organdata,
            mapping = aes(x = country, y = donors)) +
  geom_boxplot() +
  coord_flip()
```

## `reorder()` {.scrollable}

```{r}
ggplot(data = organdata,
       mapping = aes(x = reorder(country, donors, na.rm = TRUE),
                     y = donors)) +
  geom_boxplot() +
  labs(x = NULL) +
  coord_flip()
```

## Add color {.scrollable}

```{r}
ggplot(data = organdata,
       mapping = aes(x = reorder(country, donors, na.rm = TRUE),
                     y = donors, fill = world)) +
  geom_boxplot() +
  labs(x = NULL) +
  coord_flip() +
  theme(legend.position = "bottom")
```

## Strip chart {.scrollable}

```{r}
ggplot(data = organdata,
       mapping = aes(x = reorder(country, donors, na.rm = TRUE),
                     y = donors, color = world)) +
  geom_point() +
  labs(x = NULL) +
  coord_flip() +
  theme(legend.position = "bottom")
```

## Strip chart {.scrollable}

```{r}
ggplot(data = organdata,
       mapping = aes(x = reorder(country, donors, na.rm = TRUE),
                     y = donors, color = world)) +
  geom_jitter() +
  labs(x = NULL) +
  coord_flip() +
  theme(legend.position = "bottom")
```

## Calculate summary statistics {.scrollable}

```{r}
(by_country <- organdata %>%
  group_by(consent_law, country) %>%
  summarize(donors_mean = mean(donors, na.rm = TRUE),
            donors_sd = sd(donors, na.rm = TRUE),
            gdp_mean = mean(gdp, na.rm = TRUE),
            health_mean = mean(health, na.rm = TRUE),
            roads_mean = mean(roads, na.rm = TRUE),
            cerebvas_mean = mean(cerebvas, na.rm = TRUE)))
```

## Calculate summary statistics {.scrollable}

```{r}
(by_country <- organdata %>%
  group_by(consent_law, country) %>%
  summarize_if(is.numeric, funs(mean, sd), na.rm = TRUE) %>%
  ungroup())
```

## Draw the plot {.scrollable}

```{r}
ggplot(data = by_country,
       mapping = aes(x = donors_mean,
                     y = reorder(country, donors_mean),
                     color = consent_law)) +
  geom_point(size = 3) +
  labs(x = "Donor Procurement Rate",
       y = "", color = "Consent Law") +
  theme(legend.position = "top")
```

## Use facet instead of color {.scrollable}

```{r}
ggplot(data = by_country,
       mapping = aes(x = donors_mean,
                     y = reorder(country, donors_mean))) +
  geom_point(size = 3) +
  facet_wrap(~ consent_law, ncol = 1) +
  labs(x = "Donor Procurement Rate",
       y = "", color = "Consent Law")
```

## Use facet instead of color {.scrollable}

```{r}
ggplot(data = by_country,
       mapping = aes(x = donors_mean,
                     y = reorder(country, donors_mean))) +
  geom_point(size = 3) +
  facet_wrap(~ consent_law, scales = "free_y", ncol = 1) +
  labs(x = "Donor Procurement Rate",
       y = "", color = "Consent Law")
```

## Add standard deviation {.scrollable}

```{r}
ggplot(data = by_country,
       mapping = aes(x = reorder(country, donors_mean),
                     y = donors_mean)) +
  geom_pointrange(mapping = aes(ymin = donors_mean - donors_sd,
                                ymax = donors_mean + donors_sd)) +
  labs(x = "",
       y = "Donor Procurement Rate") +
  coord_flip()
```

## `geom_text()` {.scrollable}

```{r}
ggplot(data = by_country,
       mapping = aes(x = roads_mean,
                     y = donors_mean)) +
  geom_point() +
  geom_text(mapping = aes(label = country))
```

## `geom_text()` {.scrollable}

```{r}
ggplot(data = by_country,
       mapping = aes(x = roads_mean,
                     y = donors_mean)) +
  geom_point() +
  geom_text(mapping = aes(label = country), hjust = 0)
```

## `ggrepel::geom_text_repel()` {.scrollable}

```{r}
elections_historic %>%
  select(2:7)
```

```{r}
p_title <- "Presidential Elections: Popular & Electoral College Margins"
p_subtitle <- "1824-2016"
p_caption <- "Data for 2016 are provisional."
x_label <- "Winner's share of Popular Vote"
y_label <- "Winner's share of Electoral College Votes"

library(ggrepel)

ggplot(elections_historic, aes(x = popular_pct,
                               y = ec_pct,
                               label = winner_label)) +
  geom_hline(yintercept = 0.5, size = 1.4, color = "gray80") +
  geom_vline(xintercept = 0.5, size = 1.4, color = "gray80") +
  geom_point() +
  geom_text_repel() +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = x_label, y = y_label, title = p_title, subtitle = p_subtitle,
       caption = p_caption)
```

## Label outliers only {.scrollable}

```{r}
ggplot(data = by_country,
       mapping = aes(x = gdp_mean, y = health_mean)) +
  geom_point() +
  geom_text_repel(data = filter(by_country, gdp_mean > 25000),
                  mapping = aes(label = country))

ggplot(data = by_country,
       mapping = aes(x = gdp_mean, y = health_mean)) +
  geom_point() +
  geom_text_repel(data = filter(by_country,
                                gdp_mean > 25000 | health_mean < 1500 |
                                  country %in% "Belgium"),
                  mapping = aes(label = country))
```
