---
title: "Network visualization"
author: |
  | MACS 40700
  | University of Chicago
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE, include = FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(modelr)
library(forcats)
library(ggraph)
library(igraph)

options(digits = 3)
set.seed(1234)

base_size <- 22
theme_set(theme_minimal(base_size = base_size))
```

## Network

* Network/graph
* Nodes/verticies
* Links/edges

## Undirected graph

```{r undirected-edge, echo = FALSE}
simple_net <- data_frame(from = 1, to = 2) %>%
  graph_from_data_frame

ggraph(simple_net) +
  geom_edge_link() + 
  geom_node_label(aes(label = name), size = 5) +
  theme_graph(base_size = base_size)
```

## Directed graph

```{r directed-edge, echo = FALSE}
ggraph(simple_net) +
  geom_edge_link(arrow = arrow(length = unit(4, 'mm')), 
                 end_cap = circle(5, 'mm')) + 
  geom_node_label(aes(label = name), size = 5) +
  theme_graph(base_size = base_size)
```

## Political polarization

* Income inequality
* Gerrymandering/redistricting
* Activists involved in primary elections
* Political realignment in the South
* Electing more partisan members to Congress
* Shift by existing members to ideological poles
* Increasingly partisan media

## Political polarization

* Partisanship networks in U.S. House
* Varying incentives
* Cooperation on roll-call votes
* Nodes
* Edges
* Undirected network

## Visualizing polarization

* Cross-party pairs
* Same-party pairs
* Attributes

## {.scrollable}

![Figure 2 from [*The Rise of Partisanship in the U.S. House of Representatives*](http://www.mamartino.com/projects/rise_of_partisanship/)](images/journal.pone.0123507.g002.PNG)

## Force-directed placement

> [D3 example of a force-directed graph](https://bl.ocks.org/mbostock/4062045)

## {.scrollable}

![Figure 2 from [*The Rise of Partisanship in the U.S. House of Representatives*](http://www.mamartino.com/projects/rise_of_partisanship/)](images/journal.pone.0123507.g002.PNG)

## Love Actually

> [**Love Actually** trailer](https://www.youtube.com/watch?v=peiRtkXMYQ8&oref=https%3A%2F%2Fwww.youtube.com%2Fwatch%3Fv%3DpeiRtkXMYQ8&has_verified=1)

```{r lines, include = FALSE}
raw <- readLines("data/love_actually.txt")

lines <- data_frame(raw = raw) %>%
    filter(raw != "", !str_detect(raw, "(song)")) %>%
    mutate(is_scene = str_detect(raw, " Scene "),
           scene = cumsum(is_scene)) %>%
    filter(!is_scene) %>%
    separate(raw, c("speaker", "dialogue"), sep = ":", fill = "left") %>%
    group_by(scene, line = cumsum(!is.na(speaker))) %>%
    summarize(speaker = speaker[1], dialogue = str_c(dialogue, collapse = " "))
```

```{r cast, dependson = "lines", include = FALSE}
cast <- read_csv(url("http://varianceexplained.org/files/love_actually_cast.csv"))

lines <- lines %>%
    inner_join(cast) %>%
    mutate(character = paste0(speaker, " (", actor, ")"))
lines[460:465,] %>%
  knitr::kable(caption = "Example of tidied data")
```

## Speaker-by-scene matrix {.scrollable}

```{r speaker_scene_matrix, dependson = "cast"}
by_speaker_scene <- lines %>%
  count(scene, character)
by_speaker_scene

library(reshape2)
speaker_scene_matrix <- by_speaker_scene %>%
  acast(character ~ scene, fun.aggregate = length)

speaker_scene_matrix[1:5, 1:5]
dim(speaker_scene_matrix)
```

## Hierarchical clustering {.scrollable}

```{r h, dependson = "speaker_scene_matrix"}
norm <- speaker_scene_matrix / rowSums(speaker_scene_matrix)

h <- hclust(dist(norm, method = "manhattan"))

ggdendro::ggdendrogram(h)
```

## Hierarchical clustering {.scrollable}

```{r ordering, dependson = "h"}
ordering <- h$labels[h$order]
ordering
```

## Hierarchical clustering {.scrollable}

```{r scenes, dependson = "speaker_scene_matrix"}
scenes <- by_speaker_scene %>%
    filter(n() > 1) %>%        # scenes with > 1 character
    ungroup() %>%
    mutate(scene = as.numeric(factor(scene)),
           character = factor(character, levels = ordering))

ggplot(scenes, aes(scene, character)) +
    geom_point() +
    geom_path(aes(group = scene))
```

## Adjacency matrix {.scrollable}

```{r cooccur}
non_airport_scenes <- speaker_scene_matrix[, colSums(speaker_scene_matrix) < 10]
non_airport_scenes[1:5, 1:5]

cooccur <- non_airport_scenes %*% t(non_airport_scenes)
cooccur[1:5, 1:5]
```

## Adjacency matrix {.scrollable}

```{r cooccur-graphics}
heatmap(cooccur)
```

## Adjacency matrix {.scrollable}

```{r cooccur-ggplot}
cooccur %>%
  as_tibble %>%
  mutate(id1 = rownames(cooccur)) %>%
  gather(id2, n, -id1) %>%
  mutate_at(vars(id1, id2), funs(factor(., levels = ordering))) %>%
  ggplot(aes(id1, id2, fill = n)) +
  geom_tile() +
  scale_fill_continuous(low = "white", high = "red") +
  coord_fixed() +
  labs(x = NULL,
       y = NULL,
       fill = NULL) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Node-edge diagram {.scrollable}

```{r g, dependson = "cooccur", echo = TRUE}
g <- graph_from_adjacency_matrix(cooccur,
                                 weighted = TRUE,
                                 mode = "undirected",
                                 diag = FALSE)
g
```

## Node-edge diagram {.scrollable}

```{r node-edge}
ggraph(g) + 
  geom_edge_link(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none")
```

## R packages for networks

* [`igraph`](http://igraph.org/r/)
* [`ggnet2`](https://briatte.github.io/ggnet/)
* [`ggnetwork`](https://cran.r-project.org/web/packages/ggnetwork/index.html)
* [`ggraph`](https://github.com/thomasp85/ggraph)

## Layout {.scrollable}

```{r layouts, echo = FALSE}
ggraph(g) + 
  geom_edge_link(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  ggtitle("Default layout (Nicely) algorithm")

ggraph(g, layout = "kk") + 
  geom_edge_link(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  ggtitle("Kamada and Kawai spring-based algorithm")

ggraph(g, layout = "fr") + 
  geom_edge_link(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  labs(title = "Fruchterman-Reingold algorithm",
       subtitle = "Force-directed layout, 500 iterations")

ggraph(g, layout = "fr", niter = 100) + 
  geom_edge_link(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  labs(title = "Fruchterman-Reingold algorithm",
       subtitle = "Force-directed layout, 100 iterations")

ggraph(g, layout = "fr", niter = 10) + 
  geom_edge_link(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  labs(title = "Fruchterman-Reingold algorithm",
       subtitle = "Force-directed layout, 10 iterations")

ggraph(g, layout = "grid") + 
  geom_edge_link(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  ggtitle("Grid algorithm")

ggraph(g, layout = "star") + 
  geom_edge_link(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  ggtitle("Star algorithm")
```

```{r layout-circle}
ggraph(g, layout = "linear") + 
  geom_edge_arc(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  ggtitle("Linear algorithm")

ggraph(g, layout = "linear", circular = TRUE) + 
  geom_edge_arc(aes(edge_width = weight)) + 
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  scale_edge_width_continuous(range = c(.5, 3)) +
  theme_graph(base_size = base_size) +
  theme(legend.position = "none") +
  ggtitle("Star algorithm (circular)")
```

----

![Source: [Introduction to ggraph: Layouts](http://www.data-imaginist.com/2017/ggraph-introduction-layouts/)](https://www.data-imaginist.com/post/2017-02-06-ggraph-introduction-layouts_files/figure-html/unnamed-chunk-12.gif)

## Dendrograms {.scrollable}

```{r dendrograms}
dendrogram <- as.dendrogram(hclust(dist(iris[, 1:4])))

ggraph(dendrogram, 'dendrogram') + 
  geom_edge_elbow()

ggraph(dendrogram, 'dendrogram', circular = TRUE) + 
  geom_edge_elbow() + 
  coord_fixed()
```

## Nodes {.scrollable}

```{r scatter}
ggraph(g, layout = "kk") +
  geom_point(aes(x = x, y = y))

ggraph(g, layout = "kk") + 
  geom_edge_link() + 
  geom_node_point()
```

## `create_layout()`

```{r create-layout}
create_layout(g, layout = "kk")
```

## Nodes {.scrollable}

```{r geom-node}
ggraph(g, layout = "kk") + 
  geom_edge_link() + 
  geom_node_point()

ggraph(g, layout = "kk") + 
  geom_edge_link() + 
  geom_node_text(aes(label = name))

ggraph(g, layout = "kk") + 
  geom_edge_link() + 
  geom_node_label(aes(label = name))

ggraph(g, layout = "kk") + 
  geom_edge_link() + 
  geom_node_label(aes(label = name), repel = TRUE)
```

## Communicating attributes of nodes

```{r import-media-example, echo = FALSE}
# example files obtained from http://kateto.net/network-visualization

nodes <- read_csv("data/Dataset1-Media-Example-NODES.csv")
links <- read_csv("data/Dataset1-Media-Example-EDGES.csv")

links <- aggregate(links[,3], links[,-3], sum)
links <- links[order(links$from, links$to),]

net <- graph_from_data_frame(d = links, vertices = nodes, directed = TRUE)
net
```

## Communicating attributes of nodes {.scrollable}

```{r media-plot}
ggraph(net) +
  geom_edge_link() +
  geom_node_point() +
  geom_node_label(aes(label = media), repel = TRUE) +
  theme_graph(base_size = base_size)
```

## Color {.scrollable}

```{r node-color}
ggraph(net) +
  geom_edge_link() +
  geom_node_point(aes(color = type.label), size = 2) +
  geom_node_label(aes(label = media), repel = TRUE, size = 2) +
  theme_graph(base_size = base_size)
```

## Size {.scrollable}

```{r node-size}
ggraph(net) +
  geom_edge_link() +
  geom_node_point(aes(color = type.label, size = audience.size)) +
  geom_node_label(aes(label = media), repel = TRUE, size = 2) +
  theme_graph(base_size = base_size)
```

## Edges {.scrollable}

```{r edge-straight}
ggraph(net) +
  geom_edge_link() +
  geom_node_point(aes(color = type.label, size = audience.size)) +
  geom_node_label(aes(label = media), repel = TRUE, size = 2) +
  theme_graph(base_size = base_size)
```

```{r edge-fan}
ggraph(net) +
  geom_edge_fan() +
  geom_node_point(aes(color = type.label, size = audience.size)) +
  geom_node_label(aes(label = media), repel = TRUE, size = 2) +
  theme_graph(base_size = base_size)
```

## Decorating edges {.scrollable}

```{r edge-arrow}
ggraph(net) +
  geom_edge_fan(arrow = arrow(length = unit(4, 'mm'))) +
  geom_node_point(aes(color = type.label, size = audience.size)) +
  geom_node_label(aes(label = media), repel = TRUE, size = 2) +
  theme_graph(base_size = base_size)
```

```{r edge-arrow-adj}
ggraph(net) +
  geom_edge_fan(arrow = arrow(length = unit(4, 'mm')), 
                   end_cap = circle(3, 'mm')) +
  geom_node_point(aes(color = type.label, size = audience.size)) +
  geom_node_label(aes(label = media), repel = TRUE, size = 2) +
  theme_graph(base_size = base_size)
```

## Color {.scrollable}

```{r edge-color}
ggraph(net) +
  geom_edge_fan(arrow = arrow(length = unit(4, 'mm')), 
                   end_cap = circle(3, 'mm'),
                aes(color = type)) +
  geom_node_point(aes(color = type.label, size = audience.size)) +
  geom_node_label(aes(label = media), repel = TRUE, size = 2) +
  theme_graph(base_size = base_size)
```

```{r edge-color-adj}
ggraph(net) +
  geom_edge_fan(arrow = arrow(length = unit(4, 'mm')), 
                   end_cap = circle(3, 'mm'),
                aes(color = type)) +
  geom_node_point(aes(color = type.label, size = audience.size)) +
  geom_node_label(aes(label = media), repel = TRUE, size = 2) +
  theme_graph(base_size = base_size) +
  scale_color_brewer(palette = "Paired") +
  scale_edge_color_brewer(palette = "Dark2")
```

