---
title: "Visualizing scientific results"
author: |
  | MACS 40700
  | University of Chicago
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(forcats)
library(stringr)
library(ggrepel)
library(modelr)
library(plotly)

library(socviz)

options(digits = 3)
set.seed(1234)

base_size <- 18
theme_set(theme_minimal(base_size = base_size))
```

## Tables or graphs

* [Kastellec and Leoni (2007)](http://www.jstor.org.proxy.uchicago.edu/stable/pdf/20446574.pdf?refreqid=excelsior:0edf6622ccfff488e494518924b22f92)
* When to use tables vs graphs
* Paper vs. poster

## Smoothing lines

```{r diamonds-point, echo = FALSE}
p <- ggplot(diamonds, aes(carat, price)) +
  geom_point() +
  scale_y_continuous(labels = scales::dollar) +
  labs(x = "Carat size",
       y = "Price")
p
```

## Smoothing lines {.scrollable}

```{r diamonds-lm}
p +
  geom_smooth(method = "lm", se = FALSE)
```

## Smoothing lines {.scrollable}

```{r diamonds-gam}
p +
  geom_smooth(se = FALSE)
```

## Smoothing lines {.scrollable}

```{r loess, echo = FALSE, warning = FALSE, message = FALSE}
library(lattice)

mod <- loess(NOx ~ E, ethanol, degree = 1, span = .75)
fit <- augment(mod)

mod0 <- loess(NOx ~ E, ethanol, degree = 0, span = .75)
mod1 <- loess(NOx ~ E, ethanol, degree = 1, span = .75)
mod2 <- loess(NOx ~ E, ethanol, degree = 2, span = .75)

fit_all <- ethanol %>%
  gather_predictions(mod0, mod1, mod2) %>%
  mutate(model = factor(model, levels = c("mod0", "mod1", "mod2"),
                        labels = c("Constant", "Linear", "Quadratic")))

ggplot(fit_all, aes(E, NOx)) +
  geom_point() +
  geom_line(aes(y = pred, color = model)) +
  labs(title = "Local linear regression",
       x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J",
       color = "Regression")
```

## Smoothing lines {.scrollable}

```{r loess_buildup, dependson="loess", fig.show = "animate", echo = FALSE, warning = FALSE, message = FALSE}
dat <- ethanol %>%
  inflate(center = unique(ethanol$E)) %>%
  mutate(dist = abs(E - center)) %>%
  filter(rank(dist) / n() <= .75) %>%
  mutate(weight = (1 - (dist / max(dist)) ^ 3) ^ 3)

library(gganimate)

p <- ggplot(dat, aes(E, NOx)) +
  geom_point(aes(alpha = weight, frame = center)) +
  geom_smooth(aes(group = center, frame = center, weight = weight), method = "lm", se = FALSE) +
  geom_vline(aes(xintercept = center, frame = center), lty = 2) +
  geom_line(aes(y = .fitted), data = fit, color = "red") +
  labs(x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J")
gganimate(p, interval = .5)
```

## Smoothing lines {.scrollable}

```{r loess_span, dependson="loess", fig.show = "animate", echo = FALSE, warning = FALSE, message = FALSE}
spans <- c(.25, .5, .75, 1)

# create loess fits, one for each span
fits <- data_frame(span = spans) %>%
  group_by(span) %>%
  do(augment(loess(NOx ~ E, ethanol, degree = 1, span = .$span)))

# calculate weights to reproduce this with local weighted fits
dat <- ethanol %>%
  inflate(span = spans, center = unique(ethanol$E)) %>%
  mutate(dist = abs(E - center)) %>%
  filter(rank(dist) / n() <= span) %>%
  mutate(weight = (1 - (dist / max(dist)) ^ 3) ^ 3)

# create faceted plot with changing points, local linear fits, and vertical lines,
# and constant hollow points and loess fit
p <- ggplot(dat, aes(E, NOx)) +
  geom_point(aes(alpha = weight, frame = center)) +
  geom_smooth(aes(group = center, frame = center, weight = weight), method = "lm", se = FALSE) +
  geom_vline(aes(xintercept = center, frame = center), lty = 2) +
  geom_point(shape = 1, data = ethanol, alpha = .25) +
  geom_line(aes(y = .fitted, frame = E, cumulative = TRUE), data = fits, color = "red") +
  facet_wrap(~span) +
  ylim(0, 5) +
  ggtitle("x0 = ") +
  labs(x = "Equivalence ratio",
       y = "Concentration of nitrogen oxides in micrograms/J")

gganimate(p, interval = .5)
```

## Multiple smoothing lines {.scrollable}

```{r mult-lines}
library(gapminder)

ggplot(data = gapminder,
       mapping = aes(x = log(gdpPercap), y = lifeExp)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  geom_smooth(method = "lm", formula = y ~ splines::bs(x, df = 3)) +
  geom_smooth(method = "loess")
```

## Multiple smoothing lines {.scrollable}

```{r mult-lines-leg}
model_colors <- RColorBrewer::brewer.pal(3, "Set1")
model_colors

ggplot(data = gapminder,
       mapping = aes(x = log(gdpPercap), y = lifeExp)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", aes(color = "OLS", fill = "OLS")) +
  geom_smooth(method = "lm", formula = y ~ splines::bs(x, df = 3),
              aes(color = "Cubic Spline", fill = "Cubic Spline")) +
  geom_smooth(method = "loess",
              aes(color = "LOESS", fill = "LOESS")) +
  scale_color_manual(name = "Models", values = model_colors) +
  scale_fill_manual(name = "Models", values = model_colors) +
  theme(legend.position = "top")
```

## Scatterplot matricies {.scrollable}

```{r credit-import}
credit <- read_csv("data/Credit.csv") %>%
  # remove first ID column
  select(-X1)
names(credit) <- stringr::str_to_lower(names(credit))   # convert column names to lowercase
glimpse(credit)
```

## Scatterplot matricies {.scrollable}

```{r credit-scatter-matrix}
pairs(select_if(credit, is.numeric))
```

## Scatterplot matricies {.scrollable}

```{r credit-scatter-ggpairs}
library(GGally)

ggpairs(select_if(credit, is.numeric))
```

## Scatterplot matricies {.scrollable}

```{r credit-scatter-matrix-gender}
ggpairs(credit, mapping = aes(color = gender),
        columns = c("income", "limit", "rating", "cards", "age", "education", "balance"))
```

## Scatterplot matricies {.scrollable}

```{r credit-scatter-smoother}
ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = "smooth"
        )
)
```

## Scatterplot matricies {.scrollable}

```{r wrap}
ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = wrap("smooth", alpha = .1, color = "blue")
        )
)
```

## Scatterplot matricies {.scrollable}

```{r credit-scatter-point-smoother, message = FALSE, warning = FALSE}
scatter_smooth <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    # make data points transparent
    geom_point(alpha = .2) +
    # add default smoother
    geom_smooth(se = FALSE)
}

ggpairs(select_if(credit, is.numeric),
        lower = list(
          continuous = scatter_smooth
        )
)

ggpairs(credit, mapping = aes(color = gender),
        columns = c("income", "limit", "rating", "cards", "age", "education", "balance"),
        lower = list(
          continuous = scatter_smooth
        )
)
```

## Scatterplot matricies {.scrollable}

```{r diamonds-scatter-matrix, warning = FALSE, message = FALSE}
rcfss::scorecard %>%
  select(type:debt) %>%
  ggpairs
```

## Heatmap of correlation coefficients {.scrollable}

```{r mpg-cor}
(mpg_lite <- select_if(mpg, is.numeric))

(cormat <- mpg_lite %>%
  cor %>%
  round(2))
```

## Heatmap of correlation coefficients {.scrollable}

```{r cormat-tidy}
library(reshape2)
(melted_cormat <- melt(cormat))
```

## Heatmap of correlation coefficients {.scrollable}

```{r cormat-tile}
ggplot(melted_cormat, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile()
```

## Heatmap of correlation coefficients {.scrollable}

```{r cormat-reduce}
# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}

upper_tri <- get_upper_tri(cormat)
upper_tri
```

## Heatmap of correlation coefficients {.scrollable}

```{r cormat-tile-tidy}
melted_cormat <- melt(upper_tri, na.rm = TRUE)

ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white") +
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1)) +
 coord_fixed()
```

## Heatmap of correlation coefficients {.scrollable}

```{r reorder-cormat}
reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)

# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)

# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))+
  coord_fixed()

# Print the heatmap
print(ggheatmap)
```

## Heatmap of correlation coefficients {.scrollable}

```{r add-coef-heatmap}
ggheatmap + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "bottom")
```

## Heatmap of correlation coefficients {.scrollable}

```{r coef-heatmap-function}
cormat_heatmap <- function(data){
  # generate correlation matrix
  cormat <- round(cor(data), 2)
  
  # melt into a tidy table
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
  upper_tri <- get_upper_tri(cormat)
  
  # reorder matrix based on coefficient value
  reorder_cormat <- function(cormat){
    # Use correlation between variables as distance
    dd <- as.dist((1-cormat)/2)
    hc <- hclust(dd)
    cormat <-cormat[hc$order, hc$order]
  }
  
  cormat <- reorder_cormat(cormat)
  upper_tri <- get_upper_tri(cormat)
  
  # Melt the correlation matrix
  melted_cormat <- melt(upper_tri, na.rm = TRUE)
  
  # Create a ggheatmap
  ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
    geom_tile(color = "white")+
    scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                         midpoint = 0, limit = c(-1,1), space = "Lab", 
                         name="Pearson\nCorrelation") +
    theme_minimal()+ # minimal theme
    theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                     size = 12, hjust = 1))+
    coord_fixed()
  
  # add correlation values to graph
  ggheatmap + 
    geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      panel.grid.major = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      axis.ticks = element_blank(),
      legend.position = "bottom")
}

cormat_heatmap(select_if(mpg, is.numeric))
cormat_heatmap(select_if(credit, is.numeric))
cormat_heatmap(select_if(diamonds, is.numeric))
```

## Parallel coordinate plots {.scrollable}

```{r pcp, echo = FALSE}
ggparcoord(data = iris, columns = 1:4, groupColumn = 5)

# with the iris data, order the axes by overall class (Species) separation
# using the anyClass option
ggparcoord(data = iris, columns = 1:4, groupColumn = 5, order = "anyClass")

# add points to the plot, add a title, and use an alpha scalar to make the
# lines transparent
p <- ggparcoord(data = iris, columns = 1:4, groupColumn = 5, order = "anyClass", 
    showPoints = TRUE, title = "Parallel Coordinate Plot for the Iris Data", 
    alphaLines = 0.3)
p

# add some basic interactivity
ggplotly(p)
```

## Three-dimensional plots {.scrollable}

```{r add-dimensions}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point()
```

## Three-dimensional plots {.scrollable}

```{r vote96-model}
# import data
(vote <- rcfss::mental_health)

# estimate model
vote_glm <- glm(vote96 ~ age + educ, data = vote, family = "binomial")
tidy(vote_glm)

# extract predicted probabilities
vote_prob <- vote %>%
  data_grid(age = 18:89, educ = 0:20) %>%
  modelr::add_predictions(vote_glm) %>%
  # convert predicted values to probabilities
  mutate(prob = rcfss::logit2prob(pred))

ggplot(vote_prob, aes(age, educ, fill = prob)) +
  geom_tile() +
  scale_fill_gradient2(midpoint = .5, label = scales::percent) +
  labs(title = "Probability of voter turnout in 1996",
       x = "Age",
       y = "Education (in years)",
       fill = "Probability\nof voting")

# cleaner image using geom_raster and interpolate = TRUE
ggplot(vote_prob, aes(age, educ, fill = prob)) +
  geom_raster(interpolate = TRUE) +
  scale_fill_gradient2(midpoint = .5, label = scales::percent) +
  labs(title = "Probability of voter turnout in 1996",
       x = "Age",
       y = "Education (in years)",
       fill = "Probability\nof voting")
```

## `plotly` {.scrollable}

```{r 3d-vote}
plot_ly(vote_prob, x = ~age, y = ~educ, z = ~prob) %>%
  add_mesh()
```

## `plotly` {.scrollable}

```{r 3d-credit}
plot_ly(credit, x = ~limit, y = ~balance, z = ~income) %>%
  add_mesh()
```

## `plotly` {.scrollable}

```{r 3d-surface}
plot_ly(z = ~volcano) %>% add_surface()

volcano %>%
  melt %>%
  ggplot(aes(Var1, Var2, z = value)) +
  geom_contour(aes(color = ..level..))
```

## Joint distribution of sex and happiness {.scrollable}

```{r mosaic-happy, include = FALSE}
# Mosaic plot of happiness and education
library(productplots)
data("happy")

happy <- happy %>%
  na.omit
```

```{r happy-sex}
# contingency table
library(descr)
crosstab(happy$happy, happy$sex, prop.t = TRUE, plot = FALSE)
```

## Conditional distribution of sex given happiness {.scrollable}

```{r happy-sex-cond}
crosstab(happy$happy, happy$sex, prop.r = TRUE, plot = FALSE)
crosstab(happy$happy, happy$sex, prop.c = TRUE, plot = FALSE)
```

## Conditional distribution of happiness given sex and marginal distribution of sex {.scrollable}

```{r happy-sex-marg}
crosstab(happy$happy, happy$sex, prop.c = TRUE, prop.r = TRUE, plot = FALSE)
```

## `graphics::mosaicplot()` {.scrollable}

```{r graphics-mosaicplot}
mosaicplot(~ sex + happy, data = happy)
```

## `vcd::mosaic()` {.scrollable}

```{r vcd-mosaic}
library(vcd)
mosaic(~ happy + sex, happy)
```

## `productplots::prodplot()` {.scrollable}

```{r productplots}
# mosaic plot using productplots
prodplot(happy, ~ happy + sex)

# add color
prodplot(happy, ~ happy + sex) +
  aes(fill = happy) +
  theme(panel.grid = element_blank())

prodplot(happy, ~ happy + marital) +
  aes(fill = happy) +
  theme(legend.position = "none") +
  theme(panel.grid = element_blank())
```

## `productplots::prodplot()` {.scrollable}

```{r prop-barchart}
ggplot(happy, aes(marital, fill = happy)) +
  geom_bar(position = "fill")
```

## Dot plot for summary statistics {.scrollable}

```{r oj}
library(ISLR)

OJ_sum <- OJ %>%
  select(ends_with("MM"), ends_with("CH")) %>%
  gather(var, value) %>%
  group_by(var) %>%
  summarize(mean = mean(value),
            sd = sd(value),
            min = min(value),
            max = max(value),
            n = n())

# print the table
kable(OJ_sum)

# plot using a single dot plot
ggplot(OJ_sum, aes(x = fct_reorder(var, mean), y = mean)) +
  geom_linerange(aes(ymin = mean - 2 * sd,
                      ymax = mean + 2 * sd),
                  linetype = 2,
                 size = .25) +
  geom_linerange(aes(ymin = mean - sd,
                      ymax = mean + sd),
                  size = 1) +
  geom_point() +
  coord_flip() +
  labs(x = NULL,
       y = NULL)

# dodge based on OJ brand
OJ_sum %>%
  separate(var, into = c("var", "brand"), -3, remove = TRUE) %>%
  ggplot(aes(x = fct_reorder(var, mean), y = mean, color = brand)) +
  geom_linerange(aes(ymin = mean - 2 * sd,
                      ymax = mean + 2 * sd),
                  linetype = 2,
                 size = .25,
                 position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = mean - sd,
                      ymax = mean + sd),
                  size = 1,
                 position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5)) +
  coord_flip() +
  labs(x = NULL,
       y = NULL,
       color = "Brand")

# facet based on OJ brand
OJ_sum %>%
  separate(var, into = c("var", "brand"), -3, remove = TRUE) %>%
  ggplot(aes(x = fct_reorder(var, mean), y = mean)) +
  facet_grid(. ~ brand) +
  geom_linerange(aes(ymin = mean - 2 * sd,
                      ymax = mean + 2 * sd),
                  linetype = 2,
                 size = .25) +
  geom_linerange(aes(ymin = mean - sd,
                      ymax = mean + sd),
                  size = 1) +
  geom_point() +
  coord_flip() +
  labs(x = NULL,
       y = NULL,
       color = "Brand")
```

## Visualizing model results

* Present your findings in substantive terms
* Show your degree of confidence
* Show your data when you can

## Extracting model contents {.scrollable}

```{r load-gapminder}
gapminder
```

## Extracting model contents {.scrollable}

```{r lifeExp-by-country}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3)
```

## Extracting model contents {.scrollable}

```{r lifeExp-mod}
gapminder_mod <- lm(lifeExp ~ year, data = gapminder)
summary(gapminder_mod)

grid <- gapminder %>% 
  data_grid(year, country) 
grid

grid <- grid %>% 
  add_predictions(gapminder_mod) 
grid

ggplot(gapminder, aes(year, group = country)) +
  geom_line(aes(y = lifeExp), alpha = .2) +
  geom_line(aes(y = pred), data = grid, color = "red", size = 1)
```

## Extracting model contents {.scrollable}

```{r model-str}
str(gapminder_mod)
```

## `broom::tidy()` {.scrollable}

```{r tidy}
tidy(gapminder_mod)

tidy(gapminder_mod) %>%
  str()
```

## `broom::augment()` {.scrollable}

```{r augment}
augment(gapminder_mod) %>%
  as_tibble()
```

## `broom::glance()` {.scrollable}

```{r glance}
glance(gapminder_mod)
```

* [Supported models](https://github.com/dgrtwo/broom)

## `stargazer` {.scrollable}

```{r gapminder-lm, echo = FALSE}
tidy(gapminder_mod)
```

## `stargazer` {.scrollable}

```{r stargazer-single, results = "asis"}
library(stargazer)

stargazer(gapminder_mod, type = "html")
```

## `stargazer` {.scrollable}

```{r stargazer-many, results = "asis"}
gapminder_yr <- lm(lifeExp ~ year, data = gapminder)
gapminder_gdp <- lm(lifeExp ~ gdpPercap, data = gapminder)
gapminder_yr_gdp <- lm(lifeExp ~ year + gdpPercap, data = gapminder)
gapminder_gdp_year_lifeexp <- lm(gdpPercap ~ year + lifeExp, data = gapminder)

stargazer(gapminder_yr, gapminder_gdp, gapminder_yr_gdp, gapminder_gdp_year_lifeexp,
          type = "html")
```

## `stargazer` {.scrollable}

```{r stargazer-custom, results = "asis"}
stargazer(gapminder_yr, gapminder_gdp, gapminder_yr_gdp, gapminder_gdp_year_lifeexp,
          type = "html",
          dep.var.labels = c("Life expectancy", "GDP per capita"),
          covariate.labels = c("Year", "GDP per capita", "Life expectancy"),
          omit.stat = c("ser", "f"))
```

## Generate predicted values {.scrollable}

```{r gap-many}
out <- lm(formula = lifeExp ~ gdpPercap + pop + continent,
          data = gapminder)
summary(out)
```

## Generate predicted values {.scrollable}

```{r gap-df}
min_gdp <- min(gapminder$gdpPercap)
max_gdp <- max(gapminder$gdpPercap)
med_pop <- median(gapminder$pop)

pred_df <- expand.grid(gdpPercap = (seq(from = min_gdp,
                                        to = max_gdp,
                                        length.out = 100)),
                       pop = med_pop,
                       continent = c("Africa", "Americas",
                                     "Asia", "Europe", "Oceania")) %>%
  as_tibble

pred_df
```

## Generate predicted values {.scrollable}

```{r add-predict}
pred_out <- predict(object = out,
                    newdata = pred_df,
                    interval = "predict") %>%
  as_tibble
pred_out
```

## Generate predicted values {.scrollable}

```{r combine-pred}
pred_df <- bind_cols(pred_df, pred_out)
pred_df
```

## Generate predicted values {.scrollable}

```{r plot-pred}
ggplot(data = pred_df %>%
         filter(continent %in% c("Europe", "Africa")),
       aes(x = gdpPercap,
           y = fit, ymin = lwr, ymax = upr,
           color = continent,
           fill = continent,
           group = continent)) +
  geom_point(data = gapminder %>%
               filter(continent %in% c("Europe", "Africa")),
             aes(x = gdpPercap, y = lifeExp,
                 color = continent),
             alpha = 0.5,
             inherit.aes = FALSE) + 
  geom_line() +
  geom_ribbon(alpha = 0.2, color = FALSE) +
  scale_x_log10(labels = scales::dollar)
```

## Plot marginal effects {.scrollable}

```{r margins}
library(margins)
```

```{r gss_sm}
library(socviz)

glimpse(gss_sm)

# make moderates the reference category (i.e. 0)
gss_sm$polviews_m <- relevel(gss_sm$polviews, ref = "Moderate")

# build the model
out_bo <- glm(obama ~ polviews_m + sex*race,
              family = "binomial", data = gss_sm)
summary(out_bo)
```

## Plot marginal effects {.scrollable}

```{r ame}
bo_m <- margins(out_bo)
summary(bo_m)
```

## Plot marginal effects {.scrollable}

```{r clean}
bo_gg <- as_tibble(summary(bo_m))
prefixes <- c("polviews_m", "sex")
bo_gg$factor <- prefix_strip(bo_gg$factor, prefixes)
bo_gg$factor <- prefix_replace(bo_gg$factor, "race", "Race: ")

bo_gg %>%
  select(factor, AME, lower, upper) 
```

```{r ame-graph}
ggplot(data = bo_gg, aes(x = reorder(factor, AME),
                         y = AME, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, color = "gray80") +
  geom_pointrange() + coord_flip() +
  labs(x = NULL, y = "Average Marginal Effect") 
```
