---
title: "Science, art, or somewhere inbetween"
author: |
  | MACS 40700
  | University of Chicago
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(ggthemes)

options(digits = 3)
set.seed(1234)

base_size <- 18

theme_set(theme_gray(base_size = base_size))
```

## Tufte's world

* Core purpose of visualization is to communicate quantitative information
    * "Above all else show the data"
* Data-ink - non-erasable core of a graphic

    $$\text{Data-ink ratio} = \frac{\text{data-ink}}{\text{total ink used to print the graphic}}$$

* What should we consider to be part of the "data-ink"?

## What is integral?

```{r integral}
p <- ggplot(mpg, aes(cty, hwy)) +
  geom_point()
p
```

## What is integral?

```{r integral-void}
p + theme_void(base_size = base_size)
```

## Starting point

```{r integral-background}
p + theme_bw(base_size = base_size)
```

## Remove panel box

```{r integral-panelbox}
p + theme_bw(base_size = base_size) +
    theme(panel.background = element_blank(),
          panel.border = element_blank())
```

## Remove minor grid lines

```{r integral-minor}
p + theme_bw(base_size = base_size) +
    theme(panel.background = element_blank(),
          panel.border = element_blank(),
          strip.background = element_blank(),
          plot.background = element_blank(),
          axis.line = element_blank(),
          panel.grid.minor = element_blank())
```

## Remove all grid lines

```{r integral-major}
p + theme_bw(base_size = base_size) +
    theme(panel.background = element_blank(),
          panel.border = element_blank(),
          strip.background = element_blank(),
          plot.background = element_blank(),
          axis.line = element_blank(),
          panel.grid = element_blank())
```

## Remove tick marks

```{r integral-tick}
p + theme_bw(base_size = base_size) +
    theme(panel.background = element_blank(),
          panel.border = element_blank(),
          strip.background = element_blank(),
          plot.background = element_blank(),
          axis.line = element_blank(),
          panel.grid = element_blank(),
          axis.ticks = element_blank())
```

## Use serif font

```{r integral-serif}
p + theme_bw(base_size = base_size, base_family = "serif") +
    theme(panel.background = element_blank(),
          panel.border = element_blank(),
          strip.background = element_blank(),
          plot.background = element_blank(),
          axis.line = element_blank(),
          panel.grid = element_blank(),
          axis.ticks = element_blank())
```

## Chart junk

* Vibrating moire effects
* The grid
* The duck
* Asserts all chart junk is bad - on what basis?

## Line plot

```{r line}
x <- 1967:1977
y <- c(0.5, 1.8, 4.6, 5.3, 5.3, 5.7, 5.4, 5, 5.5, 6, 5)
d <- data_frame(x, y)

ggplot(d, aes(x, y)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(breaks = seq(1, 6, 1), label = sprintf("$%s", seq(300, 400, 20))) +
  labs(title = "Per capita budget expandures",
       x = "Year",
       y = "Per capita budget expandures\nin constant dollars")
```

## Minimal line plot

```{r minline}
ggplot(d, aes(x, y)) +
  geom_line() +
  geom_point(size = 3) +
  theme_tufte(base_size = base_size) +
  theme(axis.title = element_blank()) +
  geom_hline(yintercept = c(5, 6), lty = 2) +
  scale_y_continuous(breaks = seq(1, 6, 1), label = sprintf("$%s", seq(300, 400, 20))) +
  scale_x_continuous(breaks = x, label = x) +
  annotate(
    "text",
    x = c(1977, 1977.2),
    y = c(1.5, 5.5),
    adj = 1,
    family = "serif",
    label = c("Per capita\nbudget expandures\nin constant dollars", "5%")
  )
```

## Boxplot

```{r boxplot}
ggplot(quakes, aes(factor(mag), stations)) +
  geom_boxplot() +
  labs(title = "Fiji earthquakes",
       x = "Richter Magnitude",
       y = "Number of stations reporting earthquakes")
```

## Minimal boxplot

```{r minboxplot}
ggplot(quakes, aes(factor(mag), stations)) +
  theme_tufte(base_size = base_size) +
  geom_tufteboxplot(outlier.colour = "transparent") +
  theme(axis.title = element_blank()) +
  annotate(
    "text",
    x = 8,
    y = 120,
    adj = 1,
    family = "serif",
    label = c(
      "Number of stations \nreporting Richter Magnitude\nof Fiji earthquakes (n=1000)"
    )
  )
```

## Minimal boxplot

```{r offset-boxplot}
ggplot(quakes, aes(factor(mag), stations)) +
  theme_tufte(base_size = base_size) +
  geom_tufteboxplot(median.type = "line") +
  theme(axis.title = element_blank()) +
  annotate(
    "text",
    x = 8,
    y = 120,
    adj = 1,
    family = "serif",
    label = c(
      "Number of stations \nreporting Richter Magnitude\nof Fiji earthquakes (n=1000)"
    )
  )
```

## Barchart

```{r bar}
library(psych)
library(reshape2)

d <- melt(colMeans(msq[, c(2, 7, 34, 36, 42, 43, 46, 55, 68)], na.rm = T) *
            10)
d$trait <- rownames(d)

p <- ggplot(d, aes(x = trait, y = value)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(breaks = seq(1, 5, 1)) +
  labs(title = "Watson et al., 1998",
       subtitle = "N = 3896",
       x = "Negative emotion traits",
       y = "Average score")
p
```

## Default themes {.scrollable}

```{r default-themes, dependson="bar", echo = TRUE}
p + theme_bw(base_size = base_size)
p + theme_dark(base_size = base_size)
p + theme_classic(base_size = base_size)
p + theme_minimal(base_size = base_size)
p + theme_void(base_size = base_size)
```

## Minimal barchart

```{r minbar}
ggplot(d, aes(x = trait, y = value)) +
  theme_tufte(base_size = base_size, ticks = F) +
  geom_bar(width = 0.25, fill = "gray", stat = "identity") +
  theme(axis.title = element_blank()) +
  scale_y_continuous(breaks = seq(1, 5, 1)) +
  geom_hline(yintercept = seq(1, 5, 1),
             col = "white",
             lwd = 1) +
  annotate(
    "text",
    x = 3.5,
    y = 5,
    adj = 1,
    family = "serif",
    label = c(
      "Average scores\non negative emotion traits
from 3896 participants\n(Watson et al., 1988)"
    )
  )
```

## Bivariate scatterplot

```{r scatterplot}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  xlab("Car weight (lb/1000)") +
  ylab("Miles per gallon of fuel")
```

## Range-frame graph

```{r range-frame}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  geom_rangeframe() +
  theme_tufte(base_size = base_size) +
  xlab("Car weight (lb/1000)") +
  ylab("Miles per gallon of fuel") +
  theme(axis.title.x = element_text(vjust = -0.5),
        axis.title.y = element_text(vjust = 1.5))
```

## With a quartile plot

```{r range-frame-quartile}
library(devtools)
source_url('https://raw.githubusercontent.com/bearloga/Quartile-frame-Scatterplot/master/qfplot.R')

qfplot(
  x = mtcars$wt,
  y = mtcars$mpg,
  xlab = "Car weight (lb/1000)",
  ylab = "Miles per gallon of fuel"
)
```

## When is redundancy better?

![Traditional bar chart of crime in the city of San Francisco, 2009-10. Source: [Visualizing Time with the Double-Time Bar Chart](http://dougmccune.com/blog/2011/04/26/visualizing-time-with-the-double-time-bar-chart/)](http://dougmccune.com/blog/wp-content/uploads/2011/04/sk_time_chart1.png)

![Double-time bar chart of crime in the city of San Francisco, 2009-10. Source: [Visualizing Time with the Double-Time Bar Chart](http://dougmccune.com/blog/2011/04/26/visualizing-time-with-the-double-time-bar-chart/)](http://dougmccune.com/blog/wp-content/uploads/2011/04/burglary1.png)
    
## Minimalism in practice

![Chart from Harvard magazine. Source: [Involuntary head-shaking is probably not an intended consequence of data visualization](http://junkcharts.typepad.com/junk_charts/2016/11/involuntary-head-shaking-is-probably-not-an-intended-consequence-of-data-visualization.html)](http://junkcharts.typepad.com/.a/6a00d8341e992c53ef01b7c8b60cd9970b-pi)

## Minimalism in practice

![Redesigned chart from Harvard magazine. Source: [Involuntary head-shaking is probably not an intended consequence of data visualization](http://junkcharts.typepad.com/junk_charts/2016/11/involuntary-head-shaking-is-probably-not-an-intended-consequence-of-data-visualization.html)](http://junkcharts.typepad.com/.a/6a00d8341e992c53ef01b8d23fe1c7970c-pi)

## Experimental tests of Tufte's claims {.scrollable}

![Source: Figure 2 from [Bateman, Scott, et al. "Useful junk?: the effects of visual embellishment on comprehension and memorability of charts." *Proceedings of the SIGCHI Conference on Human Factors in Computing Systems*. ACM, 2010.](http://www.cedma-europe.org/newsletter%20articles/misc/The%20Effects%20of%20Visual%20Embellishment%20on%20Comprehension%20and%20Memorability%20of%20Charts.pdf)](images/tufte_experiment.png)

## Results {.scrollable}

![Source: Figures 4-6 from [Bateman, Scott, et al. "Useful junk?: the effects of visual embellishment on comprehension and memorability of charts." *Proceedings of the SIGCHI Conference on Human Factors in Computing Systems*. ACM, 2010.](http://www.cedma-europe.org/newsletter%20articles/misc/The%20Effects%20of%20Visual%20Embellishment%20on%20Comprehension%20and%20Memorability%20of%20Charts.pdf)](images/tufte_experiment_results.png)

## Rethinking Tufte's definition of visual excellence

* Too many of Tufte's claims are based on nothing
* Hockey stick chart vs. xkcd on Earth's temperature
* Experimental design
