---
title: "Text visualization"
author: |
  | MACS 40700
  | University of Chicago
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE, include = FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(modelr)
library(forcats)
library(tidytext)
library(rtweet)
library(wordcloud)
library(scales)

options(digits = 3)
set.seed(1234)

base_size <- 22
theme_set(theme_minimal(base_size = base_size))
```

## Basic workflow for text analysis

* Obtain your text sources
* Extract documents and move into a corpus
* Transformation
* Extract features
* Perform analysis

## Obtain your text sources

* Web sites
    * Twitter
* Databases
* PDF documents
* Digital scans of printed materials

## Extract documents and move into a corpus

* Corpus
* Character string

## Transformation

* Tagging segments of speech for part-of-speech (nouns, verbs, adjectives, etc.) or entity recognition (person, place, company, etc.)
* Standard text processing
    * Converting to lower case
    * Removing punctuation
    * Removing numbers
    * Removing stopwords
    * Removing domain-specific stopwords
    * Stemming

## Extract features

* Quantify your measures
* Bag-of-words model
* Term-document matrix
    * Each row is a document
    * Each column is a term
    * Each cell represents the frequency of the term appearing in the document
* Missing context

## Perform analysis

* Word frequency
* Collocation
* Dictionary tagging
* Document classification
* Corpora comparison
* Topic modeling

## Wordclouds {.scrollable}

```{r wordcloud-rstats}
library(wordcloud)

# get tweets
rt <- search_tweets(
  q = "#rstats",
  n = 3000,
  include_rts = FALSE
)
rt

# tokenize
rstats_token <- rt %>%
  unnest_tokens(word, text, token = "tweets") %>%
  filter(!word %in% stop_words$word)
glimpse(rstats_token)

# plot
rstats_token %>%
  count(word) %>%
  filter(word != "#rstats") %>%
  with(wordcloud(word, n, max.words = 100))
```

## Wordclouds {.scrollable}

```{r wordcloud-pope}
# get tweets
pope <- get_timelines("Pontifex",
                      n = 3200,
                      include_rts = FALSE)

# tokenize
pope_token <- pope %>%
  unnest_tokens(word, text, token = "tweets") %>%
  filter(!word %in% stop_words$word)
glimpse(pope_token)

# plot
pope_token %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

## Comparison cloud {.scrollable}

```{r wordcloud-pope-trump}
library(reshape2)

# get fresh trump tweets
trump <- get_timelines("realdonaldtrump",
                      n = 3200,
                      include_rts = FALSE)

# tokenize
trump_token <- trump %>%
  unnest_tokens(word, text, token = "tweets") %>%
  filter(!word %in% stop_words$word)

bind_rows(Trump = trump_token, Pope = pope_token, .id = "person") %>%
  count(word, person) %>%
  acast(word ~ person, value.var = "n", fill = 0) %>%
  comparison.cloud(max.words = 100, colors = c("blue", "red"))
```

## N-gram viewers

* N-gram
* [Google Books Ngram Viewer](https://books.google.com/ngrams)
    * [Calendar of Meaningful Dates](https://www.xkcd.com/1140/)
* [How The Internet* Talks](https://projects.fivethirtyeight.com/reddit-ngram/?keyword=triggered.safe_space.sjw.snowflake&start=20071015&end=20161231&smoothing=10)

## Geospatial visualization

* [Which Curse Words Are Popular In Your State? Find Out From These Maps.](http://www.huffingtonpost.com/entry/which-curse-words-are-popular-in-your-state_us_55a80662e4b04740a3df54b8)
* [Hate Map](http://users.humboldt.edu/mstephens/hate/hate_map.html)
* [Soda vs. Pop with Twitter](http://blog.echen.me/2012/07/06/soda-vs-pop-with-twitter/)

## Network analysis with text

* Use text features to identify edges between nodes in a network
* [How every #GameOfThrones episode has been discussed on Twitter](https://interactive.twitter.com/game-of-thrones/#?episode=1)

## Sentiment analysis

> I am happy

## Dictionaries

```{r}
get_sentiments("bing")
```

## Dictionaries

```{r}
get_sentiments("afinn")
```

## Dictionaries

```{r}
get_sentiments("nrc")
```

## Dictionaries

```{r}
get_sentiments("nrc") %>%
  count(sentiment)
```

## Measuring overall sentiment {.scrollable}

```{r}
library(janeaustenr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text,
                                     regex("^chapter [\\divxlc]",
                                           ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

janeaustensentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

ggplot(janeaustensentiment, aes(index, sentiment, fill = book)) +
        geom_bar(alpha = 0.8, stat = "identity", show.legend = FALSE) +
        facet_wrap(~book, ncol = 2, scales = "free_x")
```

## Exploring content of Donald Trump's Twitter timeline

* [Text analysis of Trump's tweets confirms he writes only the (angrier) Android half](http://varianceexplained.org/r/trump-tweets/)
* [Trump's Android and iPhone tweets, one year later](http://varianceexplained.org/r/trump-followup/)
