---
title: "Visualizing scientific results"
author: |
  | MACS 40700
  | University of Chicago
date: "April 26, 2017"
output: rcfss::cfss_slides
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages, cache = FALSE, message = FALSE, include = FALSE}
library(tidyverse)
library(knitr)
library(broom)
library(stringr)
library(modelr)
library(forcats)

options(digits = 3)
set.seed(1234)

base_size <- 22
theme_set(theme_minimal(base_size = base_size))
```

## Linear functional form

$$Y = \beta_0 + \beta_{1}X$$

```{r sim-plot}
ggplot(sim1, aes(x, y)) + 
  geom_point()
```

## Linear functional form

```{r sim-random-fit}
models <- tibble(
  a1 = runif(250, -20, 40),
  a2 = runif(250, -5, 5)
)

ggplot(sim1, aes(x, y)) + 
  geom_abline(aes(intercept = a1, slope = a2), data = models, alpha = 1/4) +
  geom_point()
```

## Least squares regression

```{r sim-error}
dist1 <- sim1 %>% 
  mutate(
    dodge = rep(c(-1, 0, 1) / 20, 10),
    x1 = x + dodge,
    pred = 7 + x1 * 1.5
  )

ggplot(dist1, aes(x1, y)) + 
  geom_abline(intercept = 7, slope = 1.5, color = "grey40") +
  geom_point(color = "grey40") +
  geom_linerange(aes(ymin = y, ymax = pred), color = "#3366FF")
```

## Linear functional form

```{r sim-lm}
sim1_mod <- lm(y ~ x, data = sim1)
```

```{r sim-lm-summary}
summary(sim1_mod)
```

## Linear functional form

```{r sim-lm-plot}
dist2 <- sim1 %>%
  add_predictions(sim1_mod) %>%
  mutate(
    dodge = rep(c(-1, 0, 1) / 20, 10),
    x1 = x + dodge
  )

ggplot(dist2, aes(x1, y)) + 
  geom_smooth(method = "lm", color = "grey40") +
  geom_point(color = "grey40") +
  geom_linerange(aes(ymin = y, ymax = pred), color = "#3366FF")
```

## Residuals and visualizations

```{r mpg-displ}
# estimate linear model
mpg_lin <- lm(hwy ~ displ, data = mpg)

# plot the model using geom_smooth()
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Residuals and visualizations

```{r mpg-drv}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  geom_smooth(method = "lm", se = FALSE)
```

## Residuals and visualizations

```{r mpg-resid-dist}
mpg %>%
  add_residuals(mpg_lin) %>%
  ggplot(aes(resid, color = drv)) +
  geom_density()
```

## Residuals and visualizations {.scrollable}

```{r mpg-drv-mod}
mpg_drv <- lm(hwy ~ displ + drv, data = mpg)

mpg %>%
  data_grid(displ, drv) %>%
  add_predictions(mpg_drv) %>%
  ggplot(aes(displ, pred, color = drv)) +
  geom_line() +
  geom_point(data = mpg, aes(displ, hwy))

mpg %>%
  add_residuals(mpg_drv) %>%
  ggplot(aes(resid, color = drv)) +
  geom_density()
```

## Non-linearity of the data

```{r mpg-nonlinear}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Non-linearity of the data {.scrollable}

```{r resid-plot}
# estimate models
mpg_lin <- lm(hwy ~ displ, data = mpg)
mpg_quad <- lm(hwy ~ poly(displ, 2, raw = TRUE), data = mpg)

mpg %>%
  add_predictions(mpg_lin) %>%
  add_residuals(mpg_lin) %>%
  ggplot(aes(pred, resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(alpha = .2) +
  geom_smooth(se = FALSE) +
  labs(title = "Residual plot for linear fit")

mpg %>%
  add_predictions(mpg_quad) %>%
  add_residuals(mpg_quad) %>%
  ggplot(aes(pred, resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(alpha = .2) +
  geom_smooth(se = FALSE) +
  labs(title = "Residual plot for quadratic fit")
```

----

```{r sim-homo}
sim_homo <- data_frame(x = runif(1000, 0, 10),
                       y = 2 + 3 * x + rnorm(1000, 0, 1))
sim_homo_mod <- glm(y ~ x, data = sim_homo)

sim_homo %>%
  add_predictions(sim_homo_mod) %>%
  add_residuals(sim_homo_mod) %>%
  ggplot(aes(pred, resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(alpha = .2) +
  geom_quantile(method = "rqss", lambda = 5, quantiles = c(.05, .95)) +
  labs(title = "Homoscedastic variance of error terms",
       x = "Predicted values",
       y = "Residuals")
```

----

```{r sim-hetero}
sim_hetero <- data_frame(x = runif(1000, 0, 10),
                       y = 2 + 3 * x + rnorm(1000, 0, (x / 2)))
sim_hetero_mod <- glm(y ~ x, data = sim_hetero)

sim_hetero %>%
  add_predictions(sim_hetero_mod) %>%
  add_residuals(sim_hetero_mod) %>%
  ggplot(aes(pred, resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(alpha = .2) +
  geom_quantile(method = "rqss", lambda = 5, quantiles = c(.05, .95)) +
  labs(title = "Heteroscedastic variance of error terms",
       x = "Predicted values",
       y = "Residuals")
```

## Outliers

```{r outlier-lm}
sim <- data_frame(x = rnorm(20),
                  y = -2 + x + rnorm(20)) %>%
  mutate(y = ifelse(row_number(x) == 15, y + 10, y),
         outlier = ifelse(row_number(x) == 15, TRUE, FALSE))

ggplot(sim, aes(x, y)) +
  geom_point(aes(color = outlier)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(data = filter(sim, outlier == FALSE),
              method = "lm", se = FALSE, linetype = 2) +
  scale_color_manual(values = c("black", "red")) +
  theme(legend.position = "none")
```

## Outliers {.scrollable}

```{r outlier-resid}
sim_lm <- lm(y ~ x, data = sim)

sim %>%
  add_predictions(sim_lm) %>%
  add_residuals(sim_lm) %>%
  ggplot(aes(pred, resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(aes(color = outlier)) +
  labs(x = "Fitted values",
       y = "Residuals") +
  scale_color_manual(values = c("black", "red")) +
  theme(legend.position = "none")

augment(sim_lm) %>%
  left_join(sim) %>%
  ggplot(aes(.fitted, .std.resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(aes(color = outlier)) +
  labs(x = "Fitted values",
       y = "Standardized residuals") +
  scale_color_manual(values = c("black", "red")) +
  theme(legend.position = "none")
```

## High-leverage points

$$\text{Influence} = \text{Leverage} \times \text{Discrepancy}$$

* Leverage
* Discrepancy

## High-leverage points

* Leverage statistic/hat value

    $$h_i = \frac{1}{n} + \frac{(x_i - \bar{x})^2}{\sum_{i'=1}^{n} (x_{i'} - \bar{x})^2}$$

* Residuals
* Cook's D

## High-leverage points

* Outcome of interest - number of federal laws struck down by SCOTUS
1. **Age** - the mean age of the members of the Supreme Court
1. **Tenure** - mean tenure of the members of the Court
1. **Unified** - a dummy variable indicating whether or not the Congress was controlled by the same party in that period

## High-leverage points {.scrollable}

```{r bubble}
library(haven)

dahl <- read_dta("data/LittleDahl.dta")
dahl_mod <-lm(nulls ~ age + tenure + unified, data = dahl)

dahl_augment <- dahl %>%
  mutate(hat = hatvalues(dahl_mod),
         student = rstudent(dahl_mod),
         cooksd = cooks.distance(dahl_mod))

# use size
ggplot(dahl_augment, aes(hat, student)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(aes(size = cooksd), shape = 1) +
  geom_text(data = dahl_augment %>%
              arrange(-cooksd) %>%
              slice(1:10),
            aes(label = Congress)) +
  scale_size_continuous(range = c(1, 20)) +
  labs(x = "Leverage",
       y = "Studentized residual") +
  theme(legend.position = "none")

# use color and geom_text_repel
ggplot(dahl_augment, aes(hat, student, color = cooksd)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point() +
  ggrepel::geom_text_repel(data = dahl_augment %>%
              arrange(-cooksd) %>%
              slice(1:10),
            aes(label = Congress)) +
  labs(x = "Leverage",
       y = "Studentized residual") +
  theme(legend.position = "none")
```

## Happiness and gender

```{r mosaic-happy}
# Mosaic plot of happiness and education
library(productplots)
data("happy")

happy <- happy %>%
  na.omit
```

## Happiness and gender

```{r happy-sex}
# contingency table
library(descr)
crosstab(happy$happy, happy$sex, prop.t = TRUE, plot = FALSE)
```

## Happiness and gender {.scrollable}

```{r happy-sex-cond}
crosstab(happy$happy, happy$sex, prop.r = TRUE, plot = FALSE)
crosstab(happy$happy, happy$sex, prop.c = TRUE, plot = FALSE)
```

## Happiness and gender

```{r happy-sex-marg}
crosstab(happy$happy, happy$sex, prop.c = TRUE, prop.r = TRUE, plot = FALSE)
```

## Mosaic plot

```{r graphics-mosaicplot}
mosaicplot(~ sex + happy, data = happy)
```

## Mosaic plot

```{r vcd-mosaic}
library(vcd)
mosaic(~ happy + sex, happy)
```

## Mosaic plot {.scrollable}

```{r productplots}
# mosaic plot using productplots
prodplot(happy, ~ happy + sex)

# add color
prodplot(happy, ~ happy + sex) +
  aes(fill = happy) +
  theme(panel.grid = element_blank())

prodplot(happy, ~ happy + marital) +
  aes(fill = happy) +
  theme(legend.position = "none") +
  theme(panel.grid = element_blank())
```

## Proportional bar chart

```{r prop-barchart}
ggplot(happy, aes(marital, fill = happy)) +
  geom_bar(position = "fill")
```

## Dot plot for summary statistics

```{r oj}
library(ISLR)

OJ_sum <- OJ %>%
  select(ends_with("MM"), ends_with("CH")) %>%
  gather(var, value) %>%
  group_by(var) %>%
  summarize(mean = mean(value),
            sd = sd(value),
            min = min(value),
            max = max(value),
            n = n())

# print the table
kable(OJ_sum)
```

----

```{r oj-dot}
# plot using a single dot plot
ggplot(OJ_sum, aes(x = fct_reorder(var, mean), y = mean)) +
  geom_linerange(aes(ymin = mean - 2 * sd,
                      ymax = mean + 2 * sd),
                  linetype = 2,
                 size = .25) +
  geom_linerange(aes(ymin = mean - sd,
                      ymax = mean + sd),
                  size = 1) +
  geom_point() +
  coord_flip() +
  labs(x = NULL,
       y = NULL)
```

----

```{r oj-dodge}
# dodge based on OJ brand
OJ_sum %>%
  separate(var, into = c("var", "brand"), -3, remove = TRUE) %>%
  ggplot(aes(x = fct_reorder(var, mean), y = mean, color = brand)) +
  geom_linerange(aes(ymin = mean - 2 * sd,
                      ymax = mean + 2 * sd),
                  linetype = 2,
                 size = .25,
                 position = position_dodge(width = 0.5)) +
  geom_linerange(aes(ymin = mean - sd,
                      ymax = mean + sd),
                  size = 1,
                 position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5)) +
  coord_flip() +
  labs(x = NULL,
       y = NULL,
       color = "Brand")
```

----

```{r oj-facet}
# facet based on OJ brand
OJ_sum %>%
  separate(var, into = c("var", "brand"), -3, remove = TRUE) %>%
  ggplot(aes(x = fct_reorder(var, mean), y = mean)) +
  facet_grid(. ~ brand) +
  geom_linerange(aes(ymin = mean - 2 * sd,
                      ymax = mean + 2 * sd),
                  linetype = 2,
                 size = .25) +
  geom_linerange(aes(ymin = mean - sd,
                      ymax = mean + sd),
                  size = 1) +
  geom_point() +
  coord_flip() +
  labs(x = NULL,
       y = NULL,
       color = "Brand")
```

## Single variable model {.scrollable}

```{r lm-plot}
library(coefplot)

model <- lm(hwy ~ displ + I(displ^2), data = mpg)

# in a table
tidy(model) %>%
  kable

# in a plot
coefplot(model)
```

## Multiple linear models {.scrollable}

```{r lm-mult, results = "asis"}
displ <- lm(hwy ~ displ + I(displ^2), data = mpg)
displ_cyl <- lm(hwy ~ displ + I(displ^2) + cyl, data = mpg)
displ_cyl_drv <- lm(hwy ~ displ + I(displ^2) + cyl + drv, data = mpg)
displ_cyl_drv_class <- lm(hwy ~ displ + I(displ^2) + cyl + drv + class, data = mpg)

# in a table
library(stargazer)
stargazer(displ, displ_cyl, displ_cyl_drv, displ_cyl_drv_class,
          type = "html",
          title = "Regression models on highway fuel efficiency",
          covariate.labels = c("Engine displacement", "$\\text{Engine displacement}\\^2$",
                               "Number of cylinders", "Front wheel drive",
                               "Rear wheel drive", "Compact", "Midsize",
                               "Minivan", "Pickup", "Subcompact", "SUV",
                               "Constant"),
          dep.var.caption = "",
          dep.var.labels.include = FALSE,
          align =TRUE)
```

## Multiple linear models {.scrollable}

```{r lm-mult-plot}
# in a plot
multiplot(displ, displ_cyl, displ_cyl_drv, displ_cyl_drv_class,
          title = "Regression models on highway fuel efficiency",
          xlab = NULL,
          ylab = NULL,
          newNames = c("displ" = "Engine displacement",
                       "I(displ^2)" = "Engine displacement^2",
                       "cyl" = "Number of cylinders",
                       "drvf" = "Front wheel drive",
                       "drvr" = "Rear wheel drive",
                       "classcompact" = "Compact",
                       "classmidsize" = "Midsize",
                       "classminivan"   = "Minivan",
                       "classpickup" = "Pickup",
                       "classsubcompact" = "Subcompact",
                       "classsuv" = "SUV")) +
  scale_color_discrete(labels = c("Model 1", "Model 2", "Model 3", "Model 4"))

multiplot(displ, displ_cyl, displ_cyl_drv, displ_cyl_drv_class,
          single = FALSE,
          names = c("Model 1", "Model 2", "Model 3", "Model 4"),
          title = "Regression models on highway fuel efficiency",
          xlab = NULL,
          ylab = NULL,
          newNames = c("displ" = "Engine displacement",
                       "I(displ^2)" = "Engine displacement^2",
                       "cyl" = "Number of cylinders",
                       "drvf" = "Front wheel drive",
                       "drvr" = "Rear wheel drive",
                       "classcompact" = "Compact",
                       "classmidsize" = "Midsize",
                       "classminivan"   = "Minivan",
                       "classpickup" = "Pickup",
                       "classsubcompact" = "Subcompact",
                       "classsuv" = "SUV")) +
  theme(legend.position = "none")
```

## Generalized linear models

```{r titanic_data, message = FALSE}
library(titanic)
titanic <- titanic_train %>%
  as_tibble()

titanic %>%
  head() %>%
  knitr::kable()
```

## Probability of survival {.scrollable}

```{r titanic-age}
survive_age <- glm(Survived ~ Age, data = titanic, family = binomial)
tidy(survive_age)
coefplot(survive_age)

# generate predicted values
library(rcfss) # need to convert log-odds to probabilities
titanic_age <- titanic %>%
  data_grid(Age) %>%
  add_predictions(survive_age) %>%
  # predicted values are in the log-odds form - convert to probabilities
  mutate(prob = logit2prob(pred))

p_age <- ggplot(titanic_age, aes(Age, prob)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Probability of surviving the Titanic",
       subtitle = "Age",
       y = "Predicted probability of survival",
       color = "Sex")
p_age
```

## Odds of survival

```{r titanic-odds}
survive_age_pred <- titanic_age %>%
  mutate(odds = prob2odds(prob))

ggplot(survive_age_pred, aes(Age, odds)) +
  geom_line(color = "blue", size = 1) +
  labs(x = "Age",
       y = "Odds of surviving the Titanic")
```

----

```{r titanic-age-woman}
survive_age_woman <- glm(Survived ~ Age + Sex, data = titanic,
                         family = binomial)

titanic_age_sex <- titanic %>%
  data_grid(Age, Sex) %>%
  add_predictions(survive_age_woman) %>%
  mutate(pred = logit2prob(pred))

p_age_sex <- ggplot(titanic_age_sex, aes(Age, pred, color = Sex)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Probability of surviving the Titanic",
       subtitle = "Age + Sex",
       y = "Predicted probability of survival",
       color = "Sex")
p_age_sex
```

----

```{r titanic-age-woman-x}
survive_age_woman_x <- glm(Survived ~ Age * Sex, data = titanic,
                           family = binomial)

titanic_age_sex_x <- titanic %>%
  data_grid(Age, Sex) %>%
  add_predictions(survive_age_woman_x) %>%
  mutate(pred = logit2prob(pred))

p_age_sex_x <- ggplot(titanic_age_sex_x, aes(Age, pred, color = Sex)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1)) +
  labs(title = "Probability of surviving the Titanic",
       subtitle = "Age x Sex",
       y = "Predicted probability of survival",
       color = "Sex")
p_age_sex_x
```

----

```{r titanic-all-models, results = "asis"}
# in a table
stargazer(survive_age, survive_age_woman, survive_age_woman_x,
          type = "html",
          title = "Probability of surviving the Titanic",
          covariate.labels = c("Age", "Male", "Age x Male", "Constant"),
          dep.var.caption = "",
          dep.var.labels.include = FALSE,
          align = TRUE)
```

----

```{r titanic-all-models-coefplot}
# in a coefficient plot
multiplot(survive_age, survive_age_woman, survive_age_woman_x,
          title = "Probability of surviving the Titanic",
          xlab = "Parameter (log-odds)",
          ylab = NULL,
          newNames = c("Sexmale" = "Male",
                       "Age:Sexmale" = "Age x Male")) +
  scale_color_discrete(labels = c("Age", "Age + Sex", "Age x Sex"))
```

----

```{r titanic-all-models-probplot}
# in a predicted probability plot
bind_rows(list("Age" = titanic_age %>%
                 mutate(pred = prob),
               "Age + Sex" = titanic_age_sex,
               "Age x Sex" = titanic_age_sex_x), .id = "id") %>%
  mutate(Sex = ifelse(is.na(Sex), "both", Sex)) %>%
  # plot the two models
  ggplot(aes(Age, pred, color = Sex, linetype = id)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = c("#1b9e77", "#d95f02", "#7570b3")) +
  labs(title = "Probability of surviving the Titanic",
       x = "Age",
       y = "Predicted probability of survival",
       color = "Sex",
       linetype = "Model")
```

