---
title: "Assignment 3"
author: "Brian Pennington"
date: "May 10, 2017"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)
library(readxl)
library(plotrix)
library(plotly)
library(shiny)
library(knitr)

import_gapminder <- function(filename, inc_name = NA){
  # import file
  indicator <- read_excel(filename)
  
  # rename first column to country, store indicator name for later
  inc_fullname <- names(indicator)[[1]]
  names(indicator)[[1]] <- "country"
  
  # tidy data frame and add indicator name as variable
  indicator <- indicator %>%
    gather(year, value, -1, convert = TRUE) %>%
    mutate(value = as.numeric(value),
           variable = ifelse(!is.na(inc_name), inc_name, inc_fullname)) %>%
    select(country, year, variable, value)
}  


arms_import <- import_gapminder("Arms imports.xlsx", inc_name = "arms import")
arms_export <- import_gapminder("Arms exports.xlsx", inc_name = "arms export") 
student <- import_gapminder("expenditure primary.xlsx", inc_name = "%GDP on Students") 
army_perc <- import_gapminder("indicator army_percent.xlsx",inc_name = "%Army of Workforce")
army_total <-import_gapminder("indicator army_total.xlsx", inc_name = "Army Total") 
ed_aid <- import_gapminder("indicator_education aid (% of total aid).xlsx", inc_name = "Ed Aid") 
GDP <- read.csv("indicator gapminder gdp_per_capita_ppp.csv")

GDP <- GDP %>%
  gather(`X1800`:`X2015`, key= "year", value= "GDP")

as.character(GDP$year)
year <- stringr::str_sub(GDP$year, 2, 5) %>%
  as.integer() %>%
  as.data.frame() 

GDP$year <- NULL

GDP <- bind_cols(GDP, year)
as.data.frame(GDP)

colnames(GDP) <- c("country", "GDP", "year")

data <- bind_rows(arms_import, arms_export, student, army_perc, army_total, ed_aid)

data <- spread(data, key= variable, value= value)

se <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))

```



With the increase prevalence for data in every facet of our days, interactivity, pertaining to graphics, is becoming a necessity. Interactivity allows for one’s audience to familiarize themselves with the data, check assumptions, and explore a narrative even if the audience lacks familiarity with subject or the knowledge base to dissect the data substantively. For this project, we attempted to examine the import and export of arms over the past half century. We attempted a semi-structured narrative for individuals to explore the relation between the global market exchange of arms over time. The reason this is semi-structured is because the default shiny application has a narrative in and of itself, but the narrative is relatively brief before individuals can begin to explore the data on their own. The interactivity portion of the graphic allows for the reader to explore the countries they may be interested in during the time period they may be interested in. It then facets and creates graphics that provide either this global average or country specific values, along with a smoothing line for cleaner interpretations. . 
The story behind the visualization is that there tends to be significantly higher amount of arms that are exported, as opposed to imported. When we look at the ratio of exports to imports, there tends to be about 2.5 times more exports than imports. However, there is significantly higher variability in the exportation of weapons than in the importation of them, but what is also fascinating is that as exportation decreases, importation increases. Finally, it my own examination, it is fascinating to notice that different trends of countries and how they utilize their manufacturing of arms. For example, we notice an exponential increase of Russia exporting arms in the later part of the 2000’s and early part of 2010. This would seem to coincide with Russia’s invasion of Georgia and potential arming of Middle Eastern and other Eastern European nations.  It would seem that there is an enormous industry in the manufacturing and distribution of weapons, and the historical perspective allows for us to notice large, macro trends over the past 50 years. It is eerily similar to George Orwell’s 1984 in explaining how war is no longer about actual conflict, bur rather used for the continued and systematic production of material goods. 
In analyzing and critiquing this visualization, it seems appropriate to begin with what is missing or what else could be added. We could use gganimate or plotly to add detail in order to clarify the changes that are occurring through time. There are also small level changes that could be added such as outlining the country with the highest or lowest number of exchanges. Finally, this data seems most important within the context of a larger narrative.  This particular data seems to more interesting within the context of potentially a fiscal policy narrative, UN/global issues, or a discourse on human nature within a historical framework. 
The visualization uses data from the World Bank, which is a reliable source. The data is presented in a way that attempts to reduce bias such as making interpretation intuitive without putting forth one particular narrative. However, between the missing values and the interactivity, there is plenty of room for the readers to make biased judgements or inferential leaps.  The interactivity portion of the visualization   allows for people to make comparisons, but they may not be as accurate or fine as they could be, which is where the use of gganimate or plotly would be valuable. The functions and computations behind the visualization should perform the correct operations, but the overall functionality might be limited by the scale of the variables, which is why this project may fit better within a larger narrative context. The graphic is aesthetically pleasing. It could possibly be tweaked to be prettier, but the aesthetics do enhance graphic, as opposed to obscuring it. The aspect of this graphic that stands out the most to be is how insightful it is, but I believe they true novice insights come into how we take this data and pair it with the other data sets such as GDP, education, conflicts occurring, and etc. 


```{r, echo= FALSE, warning=FALSE}
  ui <- fluidPage(titlePanel("The Global Arms Market: Import and Exports of Arms over the Half Century"),
                  sidebarLayout(
                    sidebarPanel(
                      sliderInput(
                        "yearInput",
                        "Year",
                        min = 1955,
                        max = 2010,
                        value = c(1955, 2010)
                      ),
                      uiOutput("armInput"),
                      selectInput(
                        "compareInput",
                        "Compare: Make sure to select only a few countries to compare at a time",
                        choices = c("Yes", "No"),
                        selected = "No",
                        multiple = FALSE
                      )),
                    mainPanel(plotOutput("Average"),
                              plotOutput("Ratio")
                  )))
```


``````{r, echo= FALSE, warning=FALSE}
server <- function(input, output) {
    
    output$armInput <- renderUI({
      selectizeInput(
        "countryInput",
        "Country",
        choices = data$country,
        multiple = TRUE
      )
    })
    
    output$Average <- renderPlot({
      
      if(!is.null(input$countryInput)){
        data <- filter(data, country %in% input$countryInput)
      }
      
      data <- data %>%
        group_by(year) %>%
        mutate(avg_imp = mean(`arms import`, na.rm = TRUE)) %>%
        mutate(imp_high = mean(`arms import`, na.rm = TRUE) + 1.96*se(`arms import`)) %>%
        mutate(imp_low = mean(`arms import`, na.rm = TRUE) - 1.96*se(`arms import`)) %>%
        mutate(avg_exp = mean(`arms export`, na.rm = TRUE)) %>%
        mutate(exp_high = mean(`arms export`, na.rm = TRUE) + 1.96*se(`arms export`)) %>%
        mutate(exp_low = mean(`arms export`, na.rm = TRUE) - 1.96*se(`arms export`)) %>%
        mutate(impexp = `arms export`/`arms import`) %>%
        mutate(mean_ratio = mean(impexp, na.rm = TRUE)) %>%
        filter(
          year >= input$yearInput[1],
          year <= input$yearInput[2])
      
      if(input$compareInput == "No"){
        ggplot(data, aes(x = year)) +
        geom_line(aes(y = avg_imp, color = "import")) +
        geom_line(aes(y = avg_exp, color = "export")) +
        geom_ribbon(aes(ymin = imp_low, ymax= imp_high), alpha = .1, fill = "blue") +
        geom_ribbon(aes(ymin = exp_low, ymax = exp_high), alpha = .1, fill = "red") +
        geom_smooth(aes(y = avg_imp), color = "black") +
        geom_smooth(aes(y = avg_exp), color = "black")
    } else{
      ggplot(data, aes(x = year)) +
        facet_wrap(~country) +
        geom_line(aes(y = `arms import`, color = "import")) +
        geom_line(aes(y = `arms export`, color = "export")) +
        geom_smooth(aes(y = `arms import`), color = "black") +
        geom_smooth(aes(y = `arms export`), color = "black")
    }
  })
    
    
    output$Ratio <- renderPlot({
      if(!is.null(input$countryInput)){
        data <- filter(data, country %in% input$countryInput)
      }
      
      data <- data %>%
        group_by(year) %>%
        mutate(impexp = mean(`arms export`,na.rm = TRUE)/mean(`arms import`, na.rm = TRUE)) %>%
        mutate(indavg = `arms export`/`arms import`) %>%
        mutate(mean_ratio = mean(impexp, na.rm = TRUE)) %>%
        filter(
          year >= input$yearInput[1],
          year <= input$yearInput[2])
      
      if(input$compareInput == "No"){
        ggplot(data, aes(x = year)) +
          geom_line(aes(y= mean_ratio, color = "export-import ratio")) +
          geom_smooth(aes(y =mean_ratio))
      } else {
        ggplot(data, aes(x = year)) +
          geom_line(aes(y= indavg, color = "export-import ratio")) +
          geom_smooth(aes(y =indavg)) +
          facet_wrap(~country)
      }
  })
}
```


``````{r, echo= FALSE, warning=FALSE}
shinyApp(ui = ui, server = server)
```
  